---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

#One-Way ANOVA {#chapter3}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

In Chapter \@ref(chapter2), tools for comparing the means of two groups
were considered. More generally, these methods are used for a quantitative
response and a categorical explanatory variable (group) which had two and only
two levels. The full prisoner rating data set actually contained three groups 
(Figure \@ref(fig:Figure3-1) with *Beautiful*, *Average*, and *Unattractive* 
rated pictures randomly assigned to the subjects for sentence ratings. In a 
situation with more than two groups, we have two choices. First, we could rely 
on our two group comparisons, performing tests for every possible pair 
(*Beautiful* vs*Average*, *Beautiful* vs *Unattractive*, and *Average* vs
*Unattractive*). We spent Chapter \@ref(chapter2) doing inferences for differences
between *Average* and *Unattractive*. The other two comparisons would lead us to
initially end up with three p-values and no direct answer about our initial 
question of interest -- is there some overall difference in the average sentences
provided across the groups? In this chapter, we will learn a new method, called 
***Analysis of Variance***, or ***One-Way ANOVA*** since there is just 
one^[In Chapter \@ref(chapter4), methods are discussed for when there are two
categorical explanatory variables that is called the Two-Way ANOVA and related 
ANOVA tests are used in Chapter \@ref(chapter8) for working with extensions of 
these models.] grouping variable. After we perform our One-Way ANOVA test for 
overall evidence of a difference, we will revisit the comparisons similar to those
considered in Chapter \@ref(chapter2) to get more details on specific differences 
among *all* the pairs of groups -- what we call  ***pair-wise comparisons***.
An issue is created when you perform many tests simultaneously and we will augment 
our previous methods with an adjusted method for pairwise comparisons to make our
results valid called  ***Tukey's Honest Significant Difference***. 

To make this more concrete, we return to the original MockJury data, making
side-by-side boxplots and beanplots (Figure \@ref(fig:Figure3-1) as well 
summarizing the sentences for the three groups using ``favstats``. 

(ref:fig3-1) Boxplot and beanplot of the sentences (years) for the three 
treatment groups.

```{r Figure3-1,fig.cap="(ref:fig3-1)",message=F,warning=F}
require(heplots)
require(mosaic)
data(MockJury)
par(mfrow=c(1,2))
boxplot(Years~Attr,data=MockJury)
require(beanplot)
beanplot(Years~Attr,data=MockJury,log="",col="bisque",method="jitter")
favstats(Years~Attr,data=MockJury)
```

There are slight differences in the sample sizes in the three groups with 37
*Unattractive*, 38 *Average* and 39 *Beautiful* group responses, providing a
data set has a total sample size of $N=114$. The *Beautiful* and *Average* groups 
do not appear to be very different with means of 4.33 and 3.97 years. In 
Chapter \@ref(chapter2), we found moderate evidence regarding the difference in
*Average*and *Unattractive*. It is less clear whether we might find evidence of a
difference between *Beautiful* and *Unattractive* groups since we are comparing 
means of 5.81 and 4.33 years. All the distributions appear to be right skewed 
with relatively similar shapes. The variability in *Average* and *Unattractive* 
groups seems like it could be slightly different leading to an overall concern of
whether the variability is the same in all the groups. 

## Situation {#section3-1}

## Linear model for One-Way ANOVA (cell-means and reference-coding) {#section3-2}

## One-Way ANOVA Sums of Squares, Mean Squares, and F-test {#section3-3}

## ANOVA model diagnostics including QQ-plots {#section3-4}

## Guinea pig tooth growth One-Way ANOVA example {#section3-5}

## Multiple (pair-wise) comparisons using Tukey's HSD and the compact letter display {#section3-6}

## Pair-wise comparisons for Prisoner Rating data {#section3-7}

## Chapter Summary {#section3-8}

## Summary of important R code {#section3-9}

## Practice problems {#section3-10}