---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

#One-Way ANOVA {#chapter3}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

In Chapter \@ref(chapter2), tools for comparing the means of two groups
were considered. 

## Situation {#section3-1}

## Linear model for One-Way ANOVA (cell-means and reference-coding) {#section3-2}

## One-Way ANOVA Sums of Squares, Mean Squares, and F-test {#section3-3}

## ANOVA model diagnostics including QQ-plots {#section3-4}

## Guinea pig tooth growth One-Way ANOVA example {#section3-5}

## Multiple (pair-wise) comparisons using Tukey's HSD and the compact letter display {#section3-6}

## Pair-wise comparisons for Prisoner Rating data {#section3-7}

## Chapter Summary {#section3-8}

## Summary of important R code {#section3-9}

## Practice problems {#section3-10}