---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Two-Way ANOVA {#chapter4}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

## Situation {#section4-1}

## Designing a two-way experiment and visualizing results	{#section4-2}

## Two-Way ANOVA models and hypothesis tests	{#section4-3}

## Guinea pig tooth growth analysis with Two-Way ANOVA	{#section4-4}

## Observational study example: The Psychology of Debt	{#section4-5}

## Pushing Two-Way ANOVA to the limit: Un-replicated designs	{#section4-6}

## Chapter summary	{#section4-7}

## Important R code	{#section4-8}

## Practice problems	{#section4-9}
