---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Chi-square tests {#chapter5}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

## Situation, contingency tables, and plots	{#section5-1}

## Homogeneity Test Hypotheses	{#section5-2}

## Independence Test Hypotheses	{#section5-3}

## Models for R by C tables	{#section5-4}

## Permutation tests for the X2 statistic	{#section5-5}

## Chi-square distribution for the X2 statistic	{#section5-6}

## Examining residuals for the source of differences	{#section5-7}

## General Protocol for X2 tests	{#section5-8}

## Political Party and Voting results: Complete Analysis	{#section5-9}

## Is cheating and lying related in students?	{#section5-10}

## Analyzing a stratified random sample of California schools	{#section5-11}

## Chapter summary	{#section5-12}

## Review of Important R commands	{#section5-13}

## Practice problems	{#section5-14}

