---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Correlation and Simple Linear Regression {#chapter6}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

## Relationships between two quantitative variables	{#section6-1}

## Estimating the correlation coefficient	{#section6-2}

## Relationships between variables by groups	{#section6-3}

## Inference for the correlation coefficient (Optional Section)	{#section6-4}

## Are tree diameters related to tree heights?	{#section6-5}

## Describing relationships with a regression model	{#section6-6}

## Least Squares Estimation	{#section6-7}

## Measuring the strength of regressions: R2	{#section6-8}

## Outliers: leverage and influence	{#section6-9}

## Residual diagnostics – setting the stage for inference	{#section6-10}

## Old Faithful discharge and waiting times {#section6-11}

## Chapter summary	{#section6-12}

## Important R code	{#section6-13}

## Practice problems	{#section6-14}

