---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Correlation and Simple Linear Regression {#chapter6}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

## Relationships between two quantitative variables	{#section6-1}

The independence test in Chapter \@ref(chapter5) provided a technique for assessing evidence of a
relationship between two categorical variables. The terms ***relationship*** and ***association***
are synonyms that, in statistics, imply that values on one variable tend to occur more often with
some other values of the other variable or that knowing
something about the level of one variable provides information about the
patterns of values on the other variable. These terms are not specific to the
"form" of the relationship -- any pattern (strong or weak, negative or positive, 
easily explained or complicated) satisfy the definition. There are two other
aspects to using these terms in a statistical context. First, they are not
directional -- an association between $x$ and $y$ is the same as saying there is an
association between $y$ and $x$. Second, they are not causal unless the levels of
the one of the variables are randomly assigned in an experimental context. We add
to this terminology the idea of correlation between variables $x$ and $y$. 
***Correlation***, in most statistical contexts, is a measure of the specific type 
of relationship between the variables: the **linear relationship between two 
quantitative variables**^[There are measures 
of correlation between categorical variables but when
statisticians say correlation they mean correlation of quantitative variables. 
If they are discussing correlations of other types, they will make that clear.]. 
So as we start to review these ideas from your previous statistics course, 
remember that associations and relationships are more general than correlations
and it is possible to have no correlation where there is a strong relationship
between variables. "Correlation" is used colloquially as a synonym for
relationship but we will work to reserve it for its more specialized usage here
to refer to the linear relationship.

Assessing and then modeling relationships between quantitative variables drives
the rest of the chapters, 
so we should get started with some motivating examples to start to think about
what relationships between quantitative variables "look like"... To motivate
these methods, we will start with a study of the effects of beer consumption on
blood alcohol levels (*BAC*, in grams of alcohol per deciliter of blood). A group 
of $n=16$ student volunteers at The Ohio State University drank a
randomly assigned number of beers^[Some of the details of this study have been lost, 
so we will assume that the
subjects were randomly assigned and that a beer means a can of beer and that
the beer was of regular strength. We don't know if any of that is actually true. 
I have been asking the Dean of Students to repeat this study as an educational
activity at MSU with no success yet.]. 
Thirty minutes later, a police officer measured their *BAC*. Your instincts, especially
as well-educated college students with some chemistry knowledge, should inform
you about the direction of this relationship -- that there is a ***positive 
relationship*** between ``Beers`` and ``BAC``. In other words,  **higher values of
one variable are associated with higher values of the other**. Similarly, 
lower values of one are associated with lower values of the other. 
In fact there are online calculators that tell you how much your *BAC* increases
for each extra beer consumed (for example:
http://www.craftbeer.com/beer-studies/blood-alcohol-content-calculator
if you plug in 1 beer). The increase
in $y$ (``BAC``) for a 1 unit increase in $x$ (here, 1 more beer) is an example of 
***slope coefficient*** that is applicable if the relationship between the 
variables is linear and something that will be
fundamental in what we will call ***simple linear regression model***. In a
simple linear regression model (simple means that there is only one explanatory
variable) the slope is the expected change in the mean response for a one unit
increase in the explanatory variable. You could also use the *BAC* calculator and
the models that we are going to develop to pick a total number of beers you
will consume and get a predicted *BAC*.

Before we get to the specifics of this model and how we measure correlation, we 
should explore the relationship between ``Beers`` and ``BAC`` in a scatterplot. 
Figure \@ref(fig:Figure6-1) shows a ***scatterplot*** of the results that display 
the expected positive relationship. Scatterplots display the response pairs for 
the two quantitative variables with the
explanatory variable on the x-axis and the response variable on the y-axis. The
relationship between ``Beers`` and ``BAC`` appears to be relatively linear but
there is possibly more variability than one might expect. For example, for
students consuming 5 beers, their *BAC*s range from 0.05 to 0.10. If you look
at the online *BAC* calculators, you will see that other factors such as weight,
sex, and beer percent alcohol can impact
the results. We might also be interested in previous alcohol consumption. In
Chapter \@ref(chapter8), we will learn how to estimate the relationship between 
``Beers`` and ``BAC`` after correcting for those "other variables" using 
***multiple linear regression***, where we incorporate more than one
quantitative explanatory variable into the linear model (somewhat like in the
2-Way ANOVA). Some of this variability might be hard or impossible to explain
regardless of the other variables available and is considered unexplained variation
and goes into the residual errors in our models, just like in the ANOVA models. 
To make scatterplots as in Figure \@ref(fig:Figure6-1), we simply use 
``plot(y~x, data=...)``.

```{r echo=F}
set.seed(5678)
```

(ref:fig6-1) Scatterplot of beers consumed versus BAC.

```{r Figure6-1,fig.cap="(ref:fig6-1)",warning=F,message=F}
BB <- read.csv("http://www.math.montana.edu/courses/s217/documents/beersbac.csv")
plot(BAC~Beers, data=BB)
```

There are a few general things to look for in scatterplots:

1. **Assess the** $\underline{\textbf{direction of the relationship}}$ -- is it 
positive or negative? 

2. **Consider the** $\underline{\textbf{strength of the relationship}}$. 
The general idea of assessing strength visually is about how hard or easy it is 
to see the pattern. If it is hard to see a pattern, then it is weak. If it is easy to
see, then it is strong. 

3. **Consider the** $\underline{\textbf{linearity of the relationship}}$. Does it 
appear to curve or does it follow a relatively straight line? Curving relationships are
called ***curvilinear*** or ***nonlinear***  and can be strong or
weak just like linear relationships -- it is all about how tightly the
points follow the pattern you identify. 

4. **Check for** $\underline{\textbf{unusual observations -- outliers}}$ -- by looking 
for points that don't follow the overall pattern. Being large in $x$ or $y$
doesn't mean that the point is an outlier. Being unusual relative to the overall
pattern makes a point an outlier in this setting.

5. **Check for** $\underline{\textbf{changing variability}}$ in one variable based 
on values of the other variable. This will tie into a constant variance assumption 
later in the regression models.

6. **Finally, look for** $\underline{\textbf{distinct groups}}$ in the scatterplot.
This might suggest that observations from two populations, say males and females, 
were combined but the relationship between the two quantitative variables
might be different for the two groups.
 
Going back to Figure \@ref(fig:Figure6-1) it appears that there is a 
moderately strong linear
relationship between ``Beers`` and ``BAC`` -- not weak but with some variability
around what appears to be a straight-line relationship. There might even be a
hint of a nonlinear relationship in the higher beer values. There are no clear
outliers because the observation at 9 beers seems to be following the overall
pattern fairly closely. There is little evidence of non-constant variance
mainly because of the limited size of the data set -- we'll check this with
better plots later. And there are no clearly distinct groups in this plot, mainly
because the # of beers was randomly assigned. These data have one more
interesting feature to be noted -- that subjects managed to consume 8 or 9
beers. This seems to be a large number. I have never been able to trace this
data set to the original study so it is hard to know if (1) they had this study
approved by a human subjects research review board to make sure it was "safe", 
(2) every subject in the study was able to consume their randomly assigned
amount, and (3) whether subjects were asked to show up to the study with *BAC*s
of 0. We also don't know the alcohol concentration of the beer consumed or
volume. But it is still a fun example to start these methods with...

In making scatterplots, there is always a choice of a variable for the 
x-axis and the y-axis. It is our
convention to put explanatory or independent variables (the ones used to
explain or predict the responses) on the x-axis. In studies where the subjects are
randomly assigned to levels of a variable, this is very clearly an explanatory
variable, and we can go as far as making causal inferences with it. In
observational studies, it can be less clear which variable explains which. In
these cases, make the most reasonable choice based on the observed variables but
remember that, when the direction of relationship is unclear, you could have
switched which axes and implication of which variable explains which.
   
## Estimating the correlation coefficient	{#section6-2}

In terms of quantifying relationships between variables, we start with 
the correlation coefficient, a
measure that is the same regardless of your choice of which variable is
considered explanatory or response. We measure the strength and direction of
linear relationships between two quantitative variables using 
***Pearson'?'s r*** or ***Pearson'?'s Product Moment Correlation Coefficient***.
For those who really like acronyms, Wikipedia even suggests calling it 
the PPMCC. However, 
its use is so ubiquitous that the lower case ***r*** or just "correlation
coefficient" are often sufficient to identify that you have used the PPMCC. 
Some of the extra distinctions arise because there are other ways of measuring
correlations in other situations (for example between two categorical
variables), but we will not consider them here. 

The correlation coefficient, ***r***, is calculated as

$$r=\frac{1}{n-1}\Sigma^n_{i=1}\left(\frac{x_i-\bar{x}}{s_x}\right)
\left(\frac{y_i-\bar{y}}{s_y}\right),$$ 

where $s_x$ and $s_y$ are the standard deviations of $x$ and $y$. This 
formula can also be written as

$$r=\frac{1}{n-1}\sum^n_{i=1}z_{x_i}z_{y_i}$$

where $z_{x_i}$ is the z-score (observation minus mean divided by 
standard deviation) for the $i^{th}$ observation on $x$ and $z_{y_i}$
is the z-score for the $i^{th}$ observation on $y$. We won't directly
use this formula, but its contents inform the behavior of ***r***.
First, because it is a sum divided by ($n-1$) it is a bit like
an average -- it combines information across all observations and, like the
mean, is sensitive to outliers. Second, it is a dimension-less measure, meaning
that it has no units attached to it. It is based on z-scores which have units
of standard deviations of $x$ or $y$ so the original units of measurement are
cancelled out going into this calculation. This also means that changing the
original units of measurement, say from Fahrenheit to Celsius or from miles to
km for one or the other variable will have no impact on the correlation. Less
obviously, the formula guarantees that ***r*** is between -1 and 1. It will
attain -1 for a perfect negative linear relationship, 1 for a perfect positive
linear relationship, and 0 for no linear relationship. We are being careful
here to say ***linear relationship*** because you can have a strong nonlinear
relationship with a correlation of 0. For example, consider 
Figure \@ref(fig:Figure6-2). 

(ref:fig6-2) Scatterplot of an amusing relationship that has $r=0$.

```{r Figure6-2,fig.cap="(ref:fig6-2)",echo=F,warning=F,message=F}
x<-seq(from=0,to=20,length.out=20)
y<-(x-mean(x))^2
x<-c(x,5,15)
y<-c(y,220,220)
plot(x,y,col='blue',main="r=0")
```

 
There are some conditions for trusting the results that the 
correlation coefficient provides:

1. Two quantitative variables measured. 

    * This might seem silly, but categorical variables can be coded 
    numerically and a meaningless correlation can be estimated if you 
    are not careful what you correlate. 
    
2. The relationship between the variables is relatively linear.

    * If the relationship is nonlinear, the correlation is meaningless 
    and can be misleading.
    
3. There should be no outliers.

    * The correlation is very sensitive (technically ***not resistant***) 
    to the impacts of certain types of outliers and you should generally 
    avoid reporting the correlation when they are present.
    
    * One option in the presence of outliers is to report the correlation 
    with and without outliers to see how they influence the estimated 
    correlation. 

The correlation coefficient is dimensionless but larger magnitude values
(closer to -1 OR 1) mean stronger linear relationships. A rough interpretation
scale based on experiences working with correlations follows, but this varies
between fields and types of research and variables measured. It depends on the
levels of correlation researchers become used to obtaining, so can even vary
within fields. Use this scale until you develop your own experience:

* $\left|\mathbf{r}\right|<0.3$: weak linear relationship,

* $0.3 < \left|\mathbf{r}\right|<0.7$: moderate linear relationship, 

* $0.7 < \left|\mathbf{r}\right|<0.7$: strong linear relationship, and

* $0.9 < \left|\mathbf{r}\right|<1.0$: very strong linear relationship.  

And again note that this scale only relates to the **linear** aspect of
the relationship between the variables. 

When we have linear relationships between two quantitative variables, 
$x$ and $y$, we can obtain estimated correlations from the ``cor``
function either using ``y~x`` or by running the ``cor`` function^[This 
interface with the ``cor`` function only works after you load the 
``mosaic`` package.] on the entire data set. When you run the ``cor``
function on a data set it produces a ***correlation matrix*** which 
contains a matrix of correlations where you can triangulate the 
variables being correlated by the row and column names, noting
that the correlation between a variable and itself is 1. A matrix of
correlations is useful for comparing more than two variables, discussed below. 

```{r}
require(mosaic)
cor(BAC~Beers, data=BB)
cor(BB)
```

Based on either version of using the function, we find that the correlation
between ``Beers`` and ``BAC`` is estimated to be 0.89. This suggests a 
strong linear relationship between the
two variables. Examples are about the only way to build up enough experience to
become skillful in using the correlation coefficient. Some additional
complications arise in more complicated studies as the next example
demonstrates. 

Gude, Cookson, Greenwood, and Haggerty (2009) explored the relationship 
between average summer
temperature (degrees F) and area burned (natural log of hectares^[The 
natural log ($\log_e$ or $\ln$) is used in statistics so much that the
function in R ``log`` actually takes the natural log and if you want a 
$\log_{10}$ you have to use the function ``log10``. When statisticians 
say log we mean natural log.] = log(hectares)) by wildfires in Montana 
from 1985 to 2007. The ***log-transformation*** is often used to reduce 
the impacts of really large observations on
non-negative (strictly greater than 0) responses with really large observations
(more on ***transformations*** and their impacts on regression models 
in Chapter \@ref(chapter7)). Based on your experiences and before 
analyzing the data, I'm sure
you would assume that summer temperature explains the area burned by wildfires. 
But could it be that more fires are related to having warmer summers? That
second direction is unlikely on a state-wide scale but could apply at a
particular weather station that is near a fire. There is another option -- some
other variable is affecting both variables. For example, drier summers might 
be the real explanatory variable that is related to having both warm summers and lots
of fires. These variables are also being measured over time making them examples
of ***time series***. In statistics, time series data is generally reserved 
for evenly spaced measurements over time, like monthly or yearly measurements. In
this situation, if there are changes over time, they might be attributed to
climate change. So there are really three relationships to explore with the
variables measured here (remembering that the full story might require
measuring even more!): log-area burned versus temperature, temperature versus
year, and log-area burned versus year. 

With more than two variables, we can use the ``cor`` function on all the 
variables and end up getting a matrix of correlations or, simply, the
***correlation matrix***. 

```{r}
mtfires <- read.csv("http://www.math.montana.edu/courses/s217/documents/climateR2.csv")
# natural log transformation of area burned
mtfires$loghectacres <- log(mtfires$hectacres) 

#Cuts the original hectacres data so only log-scale version in data.frame
mtfiresR <- mtfires[,-3] 
cor(mtfiresR)
```

If you triangulate the row and column labels, that cell provides the correlation
between that pair of variables. For example, in the first row (``Year``) 
and the last column (``loghectacres``), you can find that the correlation
coefficient is ***r***=0.362. Note the symmetry in the matrix around the 
diagonal of 1'?'s -- this further illustrates that correlation between 
$x$ and $y$ does not depend on which variable is viewed as the "response".
The estimated correlation
between ``Temperature`` and ``Year`` is -0.004 and the correlation between
``loghectacres`` (*log-hectares burned*) and ``Temperature`` is 0.81. So 
``Temperature`` has almost no linear
change over time. And there is a strong linear relationship between 
``loghectacres`` and ``Temperature``. So it appears that temperatures may 
be related to log-area burned but that the trend over time in both is less 
clear (at least the linear trends). 

The correlation matrix alone is misleading -- we need to explore scatterplots 
to check for nonlinear
relationships, outliers, and clustering of observations that may be distorting
the numerical measure of the linear relationship. The ``pairs.panels``
function from the ``psych`` package (Revelle, 2016) combines the numerical 
correlation information and scatterplots in one display. There are
some options to turn off for the moment but it is an easy function to use to
get lots of information in one place. As in the correlation matrix, you
triangulate the variables for the pairwise relationship. The upper right
diagonal of Figure \@ref(fig:Figure6-3) displays a correlation of 0.36 for 
``Year`` and ``loghectares`` and the lower left panel contains the 
scatterplot with ``Year`` on the x-axis and ``loghectares`` on the y-axis.
The correlation between ``Year`` and ``Temperature`` is really small, both
in magnitude and in display, but appears to be nonlinear (it goes down between
1985 and 1995 and then goes back up), so the correlation coefficient doesn't
mean much here since it just measures the overall linear relationship. We might
say that this is a moderate strength (moderately "clear") curvilinear
relationship. In terms of the underlying climate process, it suggests a
decrease in summer temperatures between 1985 and 1995 and then an increase in
the second half of the data set. 

(ref:fig6-3) Scatterplot matrix of Montana fires data. 

```{r Figure6-3,fig.cap="(ref:fig6-3)",warning=F,message=F}
require(psych) 
pairs.panels(mtfiresR, ellipses=F, scale=T, smooth=F, col=0)
```

As one more example, the Australian Institute of Sport collected data 
on 102 male and 100 female athletes that are available in the ``ais``
data set from the ``alr3`` package (Weisberg, 2005). They measured a 
variety of variables including the athlete'?'s Hematocrit (``Hc``, 
units of percentage of red blood cells in the blood), Body Fat Percentage 
(``Bfat``, units of percentage of total bodyweight), and height (``Ht``,
units of cm). Eventually we might be interested in predicting ``Hc`` 
based on the other variables, but for now the associations are of interest. 

(ref:fig6-4) Scatterplot matrix of athlete data. 

```{r Figure6-4,fig.cap="(ref:fig6-4)",warning=F,message=F}
require(alr3)
data(ais)
aisR <- ais[,c("Ht","Hc","Bfat")]
summary(aisR)
cor(aisR)
pairs.panels(aisR,scale=T,ellipse=F,smooth=F,col=0)
```

``Ht`` (*Height*) and ``Hc`` (*Hematocrit*) have a moderate positive 
relationship that may contain a slightly nonlinearity. It also contains one
clear outlier for a middle height athlete (around 175 cm) with an ``Hc``
of close to 60% (a result that is extremely high). One might wonder about 
whether this athlete has been doping or
if that measurement involved a recording error. We should consider removing
that observation to see how our results might change without it impacting the
results. For the relationship between ``Bfat`` (*bodyfat*) and ``Hc`` 
(*hematocrit*), that same high ``Hc`` value is a clear outlier. There is
also a high ``Bfat`` (*body fat*) athlete (35%) with a somewhat low 
``Hc`` value. This also might be influencing our impressions so we will 
remove both "unusual" values and remake the plot. The two offending
observations were found for individuals numbered 56 and 166 in the data set:

```{r}
aisR[c(56,166),]
```

We can create a reduced version of the data (``aisR2``) by removing those 
two rows using ``[-c(56, 166),]`` and then remake the plot:

(ref:fig6-5) Scatterplot matrix of athlete data with two potential 
outliers removed. 

```{r Figure6-5,fig.cap="(ref:fig6-5)",warning=F,message=F}
aisR2 <- aisR[-c(56,166),] #Removes observations in rows 56 and 166
pairs.panels(aisR2, scale=T, ellipse=F, smooth=F, col=0)
```

After removing these two unusual observations, the relationships between 
the variables are more obvious (Figure \@ref(fig:Figure6-5)). There is a 
moderate strength, relatively linear relationship between *Height* and 
*Hematocrit*. There is almost no relationship between *Height* and 
*Body Fat %* (***r***=-0.20). There is a negative, moderate strength, 
somewhat curvilinear relationship between *Hematocrit* and *Body Fat %*
(***r***=-0.54). As hematocrit increases initially, the body fat 
percentage decreases but at a certain level (around 45% for ``Hc``), the 
body fat percentage seems to
level off. Interestingly, it ended up that removing those two outliers had only
minor impacts on the estimated correlations -- this will not always be the case. 

Sometimes we want to just be able to focus on the correlations, assuming 
we trust that
the correlation is a reasonable description of the results between the
variables. To make it easier to see patterns of positive and negative
correlations, we can employ a different version of the same display from 
the ``corrplot`` package (Wei and Simko, 2016). In this case 
(Figure \@ref(fig:Figure6-6)), it tells much the same story but also allows 
the viewer to easily distinguish both size and direction and read off the 
numerical correlations if desired. 

(ref:fig6-6) Correlation plot of the athlete data with two potential 
outliers removed. Lighter (orange) circle for positive correlations 
and black for negative correlations. 


```{r Figure6-6,fig.cap="(ref:fig6-6)",fig.height=6,fig.width=6}
require(corrplot)
corrplot.mixed(cor(aisR2), col=c("black", "orange"))
```


## Relationships between variables by groups	{#section6-3}

## Inference for the correlation coefficient (Optional Section)	{#section6-4}

## Are tree diameters related to tree heights?	{#section6-5}

## Describing relationships with a regression model	{#section6-6}

## Least Squares Estimation	{#section6-7}

## Measuring the strength of regressions: R2	{#section6-8}

## Outliers: leverage and influence	{#section6-9}

## Residual diagnostics – setting the stage for inference	{#section6-10}

## Old Faithful discharge and waiting times {#section6-11}

## Chapter summary	{#section6-12}

## Important R code	{#section6-13}

## Practice problems	{#section6-14}

