---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Simple linear regression inference {#chapter7}

```{r echo=F,warning=F,message=F}
set.seed(5678)
library(pander)
require(mosaic)
library(knitr)
knitr::opts_chunk$set(cache = TRUE)
options(show.signif.stars = FALSE)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```


## Model	{#section7-1}

## Confidence Interval and Hypothesis tests for the slope and intercept	{#section7-2}

## Bozeman temperature trend	{#section7-3}

## Randomizing inferences for the slope coefficient	{#section7-4}

## Transformations part I: Linearizing relationships	{#section7-5}

## Transformations part II: Impacts on SLR interpretations: log(y), log(x), & both log(y) & log(x)	{#section7-6}

## Confidence Interval for the mean and prediction Intervals for a new observation	270

## Chapter summary	{#section7-7}

## Important R code {#section7-8}

## Practice problems	{#section7-9}

