---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Multiple linear regression {#chapter8}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```

## Going from SLR to MLR {#section8-1}

## Validity conditions in MLR	{#section8-2}

## Interpretation of MLR terms	{#section8-3}

## Comparing multiple regression models	{#section8-4}

## General recommendations for MLR interpretations and VIFs	{#section8-5}

## MLR Inference: Parameter inferences using the t-distribution	{#section8-6}

## Overall F-test in Multiple Linear Regression	{#section8-7}

## Case Study: First year college GPA and SATs	{#section8-8}

## Different intercepts for different groups: MLR with Indicator variables	{#section8-9}

## Additive MLR with more than two groups: Headache example	{#section8-10}

## Different slopes and different intercepts	{#section8-11}

## F-tests for MLR models with quantitative and categorical variables and interactions	{#section8-12}

## AICs for model selection	{#section8-13}

## Forced Expiratory Volume model selection using AICs	{#section8-14}

## Chapter summary	{#section8-15}

## Important R code	{#section8-16}

## Practice problems	{#section8-17}
