---
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
header-includes:
- \usepackage{amsmath}
- \usepackage{color}
---

# Case studies {#chapter9}

```{r echo=F,warning=F,message=F}
set.seed(3234)
library(pander)
require(mosaic)
```


```{r echo=F}
#Color Format
colFmt = function(x, color){
  outputFormat = opts_knit$get("rmarkdown.pandoc.to")
  if(outputFormat == 'latex')
    paste("\\textcolor{",color,"}{",x,"}",sep="")
  else if(outputFormat == 'html')
    paste("<font color='",color,"'>",x,"</font>",sep="")
  else
    x
}
```



## Overview of material covered	{#section9-1}

## The impact of simulated chronic nitrogen deposition on the biomass and N2-fixation activity of two boreal feather moss–cyanobacteria associations	{#section9-2}

## Ants learn to rely on more informative attributes during decision-making	{#section9-3}

## Multi-variate models are essential for understanding vertebrate diversification in deep time	{#section9-4}

## General summary	{#section9-5}






