---
output:
  html_document: default
  pdf_document: default
---

#Chi-square tests




</span>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020916">
4. 9: Is cheating and lying related in students?</a>


</h2>


A study of studentbehavior was performed at a university with a survey of *N*=319 undergraduate students (</span>
``cheating``  data set from the </span>
``poLCA``  package originally published byDayton (1998)). They were asked to answer four questions about their various
academic frauds that involved cheating and lying. Specifically, they were asked
if they had ever lied to avoid taking an exam (*LIEEXAM* with 1 for no and 2 for yes), if they had lied to avoidwith 2 for yes), if they had purchased a term paper to hand in as their own orobtained a copy of an exam prior to taking the exam (*FRAUD* with 2 for yes), and if they had copied answers during anexam from someone near them (*COPYEXAM*with 2 for yes). Additionally, their *GPA*swere obtained and put into categories: (&lt;2. 99, 3. 0 to 3. 25, 3. 26 to 3. 50, 
3. 51 to 3. 75, and 3. 76 to 4. 0). These categories were coded from 1 to 5, respectively. 
Again, the code starts with making sure the variables are treated categorically
by applying the </span>

``factor``  function. 

</span>

```{r}



```


#Survey of students</span>
 **cheating$LIEEXAM&lt;-</span>
(cheating$LIEEXAM)</span>
 **cheating$LIEPAPER&lt;-</span>
(cheating$LIEPAPER)</span>
 **cheating$FRAUD&lt;-</span>
(cheating$FRAUD)</span>
 **cheating$COPYEXAM&lt;-</span>
(cheating$COPYEXAM)</span>
 **cheating$GPA&lt;-</span>
(cheating$GPA)</span>
 **tableplot</span>
sort=</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1033" type="#_x0000_t75" style='width:450. 75pt;height:300. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image099. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=601 height=401src="chapter4_files/image100. jpg" v:shapes="_x0000_i1033">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 4-18: Tableplot of
initial cheating and lying data set. 

</span>
*


&nbsp;
</span>


We can explore someinteresting questions about the relationships between these variables. The
tableplot in Figure 4-18 again helps us to get a general idea of the data set
and to assess some complicated aspects of the relationships between variables. 
For example, the rates of different unethical behaviors seem to decrease with
higher GPA students (but do not completely disappear!). This data set also has
a few missing GPAs that we would want to carefully consider ? which sorts of
students might not be willing to reveal their GPAs? It ends up that these
students did not *admit* to any of theunethical behaviors? Note that we used the </span>
``sort=GPA``  option in the </span>
``tableplot``  function to sort the responses basedon </span>

 to see how *GPA* might explain patterns of unethical behavior. 

</span>

????????????? </span>
Whilethe relationship between GPA and presence/absence of the different behaviors is
of interest, we want to explore the types of behaviors. It is possible to group
the lying behaviors as being a different type of unethical behavior than
obtaining an exam prior to taking it, buying a paper, or copying someone else?s
answers. We want to explore whether there is some sort of relationship between
the lying and copying behaviors ? are those that engage in one type of behavior
more likely to do the other? Or are they independent of each other? This is a
hard story to elicit from the previous plot because there are so many variables
involved. 

</span>


To simplify the results, combining the two groups of variables into the four possible combinations on
each has the potential to simplify the results ? or at least allow exploration
of additional research questions. In the table plot in Figure 4-19, you can see
the four categories for each, starting with no bad behavior of either type
(which is fortunately the most popular response on both variables!). For each
variable, there are students who admitted to one of the two violations and some
that did both. The </span>

``liar``  variable has categories of *None*, *ExamLie*, *PaperLie*, and *LieBoth*. The </span>

``copier``  variable has categories of *None*, *PaperCheat, ExamCheat*, and *PaperExamCheat* (for doing both). The last category for </span>

``copier``  (plotted as a peach(?) color) seems to mostly occur at thetop of the plot which is where the students who had lied to get out of things
reside, so maybe there is a relationship between those two types of behaviors?
On the other hand, for the students who have never lied, quite a few had
cheated on exams. The contingency table can help us dig further into the
hypotheses related to the Chi-square test of Independence that is appropriate
in this situation.  ** 

</span>
** </span>

```{r}



```


(cheating$LIEEXAM, cheating$LIEPAPER)</span>
 **levels</span>
c</span>

(</span>
"None"</span>
, </span>
, </span>

, </span>

)</span>
 ** **cheating$copier&lt;-</span>
interaction</span>
(cheating$FRAUD, cheating$COPYEXAM)</span>
 **levels</span>
(cheating$copier)&lt;-</span>
c</span>
(</span>
"None"</span>
, </span>
, </span>

, </span>

)</span>
 ** **cheatlietable&lt;-</span>
tally</span>
(~liar+copier, </span>
cheating)</span>
 **cheatlietable</span>



##<spanstyle='mso-spacerun:yes'>
?????????? </span>
copier</span>
 **## liar<span style='mso-spacerun:yes'>
??????</span>
None PaperCheat ExamCheat PaperExamCheat</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
None<spanstyle='mso-spacerun:yes'>
????? </span>
207<spanstyle='mso-spacerun:yes'>
????????? </span>
7<spanstyle='mso-spacerun:yes'>
??????? </span>
46<spanstyle='mso-spacerun:yes'>
????????????? </span>
5</span>
 **##<span style='mso-spacerun:yes'>
??</span>
ExamLie<span style='mso-spacerun:yes'>
??? </span>
10<spanstyle='mso-spacerun:yes'>
????????? </span>
1<spanstyle='mso-spacerun:yes'>
???????? </span>
3<spanstyle='mso-spacerun:yes'>
????????????? </span>
2</span>
 **##<span style='mso-spacerun:yes'>
??</span>
PaperLie<span style='mso-spacerun:yes'>
?? </span>
13<spanstyle='mso-spacerun:yes'>
????????? </span>
1<spanstyle='mso-spacerun:yes'>
???????? </span>
4<spanstyle='mso-spacerun:yes'>
????????????? </span>
2</span>
 **##<span style='mso-spacerun:yes'>
??<spanstyle='mso-spacerun:yes'>
??</span>
11<span style='mso-spacerun:yes'>
?????????</span>
1<span style='mso-spacerun:yes'>
???????? </span>
4<spanstyle='mso-spacerun:yes'>
????????????? </span>
2</span>




(cheating, </span>

liar, </span>

c</span>

</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1032" type="#_x0000_t75" style='width:394. 5pt;height:263. 25pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image101. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=526 height=351src="chapter4_files/image101. png" v:shapes="_x0000_i1032">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 4-19: Tableplot of new
variables liar and copier that allow exploration of relationships between
different types of lying and cheating behaviors. </span>
*

</span>

Unfortunately, there werevery few responses in some combinations of categories even with *N*=319. For example, there was only oneresponse each in the combinations for students that copied on papers and lied
to get out of exams, papers, and both. Some other categories were pretty small
as well in the groups that only had one behavior present. To get a higher
number of counts in the combinations, we combined the single behavior only levels
into "*either*" categories and left the*none* and *both* categories for each variable. This creates two new variables
called </span>

 and </span>

``copier2``  (tableplot in Figure 4-20). The codeto create these variables and make the plot follows. 

</span>


<!--[if gte vml 1]>
<v:shape id="Picture_x0020_34" o:spid="_x0000_i1031"
 type="#_x0000_t75" style='width:401. 25pt;height:221. 25pt;visibility:visible;
 mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image102. emz" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=535 height=295src="chapter4_files/image103. png" v:shapes="Picture_x0020_34">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 4-20: Tableplot of
lying and copying variables after combining categories. 

</span>
*


&nbsp;
</span>


```{r}



```

 **cheating$liar2&lt;-cheating$liar</span>
 **levels</span>
c</span>

(</span>
"None"</span>
, </span>
, </span>

, </span>

)</span>
 **cheating$copier2&lt;-cheating$copier</span>
 **levels</span>
(cheating$copier2)&lt;-</span>
c</span>
(</span>
"None"</span>
, </span>
, </span>

, </span>

)</span>
 **tableplot</span>
sort=</span>

select=</span>

(liar2, copier2))</span>
 **cheatlietable&lt;-</span>
(~liar2+copier2, </span>

cheating)</span>
 **cheatlietable</span>



##<spanstyle='mso-spacerun:yes'>
????????????? </span>
copier2</span>
 **## liar2<span style='mso-spacerun:yes'>
????????</span>
None ExamorPaper CopyBoth</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
None<spanstyle='mso-spacerun:yes'>
???????? </span>
207<spanstyle='mso-spacerun:yes'>
????????? </span>
53<spanstyle='mso-spacerun:yes'>
??????? </span>
5</span>
 **##<span style='mso-spacerun:yes'>
??</span>
ExamorPaper<span style='mso-spacerun:yes'>
?? </span>
23<spanstyle='mso-spacerun:yes'>
?????????? </span>
9<span style='mso-spacerun:yes'>
?</span>
??????</span>
4</span>
 **##<span style='mso-spacerun:yes'>
??</span>
LieBoth<span style='mso-spacerun:yes'>
?????? </span>
11<spanstyle='mso-spacerun:yes'>
?????????? </span>
5<spanstyle='mso-spacerun:yes'>
??????? </span>
2</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


This 3x3 table is more manageable andhas few really small cells so we will proceed with the 6+ steps of hypothesis
testing applied to these data using the Independence testing methods (again a
single sample was taken from the population):

</span>


 **1) Hypotheses:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_0$: Lying and copying
behavior are independent in the population of students at this university. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_A$: Lying and copying
behavior are dependent in the population of students at this university. 

</span>


 **2) Validity conditions:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Independence:

</span>


margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Thisassumption is presumed to be met since each subject is measured only once in
the table. No other information suggests a potential issue but we don?t have
much information on how these subjects were obtained. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
All expected cell counts larger than
style='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
**  distribution to find p-values):

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Weneed to generate a table of expected cell counts to check this condition:

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
????????????? </span>
copier2</span>
 **## liar2<span style='mso-spacerun:yes'>
?????????????</span>
None ExamorPaper<span style='mso-spacerun:yes'>
? </span>
CopyBoth</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
None<spanstyle='mso-spacerun:yes'>
??????? </span>
200. 20376<spanstyle='mso-spacerun:yes'>
?? </span>
55. 658307 9. 1379310</span>
 **##<span style='mso-spacerun:yes'>
??</span>
ExamorPaper<span style='mso-spacerun:yes'>
? </span>
27. 19749<spanstyle='mso-spacerun:yes'>
??? </span>
7. 561129  **** </span>
 **##<span style='mso-spacerun:yes'>
??</span>
LieBoth<span style='mso-spacerun:yes'>
????? </span>
13. 59875<spanstyle='mso-spacerun:yes'>
??? </span>
 **** </span>




</div>


<spanstyle='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:
"Times New Roman"'>

&nbsp;
</span>


margin-bottom:0in;margin-left:. 75in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Whenwe request the expected cell counts, there is a warning message (not shown). 

</span>


margin-bottom:0in;margin-left:. 75in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Thereare three expected cell counts (in bold) below 5, so the condition is violated
and a permutation approach should be used to obtain more trustworthy p-values. 

</span>


 **3) Calculate the test statistic:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Use </span>

``chisq. test``  although this table is small enoughto do by hand if you want the practice ? see if you can find a similar answer
to what the function provides:

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
? </span>
Pearson's Chi-squared test</span>
 **## data:<span style='mso-spacerun:yes'>
?</span>
cheatlietable</span>
 **</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
The X<sup>
2</sup>
 statistic is 13. 24. 

</span>


 **4) Find the p-value:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
The parametric p-value is 0. 0102 from
style='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
**  distribution with (3-1)*(3-1) = 4 degrees of freedom that is displayed inFigure 4-21. Remember that this isn?t quite the right distribution for the test
statistic since our expected cell count condition was violated. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1030" type="#_x0000_t75" style='width:231. 4pt;height:154. 15pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image104. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=309 height=206src="chapter4_files/image105. jpg" v:shapes="_x0000_i1030">
<![endif]>
</span>


</span>

style='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Courier New"'>
Figure 4-21: Plot of </span>
</i>
 ***<spanstyle='font-size:12. 0pt;font-family:Symbol;mso-ascii-font-family:"Calibri Light";
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
***  ***<sup>
<spanstyle='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Courier New"'>
2</span>
</sup>
*** *<span style='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Courier New"'>
-distribution with 4 degrees of freedom. 

</span>
*

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
If you want to repeat the p-value
calculation directly:

</span>


```{r}



```


, </span>

4</span>

, </span>
lower. tail=</span>
F)</span>



## [1]0. 01016781</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
But since the expected cell condition
is violated again, we should use permutations as implemented in the following
code with the number of permutations increased to 10, 000 to help get a better
estimate of the p-value since it is possibly close to 0. 05:

</span>


```{r}



```

 **Tobs&lt;-</span>
chisq. test</span>
(</span>
(~liar2+copier2, </span>

cheating))$statistic</span>
 **Tobs</span>



## X-squared </span>
 **##<span style='mso-spacerun:yes'>
? </span>
13. 23844</span>



 </span>

10000</span>
 **Tstar&lt;-</span>
(</span>

NA</span>
, </span>
nrow=</span>
B)</span>
 **for (b in (</span>
:B)){</span>
 **? </span>
Tstar[b]&lt;-</span>
chisq. test</span>
tally</span>

shuffle</span>

data=</span>

cheating))$statistic</span>
 **}</span>
 **hist</span>
(Tstar)</span>
 **abline</span>
(</span>
v=</span>
col=</span>

, </span>

3</span>

(</span>

(Tstar), </span>

"Density curve of Tstar"</span>

lwd=</span>

)</span>
 **abline</span>
v=</span>

col=</span>

, </span>

3</span>

(Tstar, Tobs, </span>

F)</span>




## X-squared </span>
 **##<span style='mso-spacerun:yes'>
??? </span>
0. 0154</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1029" type="#_x0000_t75" style='width:378. 75pt;height:252. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image106. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=505 height=337src="chapter4_files/image106. png" v:shapes="_x0000_i1029">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:115%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 4-22: Plot of
permutation distributions for cheat/lie results with observed value of 13. 24
(bold, vertical line). </span>
*

</span>


&nbsp;
</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
There were 154 of *B*=10, 000 permuted data sets thatproduced larger X<sup>
2</sup>
*?s than the observed as displayed in Figure 4-22, 
so we report that the p-value was 0. 0154 using the permutation approach, which
was slightly larger than the result provided by the parametric method. 

</span>


 **5) Make a decision:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
With a p-value of 0. 0154, we can say
that there is a 1. 5% chance of observing a configuration like ours or more
extreme if the null hypothesis is true. So we should probably reject the null
hypothesis although your standards for evidence might differ ? this is moderate
to strong but not very strong evidence against the null hypothesis. 

</span>


 **6) Write a conclusion:

</span>
** 

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
If we reject the null hypothesis, 
there is enough evidence to conclude that there is a relationship between lying
and copying behavior in the population of students. There is no causal
inference possible here since neither variable was randomly assigned (really
neither is explanatory or response here either) but we can extend the
inferences to the population of students that these were select from that would
be willing to reveal their GPA (see initial discussion related to some
differences in students that wouldn?t answer that question). 

</span>


The standardized residuals can helpus more fully understand this result ? the mosaic plot only had one cell shaded
and so wasn?t needed here. 

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
????????????? </span>
copier2</span>
 **## liar2<span style='mso-spacerun:yes'>
??????????????</span>
None ExamorPaper<span style='mso-spacerun:yes'>
?? </span>
CopyBoth</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
None<spanstyle='mso-spacerun:yes'>
???????? </span>
0. 4803220<spanstyle='mso-spacerun:yes'>
? </span>
-0. 3563200 -1. 3688609</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
ExamorPaper-0. 8048695<span style='mso-spacerun:yes'>
?? </span>
0. 5232734<spanstyle='mso-spacerun:yes'>
? </span>
 **** </span>
 **##<span style='mso-spacerun:yes'>
??</span>
LieBoth<span style='mso-spacerun:yes'>
???? </span>
-0. 7047165<spanstyle='mso-spacerun:yes'>
?? </span>
0. 6271633<span style='mso-spacerun:yes'>
?</span>
1. 7507524</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


There is really one largestandardized residual for the *ExamorPaper*liars and the *CopyBoth* copiers, witha much larger observed value than expected of 2. 48. The only other medium-sized
copiers column with fewer than expected students in the *None* category and more than expected in the *LieBoth* type of lying category. So we are seeing more than expectedthat lied somehow and copied ? we can say this suggests that the people who lie
tend to copy too!

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020917">
4. 10: Analyzing a stratified random sample of California
schools</a>


</h2>


????????????? </span>
Inrecent decades, there has been a push for quantification of school performance
and tying financial punishment and rewards to growth in these metrics. One
example is the API (Academic Performance Index) in California that is based
mainly on student scores on standardized tests. It ranges between 200 and 1000
and year to year changes are of interest to assess "performance" of schools ?
calculated as one year minus the previous year (negative "growth" is also
possible!). Suppose that a researcher is interested in whether the growth
metric might differ between different levels of schools. Maybe it is easier or
harder for elementary, middle, or high schools to attain growth. The researcher
has a list of most of the schools in the state of each level that are using a
data-base that the researcher has access to. In order to assess this question, 
the researcher takes a stratified random sample<a style='mso-footnote-id:ftn6'
href="#_ftn6" name="_ftnref6" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>
, 
selecting </span>
*n<sub>
elementary</sub>
</span>
* = 100 schools from the population of
4421 elementary schools, </span>
*n<sub>
middle</sub>
</span>
* = 50 from the population of 1018
middle schools, and </span>
*n<sub>
high</sub>
</span>
* = 50 from the population of 755 high
schools. These data are available in the </span>

``survey``  package (Lumley, 2014) and the </span>
``api``  and </span>
``apistrat``  data sets. The *growth* (change!) in API scores for the schools between 1999 and
2000 (taken as the year 2000 score minus 1999 score) is used as the response
variable. The boxplot and beanplot of the growth scores are displayed in Figure
4-23. They suggest some differences in the growth rates among the different
levels. There are also a few schools flagged as being possible outliers. 

</span>


```{r}



```

</span>
(survey)</span>
</span>
 **data</span>
(api)</span>
 **tally</span>
(~stype, </span>
data=</span>
</span>
apipop) </span>
#Population counts</span>
</span>


</span>

##<spanstyle='mso-spacerun:yes'>
??? </span>
E<span style='mso-spacerun:yes'>
???</span>
H<span style='mso-spacerun:yes'>
??? </span>
M </span>
</span>
 **## 4421<span style='mso-spacerun:yes'>
? </span>
7551018</span>


</span>

tally</span>
</span>
(~stype, </span>
</span>
data=</span>
</span>
apistrat)</span>
</span>

#Sample counts</span>
</span>


</span>

##<spanstyle='mso-spacerun:yes'>
?? </span>
E<span style='mso-spacerun:yes'>
??</span>
H<span style='mso-spacerun:yes'>
?? </span>
M </span>
</span>
 **## 100<span style='mso-spacerun:yes'>
? </span>
50<spanstyle='mso-spacerun:yes'>
? </span>
50</span>


</span>

boxplot</span>
</span>
(growth~stype, </span>
</span>
data=</span>
</span>
apistrat, </span>
</span>
ylab=</span>
</span>
"Growth"</span>
</span>
, </span>
</span>
ylim=</span>
</span>
c</span>
</span>
(-</span>
</span>
55</span>
</span>
, </span>
</span>
160</span>
</span>
))</span>
</span>
 **beanplot</span>
(growth~stype, </span>
data=</span>
</span>
apistrat , </span>
log=</span>
</span>
""</span>
</span>
, </span>
col=</span>
</span>
"beige"</span>
</span>
, </span>

ylim=</span>
</span>
(-</span>

55</span>
, </span>
160</span>
</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1028" type="#_x0000_t75" style='width:471pt;height:313. 9pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image107. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=628 height=419src="chapter4_files/image108. jpg" v:shapes="_x0000_i1028">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 4-23: Boxplot and
beanplot of the API growth scores by level of school (E for elementary, M for
Middle, and H for High school). 

</span>
*


&nbsp;
</span>


The One-Way ANOVA *F*-test, provided below, suggests
evidence of some difference in the true mean growth scores among the different
p-value&lt;0. 0001). But the residuals from this model displayed in the QQ-Plot inFigure 4-24 contain a slightly long right tail, suggesting a right skewed
distribution for the residuals. In a high-stakes situation such as this, 
reporting results with violations of the assumptions probably would not be
desirable, so another approach is needed. The permutation methods would be
justified here but there is another "simpler" option available using our
Chi-square analysis methods. 

</span>


```{r}



```


data=</span>

apistrat)</span>
 **require</span>
(car)</span>
 **Anova</span>
##<spanstyle='mso-spacerun:yes'>
?????????? </span>
Sum Sq<spanstyle='mso-spacerun:yes'>
? </span>
Df F value<span style='mso-spacerun:yes'>
???</span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## stype<span style='mso-spacerun:yes'>
?????</span>
30370<span style='mso-spacerun:yes'>
?? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
23. 563 6. 685e-10 ***</span>
 **## Residuals 126957 197<spanstyle='mso-spacerun:yes'>
????????????????????? </span>
</span>
 **plot</span>
which=</span>

)</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1027" type="#_x0000_t75" style='width:366. 75pt;height:244. 5pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image109. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=489 height=326src="chapter4_files/image109. png" v:shapes="_x0000_i1027">
<![endif]>
</span>


</span>

<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

One way to get around thenormality assumption is to use a method that does not assume the responses
follow a normal distribution. If we  ***bin***  or cut the quantitative responsevariable into a set of ordered categories and apply a Chi-square test, we can
proceed without concern about the lack of normality in the residuals of the
ANOVA model. To create these bins, a simple idea would be to use the quartiles
to generate the response variable categories, binning the quantitative
responses into groups for the lowest 25%, second 25%, third 25%, and highest
25% by splitting the data at Q<sub>
1$, the Median, and Q<sub>
3</sub>
. In
R, the </span>

 function is available to turn a
quantitative variable into a categorical variable. First, we can use the
information from </span>

``favstats``  to find the cut-points:

</span>

```{r}



```


apistrat)</span>




##<spanstyle='mso-spacerun:yes'>
? </span>
min<span style='mso-spacerun:yes'>
??</span>
Q1 median Q3 max<span style='mso-spacerun:yes'>
?? </span>
mean<spanstyle='mso-spacerun:yes'>
????? </span>
sd<span style='mso-spacerun:yes'>
??</span>
n missing</span>
 **##<span style='mso-spacerun:yes'>
? </span>
-476. 75<span style='mso-spacerun:yes'>
???? </span>
25 48 133 27. 995 28. 1174200<span style='mso-spacerun:yes'>
?????? </span>
0</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The </span>
``cut``  function can provide the binnedvariable if it is provided with the end-points of the desired intervals to
create new categories with those names in a new variable called </span>

``growthcut`` . 

</span>

```{r}



```


breaks=</span>

(-</span>

47</span>
, </span>
6. 75</span>
, </span>
25</span>
, </span>
48</span>
, </span>
133</span>
include. lowest=</span>

T)</span>
 **tally</span>
data=</span>

## </span>
 **## [-47, 6. 75]<span style='mso-spacerun:yes'>
?</span>
(6. 75, 25]<span style='mso-spacerun:yes'>
??? </span>
(25, 48]<spanstyle='mso-spacerun:yes'>
?? </span>
(48, 133] </span>
 **##<span style='mso-spacerun:yes'>
????????</span>
50<span style='mso-spacerun:yes'>
???????? </span>
52<spanstyle='mso-spacerun:yes'>
???????? </span>
49<spanstyle='mso-spacerun:yes'>
???????? </span>
49</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Now that we have acategorical response variable, we need to decide which sort of Chi-squared
analysis to perform. The sampling design determines the correct analysis as
always in these situations. The stratified random sample involved samples from
each of the three populations so a Homogeneity test should be employed. In
these situations, the stacked bar chart provides the appropriate summary of the
data. It also shows us the labels of the categories that the cut function
created in the new </span>

``growthcut``  variable:

</span>

```{r}



```


apistrat)</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1026"
 type="#_x0000_t75" style='width:464. 25pt;height:284. 65pt;visibility:visible;
 mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image110. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=619 height=380src="chapter4_files/image111. jpg" v:shapes="Picture">
<![endif]>
</span>
<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>


</span>
</i>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

Figure 4-25 suggests thatthe distributions of growth scores may not be the same across the levels of the
schools with many more high growth *Elementary*schools than in either the *Middle* or *High* school groups (the "high" growth category
is labeled as (48, 133] providing the interval of growth scores placed in this
category). Similarly, the proportion of the low or negative growth (category of
(-47. 6, 6. 75] for "growth" between -47. 6 and 6. 75) is lowest in *Elementary* schools and highest in the *High* schools. Statisticians often work
across many disciplines and so may not always have the subject area knowledge
to know why these differences exist (just like you might not), but an education
researcher could take this sort of information ? because it is a useful summary
of interesting school-level data ? and generate further insights into why
growth in the API metric may or may not be a good or fair measure of school
performance. 

</span>


????????????? </span>
Ofcourse, we want to consider whether these results can extend to the population
of all California schools. The homogeneity hypotheses for assessing the growth
rate categories across the types of schools would be:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_0$: There is no difference
in the distribution of growth categories across the three levels of schools in
the population of California schools. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_a</sub>
: There is some
difference in the distribution of growth categories across the three levels of
schools in the population of California schools. 

</span>


There might be an issuewith the independence assumption in that schools within the same district might
be more similar to one another and different between one another. Sometimes
districts are accounted for in education research to account for differences in
policies and demographics among the districts. We could explore this issue by
finding district-level average growth rates and exploring whether those vary
systematically but this is beyond the scope of the current exploration. 

</span>


Checking the expectedcell counts gives insight into the assumption for using the </span>
 **<span style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:
minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
** -distribution to find the p-value:

</span>

```{r}



```


data=</span>

apistrat)</span>
 **growthtable</span>



##<spanstyle='mso-spacerun:yes'>
????? </span>
growthcut</span>
 **## stype [-47, 6. 75] (6. 75, 25] (25, 48] (48, 133]</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
E<spanstyle='mso-spacerun:yes'>
???????? </span>
14<spanstyle='mso-spacerun:yes'>
??????? </span>
22<span style='mso-spacerun:yes'>
?????</span>
27<span style='mso-spacerun:yes'>
?????? </span>
37</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
H<spanstyle='mso-spacerun:yes'>
???????? </span>
24<spanstyle='mso-spacerun:yes'>
??????? </span>
18<span style='mso-spacerun:yes'>
??????</span>
5<span style='mso-spacerun:yes'>
??????? </span>
3</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
M<spanstyle='mso-spacerun:yes'>
???????? </span>
12<spanstyle='mso-spacerun:yes'>
??????? </span>
12<span style='mso-spacerun:yes'>
????</span>

?</span>
17<spanstyle='mso-spacerun:yes'>
??????? </span>
9</span>



(growthtable)$expected<span style='mso-spacerun:yes'>
???????</span>
</span>



##<spanstyle='mso-spacerun:yes'>
????? </span>
growthcut</span>
 **## stype [-47, 6. 75] (6. 75, 25] (25, 48] (48, 133]</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
E<spanstyle='mso-spacerun:yes'>
?????? </span>
25. 0<spanstyle='mso-spacerun:yes'>
??????? </span>
26<span style='mso-spacerun:yes'>
??</span>
24. 50<span style='mso-spacerun:yes'>
??? </span>
24. 50</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
H<spanstyle='mso-spacerun:yes'>
?????? </span>
12. 5<spanstyle='mso-spacerun:yes'>
??????? </span>
13<span style='mso-spacerun:yes'>
??</span>
12. 25<span style='mso-spacerun:yes'>
??? </span>
12. 25</span>
 **##<span style='mso-spacerun:yes'>
???? </span>
M<spanstyle='mso-spacerun:yes'>
?????? </span>
12. 5<spanstyle='mso-spacerun:yes'>
??????? </span>
13<span style='mso-spacerun:yes'>
??</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The smallest expected count is12. 065, occurring in three different cells, so the assumptions for using the
parametric approach are met. 

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
? </span>
Pearson's Chi-squared test</span>
 **## data:<span style='mso-spacerun:yes'>
?</span>
growthtable</span>
 **</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The observed test statistic is X<sup>
2</sup>
=38. 67style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
** (6) distribution, the p-value is 0. 0000008. This p-value suggests that weshould reject the null hypothesis. Then we can conclude that there is strong evidence
of some difference in the distribution of API growth of schools among *Elementary*, <i style='mso-bidi-font-style:the population of schools in California between 2000 and 1999. We can concludethat there is evidence of some difference in the population (California
schools) because the schools were randomly selected from all the California
schools but because the level of schools, obviously, cannot be randomly
assigned, we cannot say that level of school causes these differences. 

</span>


The standardized residualscan enhance this interpretation, displayed in Figure 4-26:

</span>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(growthtable)$residuals</span>
</span>




<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<spanstyle='mso-spacerun:yes'>
????? </span>
growthcut</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## stype [-47, 6. 75]<spanstyle='mso-spacerun:yes'>
? </span>
(6. 75, 25]<span style='mso-spacerun:yes'>
???</span>
(25, 48]<span style='mso-spacerun:yes'>
?? </span>
(48, 133]</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
???? </span>
E-2. 2000000 -0. 7844645<span style='mso-spacerun:yes'>
? </span>
0. 5050763<spanstyle='mso-spacerun:yes'>
? </span>
2. 5253814</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
????</span>
H<span style='mso-spacerun:yes'>
? </span>
3. 2526912<spanstyle='mso-spacerun:yes'>
? </span>
1. 3867505 -2. 0714286 -2. 6428571</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
???? </span>
M-0. 1414214 -0. 2773501<span style='mso-spacerun:yes'>
? </span>
1. 3571429-0. 9285714

</span>
</span>



&nbsp;


<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
mosaicplot</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(growthcut~stype, </span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
data=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
apistrat, </span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
shade=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The *Elementary* schools have fewer low/negative growth schools and more
high growth schools than expected under the null hypothesis. The *High* schools have more low growth andfewer higher growth (over 25) schools than expected if there were no difference
in patterns of response across the school levels. The *Middle* school results were closer to the results expected if therewere no differences. 

</span>


<!--[if gte vml 1]>
<v:shape id="Picture_x0020_302"
 o:spid="_x0000_i1025" type="#_x0000_t75" style='width:510. 75pt;height:351pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter4_files/image112. emz" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=681 height=468src="chapter4_files/image113. png" v:shapes="Picture_x0020_302">
<![endif]>
</span>
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:blue;background:
#E1E2E5'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure
4-26: Mosaic plot of the API Growth rate categories versus level school with
shading for size of standardized residuals. 

</span>
</i>



&nbsp;
</span>

The binning ofquantitative variables is not a first step in analyses ? the quantitative
version is almost always preferable. However, this analysis avoided the
violation of the normality assumption that was somewhat problematic for the
ANOVA and still provided useful inferences to the differences in the types of
schools. When one goes from a quantitative to categorical version of a
variable, one loses information (the specific details of the quantitative
responses within each level created) and this almost always will result in a
loss of statistical power of the procedure. In this situation, the p-value from
the ANOVA was of the order 10<sup>
-10</sup>
 while the Chi-square test had a
p-value of order 10<sup>
-7</sup>
. This larger p-value is typical of the loss of
power in going to a categorical response when more information was available. 
In many cases, there are no options but to use contingency table analyses. This
example shows that there might be some situations where "going categorical"
could be an acceptable method for handing situations where an assumption is
violated. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020918">
4. 11: Chapter summary</a>


</h2>


Chi-square tests can begenerally used to perform two types of tests, the Independence and Homogeneity
tests. The appropriate analysis is determined based on the data collection
methodology. The parametric Chi-squared distribution for which these tests are
named is appropriate when the expected cell counts are large enough (related to
having a large enough overall sample). When this condition is violated, the
permutation approach can provide valuable inferences in these situations in
most situations. 

</span>


Data displays of thestacked barchart (Homogeneity) and mosaic plots (Independence) provide a visual
summary of the results that can also be found in contingency tables. You should
have learned how to calculate the X<sup>
2</sup>
 (X-squared) test statistic
based on first finding the expected cell counts. Under certain assumptions, it
will follow a Chi-Squared distribution with (R-1)(C-1) degrees of freedom. When
those assumptions are not met, it is better to use a permutation approach to
find p-values. Either way, the same statistic is used to test either kind of
hypothesis, independence or homogeneity. If there is evidence to reject the
null hypothesis, then it is interesting to see which cells in the table
contributed to the deviations from the null hypothesis. The standardized
residuals provide that information. Graphing them in a mosaic plot makes for a
fun display to identify the large residuals and often allows you to better
understand the results. This should tie back into the original data display and
contingency table where you identified initial patterns. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020919">
4. 12: Review of Important R commands</a>


</h2>


The main components of R code used inthis chapter follow with components to modify in red where y is a response
variable and x is a predictor are easily identified:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
Tablename`` 
``&lt;-tally(~<span style='color:red'>
x`` +<spanmargin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Thisprovides a table of the counts in the variable called </span>

``Tablename`` . 

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
``margins=T`` 
 is used if want to display row, column, and table totals. 

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Thisfunction requires that the </span>

``mosaic `` package has been loaded. 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``plot(<spanstyle='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
)

</span>


margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Makesa stacked bar chart useful for homogeneity test situations. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``mosaicplot(<spanstyle='color:red'>
Tablename`` )

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Makesa mosaic plot useful for finding patterns in the table in independence test
situations. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``chisq. test(<spanstyle='color:red'>
Tablename`` )

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
ProvidesX<sup>
2</sup>
 and p-values based on the </span>
 **<span style='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:
Calibri;mso-hansi-theme-font:minor-latin;mso-char-type:symbol;mso-symbol-font-family:
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
**  distribution with (R-1)(C-1) degrees of freedom. 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``chisq. test(<spanstyle='color:red'>
Tablename`` )$expected

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesexpected cell counts. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``pchisq(<spanstyle='color:red'>
X-squared`` , df=(<span style='color:red'>
R</span>
-1)*(<spanstyle='color:red'>
C</span>
-1), lower. tail=F)

</span>


margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
c</span>
</span>
**  **<sup>
2</span>
</sup>
**  distribution with (R-1)(C-1) degrees of freedom for observed teststatistic. 

</span>


margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
SeeSection 4. 4 for code related to finding a permutation-based p-value. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``chisq. test(<spanstyle='color:red'>
Tablename`` )$residuals^2

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
ProvidesX<sup>
2</sup>
 contributions from each cell in table. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``chisq. test(<spanstyle='color:red'>
Tablename`` )$residuals

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesstandardized residuals. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``mosaicplot(<spanstyle='color:red'>
Tablename`` , shade=T)

</span>

margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesa mosaic plot with shading based on standardized residuals. 

</span>


margin-bottom:0in;margin-left:1. 0in;margin-bottom:. 0001pt;mso-add-space:auto;<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020920">
4. 13: Practice problems</a>


</h2>


Determine which type of test isappropriate in each situation ?  ***Independence***  or  **4. 1. Concerns over diseases being
transmitted between birds and humans have led to many areas developing
monitoring plans for the birds that are in their regions. The duck pond on
campus at MSU-Bozeman is a bit like a night club for the birds that pass
through Bozeman. 

</span>


i) Suppose that a researcher randomly
samples 20 ducks at the duck pond on campus on 4 different occasions and records
the number ducks that are healthy and number that are sick on each day. The
variables in this study are the day of measurement and sick/healthy. 

</span>


ii) In another monitoring study, a
researcher goes to a wetland area and collects a random sample from all birds
present on a single day, classifies them by type of bird (ducks, swans, etc. )
and then assesses whether each is sick or healthy. The variables in this study
are type of bird and sick/healthy. 

</span>


4. 2. Psychologists performed an
experiment on 48 male bank supervisors attending a management institute to
investigate biases against women in personnel decisions. The supervisors were
asked to make a decision on whether to promote a hypothetical applicant based
on a personnel file. For half of them, the application file described a female candidate;
for the others it described a male. 

</span>


4. 3. Researchers collected data on death
penalty sentencing in Georgia. For 243 crimes, they categorized the crime by
severity from 1 to 6 with Category 1 comprising barroom brawls, liquor-induced
arguments, lovers' quarrels, and similar crimes and Category 6 including the
most vicious, cruel, cold-blooded, unprovoked crimes. They also recorded the
perpetrator's race. They wanted to know if there was a relationship between
race and type of crime. 

</span>


4. 4. Epidemiologists want to see if
Vitamin C helped people with colds. They would like to give some patients
Vitamin C and some a placebo then compare the two groups. However, they are
worried that the placebo might not be working. Since vitamin C has such a
distinct taste, they are worried the participants will know which group they are
in. To test if the placebo was working, they collected 200 subjects and
randomly assigned half to take a placebo and the other half to take Vitamin C. 
30 minutes later, they asked the subjects which supplement they received
(hoping that the patients would not know which group they were assigned to). 

</span>


4. 5. Is zodiac sign related to GPA?
300 randomly selected students from MSU were asked their birthday and their
current GPA. GPA was then categorized as &lt; 1:50 = F, 1. 51-2. 50 = D, 2. 51 -
3. 25 = C, 3. 26-3. 75 = B, 3. 76-4. 0 = A and their birthday was used to find their
zodiac sign. 

</span>


4. 6. In 1935, the statistician R. A. 
Fisher famously had a colleague claim that she could distinguish whether milk
or tea was added to a cup first. Fisher presented her, in a random order, 4
cups that were filled with milk first and 4 cups that were filled with tea first. 

</span>


4. 7. Researchers wanted to see if
people from Rural and Urban areas aged differently. They contacted 200 people
from Rural areas and 200 people from Urban areas and asked the participants
their age (&lt;40, 41-50, 51-60, &gt;60). 

</span>


The *fivethirtyeight blog *often shows up with interesting data summaries
that have general public appeal. Their staff includes a bunch of quants with
various backgrounds. When starting their blog, they had to decide on the data
is/are question that we introduced in Section 1. 0. To help them think about
this, they collected a nationally representative sample that contained three
questions about this. Based on their survey, they concluded that 

</span>


*Relevant to the </span>
*<a
href="http://fivethirtyeight. com/datalab/data-is-vs-data-are/" target="_blank">
*interests of FiveThirtyEight</span>
*</a>
* in particular, we alsoasked whether people preferred using "data" as a singular or plural noun. To
those who prefer the plural, I?ll put this in your terms: The data are pretty
conclusive that the vast majority of respondents think we should say "data is. "
The singular crowd won by a 58 percentage-point margin, with 79 percent of
respondents liking "data is" to 21 percent preferring "data are. " But only half
of respondents had put any thought to the usage prior to our survey, so it
seems that it?s not a pressing issue for most. 

</span>
*

This came from a survey thatcontained questions about *which is thecorrect usage*, (</span>
``isare`` ), *have you thoughtabout this issue* (</span>

``thoughtabout`` ) with levels Yes/No, and *doyou care about this issue *(</span>

``careabout`` ) with four levels from *Notcode loads their data set after missing responses were removed, does a littlere-ordering of factor levels to help make the results easier to understand, and
makes a tableplot to get a general sense of the results including information
on the respondents? gender, age, income, and education (or just look at the
cover of the book). 

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
csd&lt;- read. csv("http://www. math. montana. edu/courses/s217/documents/csd. csv")

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
require(tabplot)

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
csd$careabout&lt;-factor(csd$careabout, levels=levels(csd$careabout)[c(1, 4, 3, 2)])

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
csd$Education&lt;-factor(csd$Education, levels=levels(csd$Education)[c(4, 3, 5, 1, 2)])

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
csd$Household. Income&lt;-factor(csd$Household. Income, levels=levels(csd$Household. Income)[c(1, 4, 5, 6, 2, 3)])

</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
tableplot(csd[, c("isare", "careabout", 
"thoughtabout", "Gender", "Age", "Household. Income", "Education")])

</span>


4. 8. If we are interested in the
variables *isare* and *careabout*, what sort of test should weperform?

</span>


4. 9. Make the appropriate plot of the
results for the table relating those two variables relative to your answer to 4. 8. 

</span>


4. 10. Generate the contingency table
and find the expected cell counts, first "by hand" and then check them using
the output. Is the parametric procedure appropriate here? Why or why not?

</span>


4. 11. Report the value of the test
statistic, its distribution under the null, the parametric p-value, and write a
decision and conclusion, making sure to address scope of inference. 

</span>


4. 12. Make a mosaic plot with the
standardized residuals and discuss the results. Specifically, in what way do
the is/are preferences move away from the null hypothesis for people that care
more about this?

</span>


We might be fighting a losing battle on this particular wordusage, but since we are in the group that cares a lot about this, we are going
to keep trying?</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>
 <spanstyle='mso-bidi-font-size:11. 0pt'>
 </span>

</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>
 A
stratified random sample involves taking a simpler random sample from each
group or strata of the population. It is useful to make sure that each group is
represented at a chosen level (for example the sample proportion of the total
size). If a simple random sample of all schools had been taken, it is possible
that a level could have no schools selected. 



</div>


</div>


** 

</html>

