---
output:
  html_document: default
  pdf_document: default
---

#Two-Way ANOVA

</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:

</span>


```{r}



```


(ToothGrowth$dose)</span>
 **TG2 &lt;-</span>
lm</span>

data=</span>

ToothGrowth)</span>
 **Anova</span>
## AnovaTable (Type II tests)</span>
 **##<span style='mso-spacerun:yes'>
????????????</span>
Sum Sq Df F value<span style='mso-spacerun:yes'>
???</span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## supp<span style='mso-spacerun:yes'>
???????</span>
205. 35<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
15. 572 0. 0002312 ***</span>
 **## dosef<span style='mso-spacerun:yes'>
?????</span>
2426. 43<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
92. 000 &lt; 2. 2e-16 ***</span>
 **## supp:dosef<span style='mso-spacerun:yes'>
?</span>
108. 32<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
?? </span>
4. 107 0. 0218603 *<spanstyle='mso-spacerun:yes'>
? </span>
</span>
 **## Residuals<span style='mso-spacerun:yes'>
??</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
The
ANOVA *F*-test for an interactionbetween supplement type and dosage level is *F*(2, 54)= 4. 107 with a p-value of 0. 022. So there appears to be enough evidence to
reject the null hypothesis of no interaction between *Dosage* and *Delivery method*, supporting a changing effect on tooth growth of dosage based on the delivery
method in the Guinea Pigs that were assigned. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Any similarities between this correct
result and the previous WRONG result are coincidence. I (Greenwood) once
attended a Master?s defense where the results from a similar model were not as
expected (small p-values in places they didn?t expect and large p-values in
places where they thought differences existed). During the presentation, the student
showed some ANOVA tables and the four level categorical variable had 1 numerator
passed with major revisions but had to re-run  **df</i>
 and for the right number of expected model coefficients) to make
sure they match your expectations. This is one reason why you will be learning
to fill in ANOVA tables based on information about the study so that you can be
prepared to detect when your code has let you down<a style='mso-footnote-id:
ftn5' href="#_ftn5" name="_ftnref5" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>
. It
is also a great reason to explore term-plots and coefficient interpretations as
that can also help diagnose errors in model construction. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
Getting back to the previousresults, we now have enough background information to more formally write up a
focused interpretation of these results. The 6+ hypothesis testing steps in
this situation would be focused on first identifying that the best analysis
here is as a Two-Way ANOVA situation (these data were analyzed in Chapter 2 as
a One-Way ANOVA but this version is better because it can explore whether there
is an interaction between delivery method and dosage). We will use a 5%
significance level and start with assessing the evidence for an interaction. If
the interaction had not been dropped, we would have reported the test for the
interaction, re-fit the additive model and used it to explore the main effect
tests and estimates for *Dose* and *Delivery method*. 

</span>

<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
1)
margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
$H_0$</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
No interaction between *delivery method*and *dosage* on odontoblast growth inpopulation of guinea pigs

</span>


<spanstyle='font-size:12. 0pt;font-family:Symbol;mso-ascii-font-family:Calibri;
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

?s=0

</span>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
$H_A$: Interaction between *delivery method* and *dosage* on odontoblast growth in population of guinea pigs </span>

style='font-size:12. 0pt;font-family:Symbol;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 </span>

&#8800; 0

</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
2)
Validity conditions:

</span>
** 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Independence:

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
   12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
This
   assumption is presumed to be met because we don?t know of a reason why
   the independence of the measurements of tooth growth of the guinea pigs
   as studied might be violated. 

</span>
</li>

 </ul>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Constant
   variance:

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
   12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
To
   assess this assumption, we can use the diagnostic plots in Figure 3-9. 

</span>
</li>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
   12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
In
   the Residuals vs Fitted and the Scale-Location plots, the differences in
   variability among the groups (see the different x-axis positions for each
   group?s fitted values) is minor, so there is not strong evidence of a
   problem with the equal variance assumption. 

</span>
</li>

 </ul>

</ul>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
mfrow=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
c</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
2</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
2</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
))</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
plot</span>
</span>
<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(TG2, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
pch=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
16</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
) </span>
</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1034" type="#_x0000_t75"
 style='width:406. 9pt;height:271. 15pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image013. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=543 height=362src="chapter3_files/image014. jpg" v:shapes="_x0000_i1034">
<![endif]>
</span>
<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>


</span>
</i>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-9: Diagnostic plots for the interaction model for Tooth Growth. 

</span>
</i>


``color:blue;background:#E1E2E5'>

&nbsp;
`` 

margin-left:. 25in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l2 level1 lfo12'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Normality of residuals:

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l2 level2 lfo12'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
The QQ-Plot in Figure 3-9 does not
suggest a problem with this assumption. 

</span>


<spanstyle='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
3)
Calculate the test statistic for the Interaction test. 

</span>
** 

```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
 </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
lm</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(len~supp*dosef, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
data=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
ToothGrowth)</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Anova</span>
</span>
<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(TG2)</span>
</span>




<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<spanstyle='mso-spacerun:yes'>
???????????? </span>
Sum Sq Df F value<spanstyle='mso-spacerun:yes'>
??? </span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
???</span>
</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## supp<span style='mso-spacerun:yes'>
???????</span>
205. 35<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
15. 572 0. 0002312 ***</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## dosef<span style='mso-spacerun:yes'>
?????</span>
2426. 43<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
92. 000 &lt; 2. 2e-16 ***</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## supp:dosef<span style='mso-spacerun:yes'>
?</span>
108. 32<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
?? </span>
4. 107 0. 0218603 *<spanstyle='mso-spacerun:yes'>
? </span>
</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Residuals<span style='mso-spacerun:yes'>
??</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
The test statistic is *F*(2, 54)=4. 107. 

</span>
</li>

</ul>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
4) Find
:

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
The ANOVA *F*-test p-value of 0. 0219 for the   interaction. 

</span>
</li>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
To find this p-value
   directly in R, we can use the </span>

   font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";   mso-bidi-font-family:"Times New Roman"'>
pf</span>

   12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

   function:

</span>
</li>

</ul>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
4. 107</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
df1=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
2</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
df2=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
54</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
lower. tail=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
F)</span>
</span>




<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##[1] 0. 0218601</span>
</span>




</div>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
5) Make a
decision:

</span>
** 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Reject
$H_0$ since the p-value (0. 0219) is less than 0. 05. With a p-value
   of 0. 0219, there is about a 2. 19% chance we would observe interaction like
   we did (or more extreme) if none were truly present. This provides strong
   evidence against the null hypothesis of no interaction between delivery
   method and dosage on odontoblast growth so we reject the null hypothesis
   of no interaction. 

</span>
</li>

</ul>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
6) Write
a conclusion:

</span>
** 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Therefore, 
   the effects of dosage level (0. 5, 1, or 2 mg/day) on population average
   tooth odontoblast growth rates of Guinea pigs are changed by the delivery
   (OJ, Vitamin C) method (and vice versa) and we should keep the interaction
   in the model. With the random assignment of levels but not random
   selection of subjects here, we could also write this as: Different dosage
   levels cause different changes in the odontoblast growth based on the
   delivery method for these guinea pigs. <span   style='mso-spacerun:yes'>
?</span>


</span>
</li>
</ul>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
In
a Two-Way ANOVA, we need to go a little further to get to the final
interpretations since the models are more complicated. When there is an
interaction present, we should focus on the interaction plot or term-plot of
the interaction model for an interpretation of the form and pattern of the
interaction. If the interaction were unimportant, then the hypotheses and
results should focus on the additive model results, especially the estimated
model coefficients. To see why we don?t spend much time with the estimated
model coefficients in an interaction model, the model summary for this model is
provided:

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
?????????????? </span>
Estimate Std. Error t valuePr(&gt;|t|)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept)<span style='mso-spacerun:yes'>
????</span>
13. 230<span style='mso-spacerun:yes'>
????? </span>
1. 148<spanstyle='mso-spacerun:yes'>
? </span>
11. 521 3. 60e-16 ***</span>
 **## suppVC<span style='mso-spacerun:yes'>
?????????</span>
-5. 250<span style='mso-spacerun:yes'>
????? </span>
1. 624<spanstyle='mso-spacerun:yes'>
? </span>
-3. 233<span style='mso-spacerun:yes'>
?</span>
0. 00209 ** </span>
 **## dosef1<span style='mso-spacerun:yes'>
??????????</span>
9. 470<span style='mso-spacerun:yes'>
????? </span>
1. 624<spanstyle='mso-spacerun:yes'>
?? </span>
5. 831 3. 18e-07 ***</span>
 **## dosef2<span style='mso-spacerun:yes'>
?????????</span>
12. 830<span style='mso-spacerun:yes'>
????? </span>
1. 624<spanstyle='mso-spacerun:yes'>
?? </span>
7. 900 1. 43e-10 ***</span>
 **## suppVC:dosef1<span style='mso-spacerun:yes'>
??</span>
-0. 680<span style='mso-spacerun:yes'>
????? </span>
2. 297<spanstyle='mso-spacerun:yes'>
? </span>
-0. 296<span style='mso-spacerun:yes'>
?</span>
0. 76831<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## suppVC:dosef2<span style='mso-spacerun:yes'>
???</span>
5. 330<span style='mso-spacerun:yes'>
????? </span>
2. 297<spanstyle='mso-spacerun:yes'>
?? </span>
2. 321<span style='mso-spacerun:yes'>
?</span>
0. 02411 *<span style='mso-spacerun:yes'>
? </span>
</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
There are two </span>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
in the results, related to modifying the estimates for doses of 1 (-0. 68) and 2
(5. 33) for the Vitamin C group. If you want to re-construct the fitted values
from the model that are displayed in the Figure 3-10, you have to look for any
coefficients that are "turned on" for a combination of levels of interest. For
example, for the OJ group (solid line in Figure 3-10), the dosage of 0. 5 mg/day
has an estimate of an average growth of approximately 13 mm. This is the
baseline group, so the model estimate for an observation in the OJ and 0. 5 mg/day
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:108. 75pt;height:16. 5pt'>

 <v:imagedata src="chapter3_files/image015. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=145 height=22src="chapter3_files/image016. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
</span>
microns. 
For the OJ and 2 mg dosage estimate that has a value over 25 microns in the
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
y</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
&#964;</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:202. 9pt;height:16. 5pt'>

 <v:imagedata src="chapter3_files/image017. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=271 height=22src="chapter3_files/image018. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 26. 06</span>
12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
 </span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
microns. 
For the Vitamin C group, another coefficient becomes involved from its "main
effect". For the VC and 0. 5 mg dosage level, the estimate is approximately 8 microns. 
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
y</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
&#947;</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
-5. 25</m:r>
</span>
</m:e>
</m:d>
*<span style='font-size:12. 0pt;font-family: "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:219pt;height:16. 15pt'>

 <v:imagedata src="chapter3_files/image019. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=292 height=22src="chapter3_files/image020. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
7. 98</span>

 microns. 
Finally, when we consider non-baseline results for both groups, three
coefficients are required to reconstruct the results in the plot. For example, 
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
y</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
&#964;</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:   "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
&#947;</m:r>
</span>
</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
 style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family: "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  style='font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:  "Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
-5. 25</m:r>
</span>
</m:e>
</m:d>
*<span style='font-size:12. 0pt;font-family: "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:290. 25pt;height:16. 5pt'>

 <v:imagedata src="chapter3_files/image021. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=387 height=22src="chapter3_files/image022. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
17. 45 mm. We usually will by-pass all thisfun(!) with the coefficients in an interaction model and go from the ANOVA
interaction test to focusing on the pattern of the responses in the interaction
plot, but it is good to know that there are still model coefficients driving
our results. 

</span>


```{r}



```


(TG2), </span>

T, </span>

T, </span>

"bars"</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1033" type="#_x0000_t75" style='width:504. 4pt;height:336. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image023. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=673 height=449src="chapter3_files/image024. jpg" v:shapes="_x0000_i1033">
<![endif]>
</span>
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


*<spanstyle='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure 3-10: Term-plot for
the estimated interaction for the Tooth Growth data. 

</span>
*

<span style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Given the presence of an important
interaction, then the final step in the interpretation here is to interpret the
results in the interaction plot or term-plot of the interaction model, 
supported by the p-value suggesting evidence of a different effect of
supplement type based on the dosage level. To supplement this even more, 
knowing which combinations of levels differ can enhance our discussion. Tukey?s
HSD results (specifically the CLD) can be added to the original interaction
plot by turning on the </span>

``cld=T`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

option in the </span>

``intplot`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

function as seen in Figure 3-11. Sometimes it is hard to see the letters and so
there is also a </span>

``cldshift=?`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

option to move the letters up or down, here a value of 1 seemed to work. 

</span>


```{r}



```


ToothGrowth, </span>

c</span>

(</span>
1</span>
, </span>
2</span>
), </span>
cldshift=</span>
, </span>

T, </span>

"Interaction Plot with CLD"</span>

)</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1032" type="#_x0000_t75" style='width:504. 4pt;height:336. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image025. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=673 height=449src="chapter3_files/image026. jpg" v:shapes="_x0000_i1032">
<![endif]>
</span>
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Courier New"'>
Figure
3-11: Interaction plot with added CLD from Tukey?s HSD. 

</span>
</i>


<i12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;

&nbsp;
</span>
</i>


<!--[if gte vml 1]>
<v:shape id="Ink_x0020_60" o:spid="_x0000_s1026" style='position:absolute;
 margin-left:699. 9pt;margin-top:190. 85pt;width:1. 7pt;height:1. 7pt;z-index:251659264;
 visibility:visible;mso-wrap-style:square;mso-wrap-distance-left:9pt;
 mso-wrap-distance-top:0;mso-wrap-distance-right:9pt;
 mso-wrap-distance-bottom:0;mso-position-horizontal:absolute;
 mso-position-horizontal-relative:text;mso-position-vertical:absolute;
 mso-position-vertical-relative:text' coordsize="0, 0" path="m, l, , , e" filled="f"
 strokecolor="#3165bb" strokeweight=". 58331mm">

 <v:stroke endcap="round"/>

 <v:path shadowok="f" fillok="f" insetpenok="f"/>

 <o:lock v:ext="edit" rotation="t" aspectratio="t" verticies="t" text="t"
 shapetype="t"/>

 <o:ink i="AFodAggIARBYz1SK5pfFT48G+LrS4ZsiAxVIEUSxyu0FRTobAjmlAEY6GwI5pQAFAgtkGRQyCAD+
PwIAAC9DMwgA/j8CFx2bQwoRA4IAggAKABEgIHnz5QwAzwE=
" annotation="t"/>

absolute;z-index:251659264;margin-left:1872px;margin-top:5511px;width:22px;style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Courier New"'>
???????????????? </span>
The interpretation of theprevious hypothesis test result can be concluded with the following discussion. 
Generally increasing the dosage increases the amount of mean growth except for
the 2 mg/day dosage level where the increase levels off in the OJ group (OJ 1
and 2 mg/day are not detectably different) and the differences between the two delivery
methods disappear at the highest dosage level. But for 0. 5 and 1 mg/day
dosages, OJ is clearly better than VC by about 10 microns of growth on average. <spanstyle='mso-no-proof:yes'>
 </span>


</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020901">
<span style='mso-fareast-font-family:"Times New Roman"'>
3. 4:Observational study example: The Psychology of Debt</span>
</a>
<spanstyle='mso-fareast-font-family:"Times New Roman"'>


</span>
</h2>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
In this section, the analysis of aadults will be analyzed from a survey conducted by Lee, Webley, and Walker (1995)and available in the </span>

``debt`` style='font-size:12. 0pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
 </span>

Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
data set from the </span>

style='font-size:12. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
faraway</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

package (Faraway, 2016). The subjects responded to a variety of questions
including whether they buy cigarettes (</span>

font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
cigbuy</span>

12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
: 0 if
no, 1 if yes), their housing situation (house: 1 = rent, 2 = mortgage, and 3 =
owned outright), their income group (</span>

font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
incomegp</span>

12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
: 1 =
lowest, 5 = highest), and their score on a continuous scale of attitudes about
debt (</span>

``prodebt`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
was derived as the average of a series of questions about debt with eachquestion measured on an  ***ordinal***  1 to 5 scale, with highervalues corresponding to more positive responses about <u>
going into debt</u>
 of
various kinds. The ordered scale on surveys that try to elicit your opinions on
topics with scales from 1 to 5 or 1 to 7 or even, sometimes, 1 to 10 is called
a  ***Likertscale***  (Likert, 1932). It is not a quantitative scale and really shouldbe handled more carefully than taking an average of a set responses. That said, 
it is extremely common practice in social science research to treat ordinal
responses as if they are quantitative and take the average of many of them to
create a more continuous response variable like the one we are using here. If
you continue your statistics explorations, you will see some better techniques
for analyzing responses obtained in this fashion. That said, the scale of the
response is relatively easy to understand as an amount of willingness to go
into debt on a scale from 1 to 5 with higher values corresponding to more
willingness to be in debt. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
This data set is typical of survey data
where respondents were not required to answer all questions and there are some
missing responses. We will clean out any individuals that failed to respond to
all questions using the </span>

``na. omit`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

function, which will return only subjects that responded to every question in
the data set. But is this dangerous? Suppose that people did not want to
provide their income levels if they were in the lowest or, maybe, highest
income groups. Then we would be missing responses systematically and conclusions
could be biased because of ignoring these types of subjects. This is another
topic for more advanced statistical methods to try to handle but something
every researcher should worry about when selected subjects do not respond at
all or fail to answer some questions. Is there bias because of responses that
were not observed that could invalidate all my hard won statistical
conclusions? This ties back into our discussion of who was sampled. We need to
think carefully about who was part of the sample but refused to participate and
how that might impact our inferences. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Ignoring this potential for bias in the
results for the moment, we are first interested in whether buying
cigarettes/not and income groups interact in their explanation of the
respondent?s mean opinions on being in debt. The interaction plot (Figure 3-12)
may suggest an interaction between *cigbuy*and *incomegp* from income levels 1 to3 where the lines cross but it is not as clear as the previous examples. The
assess evidence for that interaction. Based on the plot, there do not appear tobe differences based on cigarette purchasing but there might be some
differences between the income groups. If there is no interaction present, then
this suggests that we might be in Scenario 2 or 3 where a single main effect of
interest is present. 

</span>


```{r}



```

 **data</span>
(debt)</span>
 **debt$incomegp&lt;-</span>
(debt$incomegp)</span>
 **debt$cigbuy&lt;-</span>
(debt$cigbuy)</span>
 **debtc&lt;-</span>
(debt)</span>
 **intplot</span>
data=</span>

col=</span>

(</span>

1</span>
, </span>
3</span>
), </span>
lwd=</span>
2</span>
</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1031" type="#_x0000_t75" style='width:444pt;height:295. 9pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image028. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=592 height=395src="chapter3_files/image029. jpg" v:shapes="_x0000_i1031">
<![endif]>
</span>
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


*<spanstyle='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure 3-12: Interaction
plot of Prodebt by income group and buy cigarettes (0=no, 1=yes). 

</span>
*

<span style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
As in other situations, and especially
with observational studies where a single large sample is analyzed, it is
important to check for balance - whether all the combinations of the two
predictor variables are similarly represented. Even more critically, we need to
check whether all the combinations of levels of factors are measured. If a
combination is not measured, then we lose the ability to estimate the mean for
that combination and the ability to test for an interaction. A solution to that
problem would be to collapse the categories of one of the variables, changing
the definitions of the levels but if you fail to obtain information for all
combinations, you can?t work with the interaction model. In this situation, we
barely have enough information to proceed (the smallest </span>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

is 8 for income group 4 that buys cigarettes). We have a very unbalanced design
with counts between 8 and 51 in the different combinations. 

</span>


```{r}



```


debtc)</span>




##<spanstyle='mso-spacerun:yes'>
?????? </span>
incomegp</span>
 **## cigbuy<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
3<span style='mso-spacerun:yes'>
?</span>
4<span style='mso-spacerun:yes'>
? </span>
5</span>
 **##<span style='mso-spacerun:yes'>
????? </span>
0 24 4045 47 51</span>
 **##<span style='mso-spacerun:yes'>
????? </span>
1 23 2918<span style='mso-spacerun:yes'>
? </span>
 **** 19</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
The test for the interaction is always
how we start our modeling in Two-Way ANOVA situations. The ANOVA table suggests
that there is little evidence of interaction between the income level and
buying cigarettes on the opinions of the respondents towards debt (*F*(4, 294)=1. 0003, p-value=0. 408). Thissuggests that the initial assessment that the interaction wasn?t too prominent
was correct. We should move to the additive model here but first need to check
the assumptions to make sure we can trust this initial test. 

</span>


```{r}



```


data=</span>

debtc)</span>
 **Anova</span>
## AnovaTable (Type II tests)</span>
 **## Response: prodebt</span>
 **##<span style='mso-spacerun:yes'>
?????????????????</span>
Sum Sq<span style='mso-spacerun:yes'>
? </span>
Df F value<spanstyle='mso-spacerun:yes'>
?? </span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
??</span>
</span>
 **## incomegp<span style='mso-spacerun:yes'>
?????????</span>
9. 018<span style='mso-spacerun:yes'>
?? </span>
4<spanstyle='mso-spacerun:yes'>
? </span>
4. 5766 0. 001339 **</span>
 **?????????</span>
0. 703<span style='mso-spacerun:yes'>
??</span>
1<span style='mso-spacerun:yes'>
? </span>
1. 4270 0. 233222<spanstyle='mso-spacerun:yes'>
?? </span>
</span>
 **## incomegp:cigbuy<span style='mso-spacerun:yes'>
??</span>
1. 971<span style='mso-spacerun:yes'>
?? </span>
4<spanstyle='mso-spacerun:yes'>
? </span>
1. 0003 0. 407656<spanstyle='mso-spacerun:yes'>
?? </span>
</span>
 **## Residuals<span style='mso-spacerun:yes'>
??????</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
The
diagnostic plots (Figure 3-13) seem to be pretty well-behaved with no apparent
violations of the normality assumption and no clear evidence of a violation of
the constant variance assumption. The observations would seem to be independent
because there is no indication of structure to the measurements of the survey
respondents that might create dependencies. In observational studies, violations
of the independence assumption might come from repeated measures of the same
person or multiple measurements within the same family/household or samples
that are clustered geographically. The random sampling from a population should
allow inferences to a larger population except for that issue of removing
partially missing responses. We also don?t have much information on the
population sampled, so will just leave this vague here but know that there is a
population these conclusions apply to since it was random sample. All of this
suggests proceeding to fitting and exploring the additive model is reasonable
here. No causal inferences are possible because this is an observational study. 

</span>


```{r}



```


c</span>

(</span>
2</span>
, </span>
2</span>
))</span>
 **plot</span>
(debt1)</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1030" type="#_x0000_t75" style='width:418. 15pt;height:278. 65pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image030. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=558 height=372src="chapter3_files/image031. jpg" v:shapes="_x0000_i1030">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure 3-13: Diagnostic
plot for Prodebt by income group and buy cigarettes/not interaction model. 

</span>
*

<span style='font-size:8. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
1)
Hypotheses (Two sets apply when the additive model is the focus!):<spanstyle='mso-spacerun:yes'>
? </span>


</span>
** 
margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
$H_0$</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
No difference in means for prodebt for income groups in population, given
cigarette buying in model </span>

Symbol;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
? </span>
All </span>
<i style='mso-bidi-font-style:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
= 0 in additive model

</span>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
$H_A$</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
Some difference in means for prodebt for income group in population, given
cigarette buying in model </span>

Symbol;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
? </span>
Not all </span>
<i style='mso-bidi-font-style:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
= 0 in additive model

</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
$H_0$</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
No difference in means for prodebt for cigarette buying/not in population, 
given income group in model </span>

Symbol;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
? </span>
All </span>
<i style='mso-bidi-font-style:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
= 0 in additive model

</span>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l5 level1 lfo9'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
$H_A$</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
:
Some difference in means for prodebt for cigarette buying/not in population, 
given income group in model </span>

Symbol;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
? </span>
Not all </span>
<i style='mso-bidi-font-style:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
= 0 in additive model

</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
2)
Validity conditions ? discussed above but with new plots for the additive:

</span>
** 

```{r}



```


data=</span>

debtc)</span>
 **plot</span>
</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1029" type="#_x0000_t75"
 style='width:377. 25pt;height:251. 65pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image032. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=503 height=336src="chapter3_files/image032. png" v:shapes="_x0000_i1029">
<![endif]>
</span>
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-14: Diagnostic plot for Prodebt by income group and buy cigarettes/not


</span>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Constant
   variance:

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
   12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
In
   the Residuals vs Fitted and the Scale-Location plots in Figure 3-14, the
   differences in variability among groups is minor and nothing suggests a
   you should always revisit the diagnostic plots to make sure you didn?t   create problems that were not present in more complicated models. 

** </span>
</li>

 </ul>

</ul>


margin-left:. 25in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l2 level1 lfo12'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Normality of residuals:

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l2 level2 lfo12'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
The QQ-Plot in Figure 3-14 does not
suggest a problem with this assumption. 

</span>


<spanstyle='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
3)
Calculate the test statistic for the two main effect tests. 

</span>
** 

```{r}



```





## AnovaTable (Type II tests)</span>
 **## Response: prodebt</span>
 **##<span style='mso-spacerun:yes'>
???????????</span>
Sum Sq<span style='mso-spacerun:yes'>
? </span>
Df F value<spanstyle='mso-spacerun:yes'>
?? </span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
??</span>
</span>
 **## incomegp<span style='mso-spacerun:yes'>
???</span>
9. 018<span style='mso-spacerun:yes'>
?? </span>
4<spanstyle='mso-spacerun:yes'>
? </span>
4. 5766 0. 001335 **</span>
 **## cigbuy<span style='mso-spacerun:yes'>
?????</span>
0. 703<span style='mso-spacerun:yes'>
?? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1. 4270 0. 233210<spanstyle='mso-spacerun:yes'>
?? </span>
</span>
 **## Residuals 146. 806 298<spanstyle='mso-spacerun:yes'>
??????????????????? </span>
</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
The test statistics are *F*(4, 298)=4. 577 and *F*(1, 298)=1. 427. 

</span>
</li>

</ul>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
4) Find
:

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
The ANOVA *F*-test p-values are 0. 001335 for   the income group variable (conditional on cigarette buy) and 0. 2232 for
   the cigarette buy variable (conditional on income group). 

</span>
</li>

</ul>


 **<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";

&nbsp;
</span>
** 

 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
5) Make
decisions:

</span>
** 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Reject
$H_0$ of no income group differences (p-value=0. 0013) and fail to
$H_0$ of no cigarette buying differences (p-value=0. 2232), 
   each after controlling for the other variable. 

</span>
</li>

</ul>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
6) Write
a conclusion:

</span>
** 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
There
   was initially no evidence to support retaining the interaction of income
   group and cigarette buying on pro-debt feelings (*F*(4, 294)=1. 00, p-value =0. 408) so the interaction was dropped   from the model. There is strong evidence of some difference in the mean
   pro-debt feelings in the population across the income groups, after
   adjusting for cigarette buying. There is little to no evidence of a
   difference in the mean pro-debt feelings in the population based on
   cigarette buying/not, after adjusting for income group. 

</span>
</li>

</ul>


<spanstyle='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
So we learned that the additive model
was more appropriate for these responses and that the results resemble Scenario
2 or 3 with only one main effect being important. In the additive model, the
coefficients can be interpreted as shifts from the baseline after controlling
for the other variable in the model. Figure 3-15 shows the increasing average
comfort with being in debt as the income groups go up. Being a cigarette buyer
was related to a lower comfort level with debt. But compare the y-axis scales
in the two plots - the differences in the means across income groups are almost
0. 5 points on a 5 point scale whereas the difference across </span>

style='font-size:12. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
cigbuy</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?s
two levels is less than 0. 15 units. The error bars for the 95% confidence
intervals are of similar width but the differences in means show up clearly in
the income group term-plot. This is all indirectly related to the size of the
p-values for each term in the additive model but hopefully helps to build some
intuition on the reason for differences. 

</span>


```{r}



```


(debt1r))</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
The estimated coefficients can also be
interesting to interpret for the additive model. Here is the model summary:

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
???????????? </span>
Estimate Std. Error t valuePr(&gt;|t|)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept)<span style='mso-spacerun:yes'>
?</span>
3. 05484<span style='mso-spacerun:yes'>
??? </span>
0. 11127<spanstyle='mso-spacerun:yes'>
? </span>
27. 454<span style='mso-spacerun:yes'>
?</span>
&lt; 2e-16 ***</span>
 **## incomegp2<span style='mso-spacerun:yes'>
???</span>
0. 01641<span style='mso-spacerun:yes'>
??? </span>
0. 13289<spanstyle='mso-spacerun:yes'>
?? </span>
0. 123 0. 901826<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## incomegp3<span style='mso-spacerun:yes'>
???</span>
0. 17477<span style='mso-spacerun:yes'>
??? </span>
0. 13649<spanstyle='mso-spacerun:yes'>
?? </span>
1. 280 0. 201385<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## incomegp4<span style='mso-spacerun:yes'>
???</span>
0. 16901<span style='mso-spacerun:yes'>
??? </span>
0. 14275<spanstyle='mso-spacerun:yes'>
?? </span>
1. 184 0. 237381<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## incomegp5<span style='mso-spacerun:yes'>
???</span>
0. 46833<span style='mso-spacerun:yes'>
??? </span>
0. 13378<spanstyle='mso-spacerun:yes'>
?? </span>
3. 501 0. 000535 ***</span>
 **## cigbuy1<span style='mso-spacerun:yes'>
????</span>
-0. 10640<span style='mso-spacerun:yes'>
??? </span>
0. 08907<spanstyle='mso-spacerun:yes'>
? </span>
-1. 195 0. 233210<spanstyle='mso-spacerun:yes'>
??? </span>
</span>



</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1028" type="#_x0000_t75" style='width:426. 4pt;height:284. 25pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image033. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=569 height=379src="chapter3_files/image034. jpg" v:shapes="_x0000_i1028">
<![endif]>
</span>
style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-15: Term-plots for the Prodebt additive model with left panel for income
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
In the model, the baseline groupis for non-cigarette buyers (</span>

"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
cigbuy</span>

Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:  "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:9pt;height:17. 65pt'>

 <v:imagedata src="chapter3_files/image035. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=12 height=24src="chapter3_files/image036. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=
3. 055 points. Regardless of the </span>

font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
cigbuy</span>

12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
 level, 
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:12. 75pt;height:17. 65pt'>

 <v:imagedata src="chapter3_files/image037. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=17 height=24src="chapter3_files/image038. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=0. 016, 
an increase in the mean score of 0. 016 points. Similarly, the difference
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:12. 75pt;height:17. 65pt'>

 <v:imagedata src="chapter3_files/image039. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=17 height=24src="chapter3_files/image040. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=0. 175
points, regardless of cigarette smoking status. The estimated difference
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
  mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 15pt;height:17. 65pt'>

 <v:imagedata src="chapter3_files/image041. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=18 height=24src="chapter3_files/image042. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=
-0. 106 points for any income group, remember that this variable had a moderately
large p-value in this model. The additive model-based estimates for all six combinations
can be found in Table 3-2. 

</span>


<span style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;

&nbsp;
</span>


*<span style='font-size:12. 0pt;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;
Table 3-2: Calculations to construct
the estimates for all combinations of variables for the Prodebt additive model. 


</span>
*

<table class=MsoNormalTable border=1 cellspacing=0 cellpadding=0
 style='border-collapse:collapse;mso-table-layout-alt:fixed;border:none;
 . 5pt solid windowtext'>

 <tr style='mso-yfti-irow:0;mso-yfti-firstrow:yes;height:15. 0pt'>

</td>

 <td width=633 nowrap colspan=5 valign=bottom style='width:475. 1pt;border:

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:#0070C0'>
Income Group

</span>
** 
 </td>

 </tr>

 <tr style='mso-yfti-irow:1;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='mso-ascii-font-family:Calibri;mso-fareast-font-family: "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:red'>
Cig buy

</span>

 </td>

 <td width=82 nowrap valign=bottom style='width:61. 35pt;border-top:none;

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
 **<span style='font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 style='font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family: "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>
** 
 </td>

 <td width=150 nowrap valign=bottom style='width:112. 5pt;border-top:none;

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:#0070C0'>
2

</span>
** 
 </td>

 <td width=144 nowrap valign=bottom style='width:1. 5in;border-top:none;

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:#0070C0'>
3

</span>
** 
 </td>

 <td width=138 nowrap valign=bottom style='width:103. 5pt;border-top:none;

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:#0070C0'>
4

</span>
** 
 </td>

 <td width=120 nowrap valign=bottom style='width:89. 75pt;border-top:none;

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:#0070C0'>
5

</span>
** 
 </td>

 </tr>

 <tr style='mso-yfti-irow:2;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='mso-ascii-font-family:Calibri;mso-fareast-font-family: "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:red'>
0: No

</span>

 </td>

 <td width=82 nowrap valign=bottom style='width:61. 35pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:9pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image035. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=12 height=24 src="chapter3_files/image036. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
 mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman"'>
?</span>
</span>
 mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman"; mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
= 3. 055<span style='color:black'>


</span>
</span>
 </td>

 <td width=150 nowrap valign=bottom style='width:112. 5pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<m:acc>
<m:accPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:38. 25pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image043. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=51 height=24 src="chapter3_files/image044. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
 mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman"'>


</span>

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055+0. 016 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=30. 71</span>

 style='font-size:12. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family: "Times New Roman"'>


</span>

 </td>

 <td width=144 nowrap valign=bottom style='width:1. 5in;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:38. 25pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image045. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=51 height=24 src="chapter3_files/image046. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>


</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=
 3. 055+ 0. 175

</span>

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
=3. 230</span>
<span style='font-size:12. 0pt; font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>


</span>
 </td>

 <td width=138 nowrap valign=bottom style='width:103. 5pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:38. 25pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image047. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=51 height=24 src="chapter3_files/image048. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>


</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=
 3. 055+ 0. 169

</span>

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
=3. 224</span>
<span style='font-size:12. 0pt; font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>


</span>
 </td>

 <td width=120 nowrap valign=bottom style='width:89. 75pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:38. 25pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image049. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=51 height=24 src="chapter3_files/image050. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>


</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=
 3. 055+ 0. 468

</span>

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
=3. 523</span>
<span style='font-size:12. 0pt; font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>


</span>
 </td>

 </tr>

 <tr style='mso-yfti-irow:3;mso-yfti-lastrow:yes;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='mso-ascii-font-family:Calibri;mso-fareast-font-family: "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:red'>
1: Yes

</span>

 </td>

 <td width=82 nowrap valign=bottom style='width:61. 35pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:9pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image035. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=12 height=24 src="chapter3_files/image036. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;
 mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman"'>
?</span>
+</span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
 </m:r>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:16. 5pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image051. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=22 height=24 src="chapter3_files/image052. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055 ? 0. 106 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=2. 949<span style='color:black'>


</span>
</span>
 </td>

 <td width=150 nowrap valign=bottom style='width:112. 5pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
 </m:r>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:77. 65pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image053. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=104 height=24 src="chapter3_files/image054. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055+ 0. 016 ? 0. 106 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=2. 965</span>

 style='font-size:12. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family: "Times New Roman"'>


</span>

 </td>

 <td width=144 nowrap valign=bottom style='width:1. 5in;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
 </m:r>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:77. 65pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image055. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=104 height=24 src="chapter3_files/image056. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055+ 0. 175 ? 0. 106 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
= 3. 124</span>

 style='font-size:12. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family: "Times New Roman"'>


</span>

 </td>

 <td width=138 nowrap valign=bottom style='width:103. 5pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
 </m:r>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:77. 65pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image057. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=104 height=24 src="chapter3_files/image058. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055+ 0. 169 ? 0. 106 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
= 3. 118</span>

 style='font-size:12. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family: "Times New Roman"'>


</span>

 </td>

 <td width=120 nowrap valign=bottom style='width:89. 75pt;border-top:none;

 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   mso-bidi-font-family:"Times New Roman"'>
<m:r>
$\alpha$ </m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
&#964;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
   mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:"Cambria Math";
    mso-bidi-font-family:"Times New Roman";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
    mso-bidi-font-family:"Times New Roman"'>
<m:r>
 </m:r>
<m:r>
&#947;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
<m:r>
<m:rPr>
<m:scr
 style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif; mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;position:relative;top:4. 5pt;mso-text-raise:
 -4. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
 AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
  style='width:77. 65pt;height:17. 65pt'>

  <v:imagedata src="chapter3_files/image059. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=104 height=24 src="chapter3_files/image060. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
?</span>
= 3. 055+ 0. 468 ? 0. 106 

</span>
 <p class=MsoNormal align=center style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:center;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
=3. 417</span>

 style='font-size:12. 0pt;font-family:"Times New Roman", serif;mso-fareast-font-family: "Times New Roman"'>


</span>

 </td>

 </tr>

</table>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
One final plot of the fitted values
from this additive model in Figure 3-16 hopefully crystalizes the implications
of an additive model and reinforces that this model creates and assumes that
the differences across levels of one variable are the same regardless of the
level of the other variable and that this creates parallel lines. The
difference between </span>

``cigbuy`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

levels across all income groups is a drop in -0. 106 points. The income groups
have the same differences regardless of cigarette buying or not, with income
group 5 much higher than the other four groups. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1027" type="#_x0000_t75" style='width:504. 4pt;height:336. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image061. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=673 height=449src="chapter3_files/image062. jpg" v:shapes="_x0000_i1027">
<![endif]>
</span>
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-16: Illustration of the results from Table 3-2 showing the combined impacts
of the components of the additive model for Prodebt. Panel (a) uses income
groups on the x-axis and different lines for cigarette buyers (1) or not (0). 
Panel (b) displays the different income groups as lines with the cigarette
buying status on the x-axis. 

</span>
*

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>

&nbsp;
`` 

 **<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
In
general, we proceed through the following steps in any 2-WAY ANOVA situation:

</span>
** 

<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
1)
Make an interaction plot. 

</span>


<span style='font-size:9. 0pt;mso-bidi-font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
2)
Fit the interaction model; examine the test for the interaction. 

</span>


<span style='font-size:9. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
3)
Check the residual diagnostic plots for the interaction model (especially
normality and equal variance). 

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l9 level1 lfo5'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
If there is a problem with normality or
equal variance, consider a "transformation" of the response as discussed in
Chapter 6. This can help make the responses have similar variances or responses
to be more normal, but sometimes not both. 

</span>


<span style='font-size:9. 0pt;mso-bidi-font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
4)
If the interaction test has a small p-value, that is your main result. Focus on
the interaction plot from (1) to fully understand the results, adding Tukey?s
HSD results to see which means of the combinations of levels are detected as
being different. 

</span>


<span style='font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
5)
If the interaction is not considered important, then re-fit the model without
the interaction (additive model) and re-check the diagnostic plots. If the
diagnostics are reasonable to proceed:

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l9 level1 lfo5'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Focus on the results for each explanatory
variable, using Type II tests especially if the design is not balanced. 

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l9 level1 lfo5'>
<![if !supportLists]>
<span style='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Report the initial interaction test
results and the results for the test for each variable from the model that is re-fit
without the interaction. 

</span>


margin-left:. 75in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-list:l9 level1 lfo5'>
<![if !supportLists]>
<span style='font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Model coefficients are interesting as
they are shifts from baseline for each level of each variable, controlling for
the other variable ? interpret those differences if the number of levels is not
too great. </span>

<span style='font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>


</span>

<spanstyle='font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Whether
you end up favoring an additive or interaction model, all steps of the
hypothesis testing protocol should be engaged. 

</span>


 **<span style='font-size:8. 0pt;mso-bidi-font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>
** 

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020902">
<span style='mso-fareast-font-family:"Times New Roman"'>
3. 5:Pushing Two-Way ANOVA to the limit: Un-replicated designs</span>
</a>
<spanstyle='mso-fareast-font-family:"Times New Roman"'>


</span>
</h2>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
In some situations, it is too expensive
to replicate combinations of treatments and only one observation at each
combination of the two explanatory variables, A and B, is possible. In these
situations, even though we have information about all combinations of A and B, 
it is no longer possible to test for an interaction. Our regular rules for
degrees of freedom show that we have nothing left for the error degrees of
freedom and so we have to drop the interaction and call that potential
interaction variability "error". 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
We can still perform an analysis of the
responses but an issue occurs with trying to estimate the interaction *F*-test statistic ? we run out of degreesof freedom for the error. To illustrate these methods, the paper towel example
is revisited except that only one response for each combination is used. Now
the entire data set can be displayed:

</span>


```{r}



```


"http://www. math. montana. edu/courses/s217/documents/ptR. csv"</span>

)</span>
 **ptR$dropsf&lt;-</span>
factor</span>
(ptR$drops)</span>
 **##<spanstyle='mso-spacerun:yes'>
?? </span>
brand drops responses dropsf</span>
 **## 1<span style='mso-spacerun:yes'>
??? </span>
B1<spanstyle='mso-spacerun:yes'>
??? </span>
10 1. 9064356<spanstyle='mso-spacerun:yes'>
???? </span>
10</span>
 **## 2<span style='mso-spacerun:yes'>
??? </span>
B2<spanstyle='mso-spacerun:yes'>
??? </span>
10 3. 0504173<spanstyle='mso-spacerun:yes'>
???? </span>
10</span>
 **## 3<span style='mso-spacerun:yes'>
??? </span>
B1<spanstyle='mso-spacerun:yes'>
??? </span>
20 0. 7737965<spanstyle='mso-spacerun:yes'>
???? </span>
20</span>
 **## 4<span style='mso-spacerun:yes'>
??? </span>
B2<spanstyle='mso-spacerun:yes'>
??? </span>
20 2. 8384124<spanstyle='mso-spacerun:yes'>
???? </span>
20</span>
 **## 5<span style='mso-spacerun:yes'>
??? </span>
B1<spanstyle='mso-spacerun:yes'>
??? </span>
30 1. 5557071<spanstyle='mso-spacerun:yes'>
???? </span>
30</span>
 **## 6<span style='mso-spacerun:yes'>
??? </span>
B2<spanstyle='mso-spacerun:yes'>
??? </span>
30 0. 5470565<spanstyle='mso-spacerun:yes'>
???? </span>
30</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Upon
first inspection the interaction plot in Figure 3-17 looks like there might be
some interesting interactions present. But remember now that there is only a
single observation at each combination of the brands and water levels so there
is not much power to detect differences in this sort of situation and no replicates
at combination that allow estimation of SEs so no bands are produced in the
plot. 

</span>


```{r}



```


ptR, </span>

2</span>

</div>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
<!--[if gte vml 1]>
<v:shape
 id="Picture_x0020_501" o:spid="_x0000_i1026" type="#_x0000_t75" style='width:347. 25pt;
 height:180pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image063. emz" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=463 height=240src="chapter3_files/image064. png" v:shapes="Picture_x0020_501">
<![endif]>
</span>
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-17: Interaction plot in paper towel data set with no replication. 

</span>
</i>


<span style='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
The next step would be to assess the
statistical evidence related to an interaction between *Brand* and *Drops*. Aproblem will arise in trying to form the ANOVA table as you would see this in
the console:

</span>


``"Times New Roman";mso-bidi-font-family:"Courier New";color:blue;background:
#E1E2E5'>
anova(lm(responses~dropsf*brand, data=ptR))

`` 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
Analysis of Variance Table

`` 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
Response: responses

`` 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
???????????? `` Df<spanstyle='mso-spacerun:yes'>
? </span>
Sum Sq Mean Sq F value Pr(&gt;F)

</span>

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
dropsf<span style='mso-spacerun:yes'>
??????? `` 2 2. 038721. 01936<span style='mso-spacerun:yes'>
?????????????? </span>


</span>

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
brand<span style='mso-spacerun:yes'>
???????? `` 1 0. 806630. 80663<span style='mso-spacerun:yes'>
?????????????? </span>


</span>

``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>
dropsf:brand<span style='mso-spacerun:yes'>
? `` 2 2. 487731. 24386<span style='mso-spacerun:yes'>
?????????????? </span>


</span>

 **Residuals<spanstyle='mso-spacerun:yes'>
???? </span>
0 0. 00000<spanstyle='mso-spacerun:yes'>
?????????????????????? </span>


</span>
** 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:#C5060B;background:
#E1E2E5'>
Warning message:

`` 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:#C5060B;background:
#E1E2E5'>
In anova. lm(lm(responses~dropsf*brand, data=ptR)):

`` 

``"Times New Roman";mso-bidi-font-family:"Courier New";color:#C5060B;background:
#E1E2E5'>
? `` ANOVA F-tests on anessentially perfect fit are unreliable</span>
font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>


</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Warning
messages in R output show up in red after you run functions that contain
problems and are generally not a good thing, but can sometimes be ignored. In
this case, the warning message is not needed ? there are no *F*-statistics or p-values in the resultsso we know there are some issues with the results. The bolded line is key here
squares of 0. Without replication, there are no degrees of freedom left toestimate the residual error. My (Greenwood?s) first statistics professor, 
Gordon Bril at Luther College, used to refer to this as "shooting your load" by
fitting too many terms in the model given the number of observations available. 
Maybe this is a bit graphic but hopefully will help you remember the need for
replication if you want to estimate interactions ? it did for me. Without
replication of observations, we run out of information to estimate and test all
the desired model components. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
So what can we do if we can?tthat the interaction does not exist and use those degrees of freedom andvariability as the error variability. When we drop the interaction from Two-Way
models, the interaction variability is added into the SS<sub>
E</sub>
 so this is
interaction between the variables. We are not able to test for an interactionso must rely on the interaction plot to assess whether an interaction might be
present. Figure 3-17 suggests there might be an interaction in these data (the
two brands lines cross noticeably suggesting non-parallel lines). So in this
case, assuming no interaction is present is hard to justify. But if we proceed
under this dangerous assumption, tests for the main effects can be developed. 

</span>


```{r}



```


data=</span>

ptR)</span>
 **Anova</span>
## AnovaTable (Type II tests)</span>
 **## Response: responses</span>
 **##<span style='mso-spacerun:yes'>
???????????</span>
Sum Sq Df F value Pr(&gt;F)</span>
 **## dropsf<span style='mso-spacerun:yes'>
???</span>
2. 03872<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
0. 8195 0. 5496</span>
 **## brand<span style='mso-spacerun:yes'>
????</span>
0. 80663<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
0. 6485 0. 5052</span>
 **## Residuals 2. 48773<span style='mso-spacerun:yes'>
?</span>
2</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
In
the additive model, the last row of the ANOVA that is called the </span>

style='font-size:12. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
Residuals</span>

style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

row is really the interaction row from the interaction model ANOVA table. 
Neither main effect had a small p-value (*Drops*:*F*(2, 2)=0. 82, p-value=0. 55 and *Brand*: *F*(1, 2)=0. 65, p-value=0. 51) in the additive model. To get smallp-values with small sample sizes, the differences would need to be  ****  large because the residual degreesof freedom have become very small. The term-plots in Figure 3-18 show that the
differences among the levels are small relative to the residual variability as
seen in the error bars around each point estimate. 

</span>


```{r}



```


(norep1))</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75"
 style='width:424. 15pt;height:282. 75pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter3_files/image065. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=566 height=377src="chapter3_files/image066. jpg" v:shapes="Picture">
<![endif]>
</span>
style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-fareast-font-family:"Times New Roman";mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman"'>
Figure
3-18: Term-plots for the additive model in paper towel data set with no
replication. 

</span>
</i>


<span style='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Hopefully by pushing the limits there
are two conclusions available from this section. First, replication is
important, both in being able to perform tests for interactions and for having
enough power to detect differences for the main effects. Second, dropping from
the interaction model to additive model, the variability explained by the
interaction term is pushed into the error term, whether replication is
available or not. 

</span>


<span style='font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020903">
<span style='mso-fareast-font-family:"Times New Roman"'>
3. 6:Chapter summary</span>
</a>
<span style='mso-fareast-font-family:"Times New Roman"'>


</span>
</h2>

<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
In this chapter, methods forhandling two different categorical predictors in the same model with a
continuous response were developed. The methods build on techniques from
Chapter 2 for the One-Way ANOVA and there are connections between the two
models. This was most clearly seen in the guinea pig data set that was analyzed
in both chapters. When two factors are available, it is better to start with
the methods developed in this chapter because the interaction between the
factors can, potentially, be separated from their main effects. The additive
model is easier to interpret but should only be used when no evidence of an
interaction is present. When an interaction is determined to be present, the
main effects should not be interpreted and the interaction plot in combination
with Tukey?s HSD provides information on the important aspects of the results. 

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
If the interaction is retained
   in the modele, there are two things you want to do with interpreting the
   interaction:

</span>
</li>

</ul>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-ascii-font-family:Calibri;mso-fareast-font-family:Calibri;mso-hansi-font-family:Calibri;mso-bidi-font-family:Calibri'>
1. <spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Describe
the interaction, going through the changes from left to right in the
interaction plot or term-plot for each level of the other variable. 

</span>


<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


margin-left:1. 0in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-ascii-font-family:Calibri;mso-fareast-font-family:Calibri;mso-hansi-font-family:Calibri;mso-bidi-font-family:Calibri'>
2. <spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
Suggest
optimal combinations of the two variables to either get the highest or lowest
possible responses. 

</span>


margin-left:1. 5in;margin-bottom:. 0001pt;text-indent:-. 25in;line-height:normal;mso-ascii-font-family:Calibri;mso-fareast-font-family:Calibri;mso-hansi-font-family:Calibri;mso-bidi-font-family:Calibri'>
a. <spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
For
example, you might want to identify a dosage and delivery method for the guinea
pigs to recommend and one to avoid if you want to optimize odontoblast growth. 

</span>


<spanstyle='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height:
   Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:   Calibri;mso-bidi-font-family:"Times New Roman"'>
If there is no
   interaction, then the additive model provides information on each of the
   variables and the differences across levels of each variable are the same
   regardless of the levels of the other variable. 

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";   mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
You
   can describe the deviations from baseline as in Chapter 2, but for each
   variable, noting that you are controlling for the variable. 

</span>
</li>

 </ul>

</ul>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
Some statisticians might have different
recommendations for dealing with interactions and main effects, especially in
the context of evidence of an interaction. We have chosen to focus on tests for
interactions to screen for "real" interactions and then interpret the
interaction plots aided by the Tukey?s HSD for determining which combinations
of levels are detectably different. Others might suggest exploring the main
effects tests even with interactions present. In some cases, those results are
interesting but in others the results can be misleading and wanted to avoid
trying to tell you when that might happen. Consider two scenarios, one where
the main effects have large p-values but the interaction has a small p-value
and the other where the main effects and the interaction all have small
p-values. The methods discussed in this chapter allow us to effectively arrive
at the interpretation of the differences in the results across the combinations
of the treatments due to the interaction having a small p-value. The main
effects results are secondary results at best when the interaction is important
because we know that impacts of one explanatory variable is changing based on
the levels of the other variable. 

</span>


<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>
????????????? </span>
Chapter 4 presents a bit of adifferent set of statistical methods that allow analyses of data sets similar
to those considered in the last two chapters but with a categorical response
variable. The methods are very different but are quite similar in overall goals
to those in Chapter 2 where differences in responses where explored across
groups. After Chapter 4, the rest of the semester will return to fitting models
using the </span>

``lm`` style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

function as used here, but incorporating quantitative predictor variables and
then eventually incorporating both categorical and quantitative predictor
variables. The methods in Chapter 7 are actually quite similar to those considered
here. 

</span>


<span style='font-size:10. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020904">
<span style='mso-fareast-font-family:"Times New Roman"'>
3. 7:Important R code</span>
</a>
<span style='mso-fareast-font-family:"Times New Roman"'>


</span>
</h2>

The main components of R code used inthis chapter follow with components to modify in red, remembering that any R
packages mentioned need to be installed and loaded for this code to have a
chance of working:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``tally(<spanstyle='color:red'>
A`` ~<span style='color:red'>
B</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
)

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe </span>

 package be loaded. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesthe counts of observations in each combination of categorical predictor
variables A and B, used to check for balance and understand sample sizes in
each combination. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
DATASETNAME`` 
``$<span style='color:red'>
VARIABLENAME`` &lt;-factor(<spanmargin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Usethe </span>

 function on any numerically coded
explanatory variable where the numerical codes represent levels of a
categorical variable. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``intplot(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
A</span>
*<spanmargin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='mso-bidi-font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Download and install using: </span>

style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:115%;font-family:"Lucida Console"'>
source("http://www. math. montana. edu/courses/s217/documents/intplot. R")</span>



</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesinteraction plot. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
INTERACTIONMODELNAME `` 
``&lt;-lm(<span style='color:red'>
Y`` ~<spanstyle='color:red'>
A</span>
*<span style='color:red'>
B</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
)

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Fitsthe interaction model with main effects for A and B and an interaction between
them. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Thisis the first model that should be fit in Two-Way ANOVA modeling situations. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
ADDITIVEMODELNAME`` 
``&lt;-lm(<span style='color:red'>
Y`` ~<spanstyle='color:red'>
A</span>
+<span style='color:red'>
B</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
)

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Fitsthe additive model with only main effects for A and B but no interaction
between them. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Shouldonly be used if the interaction has been decided to be unimportant using a test
for the interaction. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``summary`` 
(</span>
``MODELNAME)`` 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesmodel summary information including the estimated model coefficients, SEs, 
t-tests, and p-values. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``Anova`` 
(</span>
``MODELNAME)`` 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe </span>

 package to be loaded. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesa Type II Sums of Squares the ANOVA table that is useful for both additive and
interaction models, but it most important to use when working with the additive
model as it provides inferences for each term conditional on the other one. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``par(mfrow=c(2, 2));
plot(MODELNAME)

`` 

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesfour diagnostic plots including the Residuals vs Fitted and Normal Q-Q. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``plot(allEffects(<spanstyle='color:red'>
MODELNAME`` ))</span>


</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Plotsthe results from the estimated model. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe </span>

 package be loaded. 

</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020905">
<span style='mso-fareast-font-family:"Times New Roman"'>
3. 8:Practice problems</span>
</a>
<span style='mso-fareast-font-family:"Times New Roman"'>


</span>
</h2>

<span style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
To practice the Two-Way ANOVA, consider
Mathematics Usage Test scores from 1987. The test was given to a sample of highschool seniors who met one of three profiles of high school mathematics course
work: (a) Algebra I only; (b) two Algebra courses and Geometry; and (c) two
Algebra courses, Geometry, Trigonometry, Advanced Mathematics, and Beginning
Calculus. These data were generated from summary statistics for one particular
form of the test as reported by Doolittle (1989). The source of this version of
the data set is Ramsey and Schafer (2002) and the </span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
Sleuth2</span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 package (Ramsey and Schafer, 2012). 
First install and then load that package. 

</span>


``"Times New Roman"'>
require(Sleuth2)

`` 

``"Times New Roman"'>
require(mosaic)

`` 

``"Times New Roman"'>
math &lt;- ex1320

`` 

``"Times New Roman"'>
names(math)

`` 

``"Times New Roman"'>
favstats(Score~Sex+Background, data = math)

`` 

``"Times New Roman"'>

&nbsp;
`` 

<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 1. Use the </span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
favstats</span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
 summary to discuss whether the design
was balanced or not. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 2. Make a side-by-side beanplot and
interaction plot of the results and discuss the relationship between Sex, 
Background, and ACT Score. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 3. Write out the interaction model in
terms of the Greek letters, making sure to define all the terms and don?t
forget the error terms in the model. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 4. Fit the interaction plot and find
the ANOVA table. For the test you should consider first (the interaction), 
write out the hypotheses, report the test statistic, p-value, distribution of
the test statistic under the null, and write a conclusion related to the
results of this test. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 5. Re-fit the model as an additive
model (why is this reasonable here?) and use Anova to find the Type II sums of
squares ANOVA. Write out the hypothesis for the Background variable, report the
test statistic, p-value, distribution of the test statistic under the null, and
write a conclusion related to the results of this test. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 6. Use the </span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
effects</span>

style='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Times New Roman"'>
 package to make a term-plot from the
additive model from 3. 5 and discuss the results. Specifically, discuss what you
can conclude about the average relationship across both sexes, between
Background and average ACT score?

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 7. Make our standard diagnostic plots
and assess the assumptions using these plots. Can you assess independence using
these plots? Discuss this assumption in this situation. 

</span>


<spanstyle='font-size:12. 0pt;mso-bidi-font-size:8. 0pt;mso-fareast-font-family:"Times New Roman";
3. 8. Use the estimated model
coefficients to determine which of the combinations of levels provides the
highest estimated average score. 

</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>




</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>




</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>

Just so you don?t think that perfect R code should occur on the first try, we
have all made similarly serious coding mistakes even after accumulating more
than decade of experience with R. It is finding those mistakes that matters. 



</div>


</div>


** 

</html>

