The previous results used the </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
lm</span>

 function as a "black box" to generate theestimated coefficients. The lines produced probably look reasonable but you
could imagine drawing other lines that might look equally plausible. Because we
are interested in explaining variation in the response variable, we want a
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
e</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:72. 75pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image052. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=97 height=24src="chapter5_files/image053. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
) to find a
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:40. 5pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image054. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=54 height=24src="chapter5_files/image055. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
as small as possible. We can?t just add theseup because it would always be 0 (remember why we use the variance to measure
spread from introductory statistics?). We use a similar technique in
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
e</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
(</m:r>
<m:r>
y</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
    font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:93. 75pt;height:18pt'>

 <v:imagedata src="chapter5_files/image056. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=125 height=24src="chapter5_files/image057. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
over all the observations, minimizing the  ***Sum of Squared Residuals*** =</span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:nary>
<m:naryPr>
<m:chr
  m:val="&#8721;"/>
<m:limLoc m:val="undOvr"/>
<m:subHide m:val="on"/>
<m:supHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
    font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
e</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:29. 65pt;height:18pt'>

 <v:imagedata src="chapter5_files/image058. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=40 height=24src="chapter5_files/image059. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. </span>

 Findingthe estimated regression coefficients that minimize the sum of squared
residuals is called the  ***least squares estimation*** and provides us a reasonable method for finding the "best" estimatedregression line of all the possible choices. 

</span>


Forthe *Beers* vs *BAC* data, Figure 5-16 shows the result of a search for the optimalslope coefficient between values of 0 and 0. 03. The plot shows how the sum of
the squared residuals was minimized for the value that </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
lm</span>

 returned at 0. 018. The main point of this isthat if any other slope coefficient was tried, it did not do as good  ****  as the least squares estimates. 

</span>

<!--[if gte vml 1]>
<v:shape id="_x0000_i1034" type="#_x0000_t75" style='width:267. 4pt;height:167. 25pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image060. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=357 height=223src="chapter5_files/image061. jpg" v:shapes="_x0000_i1034">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-16: Plot of sum of
squared residuals vs possible slope coefficients for Beers vs BAC data, with
vertical line for the least squares estimate that minimizes the sum of squared
residuals. 

</span>
*

*<spanstyle='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>

&nbsp;
</span>
*

????????????? </span>
Sometimes it is helpful to have a
go at finding the estimates yourself. If you install and load the </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
tigerstats</span>

 and </span>
``manipulate``  packagesin R-studio and then run </span>

``FindRegLine()`` , you get achance to try to find the optimal slope and intercept for a fake data set. 
Click on the "sprocket" icon in the upper left of the plot and you will see
something like Figure 5-17. This interaction can help you see how the residuals
are being measuring in the y-direction and appreciate that </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
lm</span>

 takes care of this for us. 

</span>

```{r}

require(tigerstats)

```



```{r}

require(manipulate)

```



```{r}

FindRegLine()

```



normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>

&nbsp;
</span>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>
Equation of the regression line is:

</span>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>
y = 4. 34 + -0. 02x

</span>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>

&nbsp;
</span>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>
Your final score is 13143. 99 

</span>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>
Thanks for playing!

</span>


<!--[if gte vml 1]>
<v:shape id="Picture_x0020_56" o:spid="_x0000_i1033"
 type="#_x0000_t75" style='width:330pt;height:179. 65pt;visibility:visible;
 mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image062. emz" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=440 height=240src="chapter5_files/image063. png" v:shapes="Picture_x0020_56">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-17: Results of
running </span>
*``FindRegLine()`` *<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
 where the user (Greenwood) didn?t
quite find the least squares line. The correct line is the bold (red) line and
produced a smaller sum of squared residuals than the guessed thinner (black)
line. </span>
*

</span>

????????????? </span>
It ends up that the least squares
criterion does not require a search across coefficients or trial and error?
there are some "simple" equations available for calculating the estimates of
the y-intercept and slope: 

</span>


<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:fPr>
<m:num>
<m:nary>
<m:naryPr>
<m:chr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:naryPr>
<m:sub>
*<span style='font-size:14. 0pt;    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
     font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
     14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;     mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
     mso-bidi-font-weight:bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
    font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
    bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
</m:d>
</m:num>
<m:den>
<m:nary>
<m:naryPr>
<m:chr
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:naryPr>
<m:sub>
*<span style='font-size:14. 0pt;    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
      style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:      12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
      mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
      font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;      mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;      mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
      14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;      mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
      mso-bidi-font-weight:bold;font-style:italic;mso-bidi-font-style:
      normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;      mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
</m:d>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
</m:e>
</m:nary>
</m:den>
</m:f>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:sub>
</m:sSub>
</m:den>
</m:f>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-10. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:145. 5pt;height:28. 15pt'>

 <v:imagedata src="chapter5_files/image064. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=194 height=38src="chapter5_files/image065. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
  bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
  bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:77. 65pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image066. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=104 height=24src="chapter5_files/image067. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. 

</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;mso-bidi-font-weight:bold'>
You will never
need to use these equations but they do inform some properties of the
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 15pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image043. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=18 height=24src="chapter5_files/image044. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, is based on
the variability in *x* and *y* and the correlation between them. If  ***r*** =0, then the slope coefficient will
also be 0. The intercept is a function of the means of x and *y* and what the estimated slopecoefficient is **** </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
b</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
1</m:r>
</span>
*** </m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:21pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image068. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=28 height=24src="chapter5_files/image069. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
is0, then </span>
** <!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
b</m:r>
</span>
*** </m:e>
<m:sub>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
0</m:r>
</span>
*** </m:sub>
</m:sSub>
 ***<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
  m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
=</m:r>
</span>
*** <m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-weight:bold;
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:35. 65pt;height:15pt'>

 <v:imagedata src="chapter5_files/image070. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=48 height=20src="chapter5_files/image071. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
(which is just the mean of the response? this is a very boring model!). The slope is 0 when the correlation is 0. Sowhen there is no linear relationship between x and y (*r*=0), the least squares regression line is a horizontal line with height  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:8. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image072. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=11 height=24src="chapter5_files/image073. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, and the line
produces the same fitted values for all x values. You can also think about this
as when there is no relationship between x and *y*, the best prediction of *y*is the mean of the y-values and it doesn?t change based on the values of x. It
is less obvious in these equations, but they also imply that  **** </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
<m:chr  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
  minor-fareast;mso-hansi-font-family:"Cambria Math";font-weight:bold;
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
x</m:r>
</span>
*** </m:e>
</m:acc>
 ***<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
 mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
  m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
, </m:r>
</span>
*** <m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
  minor-fareast;mso-hansi-font-family:"Cambria Math";font-weight:bold;
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:18. 4pt;height:15pt'>

 <v:imagedata src="chapter5_files/image074. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=25 height=20src="chapter5_files/image075. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;mso-bidi-font-weight:bold'>
 It provides a sort of anchor point
for all regression lines. </span>



</span>

????????????? </span>
For one more example, we can
revisit the Montana wildfire areas burned (log-hectares) and the average summer
0. 81. The interpretations of the different parts of the regression model followthe least squares estimation provided by </span>

``lm`` :

</span>

```{r}



```


data=</span>

mtfires)</span>
 **summary</span>
## Coefficients:</span>
 **##<span style='mso-spacerun:yes'>
????????????</span>
Estimate Std. Error t value Pr(&gt;|t|)<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept) -69. 7845<spanstyle='mso-spacerun:yes'>
??? </span>
12. 3132<span style='mso-spacerun:yes'>
?</span>
-5. 667 1. 26e-05 ***</span>
 **## Temperature<span style='mso-spacerun:yes'>
??</span>
1. 3884<span style='mso-spacerun:yes'>
???? </span>
0. 2165<spanstyle='mso-spacerun:yes'>
?? </span>
6. 412 2. 35e-06 ***</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Regression Equation   (Completely Specified):

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
     font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;    style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;    mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;
    mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
    mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
    position:relative;top:4. 5pt;mso-text-raise:-4. 5pt;mso-ansi-language:
    EN-US;mso-fareast-language:EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape
    id="_x0000_i1025" type="#_x0000_t75" style='width:53. 65pt;height:19. 15pt'>

    <v:imagedata src="chapter5_files/image076. png" o:title="" chromakey="white"/>

    </v:shape>
<![endif]-->
<![if !vml]>
<img width=72 height=26    src="chapter5_files/image077. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
    mso-bidi-font-weight:bold'>
?</span>
=    -69. 78 +1. 39 Temp </span>
????????????? </span>
OR

</span>
</li>
  <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
     font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;    style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;    mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;
    mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
    mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
    position:relative;top:4. 5pt;mso-text-raise:-4. 5pt;mso-ansi-language:
    EN-US;mso-fareast-language:EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape
    id="_x0000_i1025" type="#_x0000_t75" style='width:8. 25pt;height:17. 65pt'>

    <v:imagedata src="chapter5_files/image037. png" o:title="" chromakey="white"/>

    </v:shape>
<![endif]-->
<![if !vml]>
<img width=11 height=24    src="chapter5_files/image038. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
    mso-bidi-font-weight:bold'>
= -69. 78+ 1. 39 x </span>

? </span>
with  **** 

</span>
</li>
 </ul>
 </ul>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Response Variable: Yearly *log* Hectares burned by wildfires 

</span>
</li>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Explanatory Variable: Average   Summer Temperature

</span>
</li>

</ul>


style='font-size:9. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Estimated y-Intercept (</span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
    font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;   mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;
   mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
   mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
   position:relative;top:4. 5pt;mso-text-raise:-4. 5pt;mso-ansi-language:EN-US;
   mso-fareast-language:EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape
   id="_x0000_i1025" type="#_x0000_t75" style='width:13. 5pt;height:17. 65pt'>

   <v:imagedata src="chapter5_files/image041. png" o:title="" chromakey="white"/>

   </v:shape>
<![endif]-->
<![if !vml]>
<img width=18 height=24   src="chapter5_files/image042. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
):
   -69. 78

</span>
</li>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Estimated slope (</span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;
    font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;   mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;
   mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
   mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
   position:relative;top:4. 5pt;mso-text-raise:-4. 5pt;mso-ansi-language:EN-US;
   mso-fareast-language:EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape
   id="_x0000_i1025" type="#_x0000_t75" style='width:13. 15pt;height:17. 65pt'>

   <v:imagedata src="chapter5_files/image043. png" o:title="" chromakey="white"/>

   </v:shape>
<![endif]-->
<![if !vml]>
<img width=18 height=24   src="chapter5_files/image044. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span   style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
   mso-fareast-theme-font:minor-fareast;mso-bidi-font-weight:bold'>
)</span>

: 1. 39

</span>
</li>
</ul>


style='font-size:9. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Slope Interpretation:<span   style='mso-spacerun:yes'>
? </span>
For a 1 degree Fahrenheit increase in
   Average Summer Temperature we would expect,  **** , a 1. 39 <i>
<u>
change</u>
*   in log(Hectares) burned. 

</span>
</li>
</ul>


style='font-size:9. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;mso-list:
   l5 level1 lfo6;tab-stops:list . 5in'>
Y-intercept   Interpretation:<span style='mso-spacerun:yes'>
? </span>
If temperature were   0 degrees F, we would expect -69. 78 log(Hectares) burned on average. 

</span>
</li>
</ul>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1032" type="#_x0000_t75" style='width:370. 9pt;height:231. 75pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image078. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=495 height=309src="chapter5_files/image079. jpg" v:shapes="_x0000_i1032">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-18: Scatterplot of
log-hectares burned versus temperature with estimated regression line. </span>
*

</span>


&nbsp;
</span>


One other use of regression equations is for prediction. It is a trivial
exercise (or maybe not ? we?ll see when you try it!) to plug an x-value of
interest into the regression equation and get an estimate for *y* at that x. Basically, the regressionlines displayed in the scatterplots show the predictions from the regression
involves estimating the response for a particular value of x. We know that itwon?t be perfect but it is our best guess. Suppose that we are interested in
predicting the log-area burned for a summer that had an average temperature of
59</span>

Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F. If we plug 59</span>
line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;
<spanstyle='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F into the  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:46. 15pt;height:16. 5pt'>

 <v:imagedata src="chapter5_files/image080. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=62 height=22src="chapter5_files/image081. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
= -69. 78 +1. 39?Temp, we get

</span>

<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:46. 15pt;height:16. 5pt'>

 <v:imagedata src="chapter5_files/image080. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=62 height=22src="chapter5_files/image081. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
= -69. 78log-hectacres+1. 39log-hectacres/</span>
style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F ? 59</span>
line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;
<spanstyle='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F 

</span>

=
12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:
minor-latin;mso-char-type:symbol;mso-symbol-font-family:Symbol;mso-bidi-font-weight:
bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
</s>
<s>
F</span>
</s>
<spanstyle='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
</s>
<s>
F</span>
</s>
 

</span>

= 12. 23
log-hectares. 

</span>


We did notobserve any summers at exactly *x*=59but did observe some nearby and this result seems relatively reasonable. 

</span>

Nowsuppose someone asks you to use this equation for predicting *Temperature*=65</span>
style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F. We can run that through the equation:-69. 78+1. 39*65 = 20. 57 log-hectares but can we trust this prediction? We did
not observe any summers over 60 degrees F so we are now predicting outside the
scope of our observations ? performing extrapolation. Having a scatterplot in
hand helps us to assess the range of values where we can reasonably use the
equation ? between 54 and 60 degrees F. 

</span>


```{r}



```

(loghectacres~Temperature, </span>
mtfires, </span>

T, </span>

"Scatterplot with regression line for Area burned vs
Temperature"</span>

</div>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020929">
5. 7: Measuring the strength of regressions: R<sup>
2</sup>
</a>
<sup>


</sup>
</h2>


????????????? </span>
At the beginning of the chapter, 
we used the correlation coefficient to measure the strength and direction of
the linear relationship. The regression line provides an even more detailed
description of the direction of the linear relationship than the correlation
provided; in regression we addressed the question of "for a unit change in x, 
on average?" whereas the correlation just addressed whether the relationshipwas positive or negative. However, the  ****   **** . Consider the three scatterplotsin Figure 5-19: the left panel is the original BAC data and the two right
panels have fake data that generated the same estimated regression model with a
weaker (middle panel) and then a stronger (right panel) linear relationship
between *Beers* and *BAC*. This suggests that the regressionline is a useful but incomplete characterization of relationships between
variables ? we need a measure of strength of the relationship to go with the
equation. 

</span>


Weagain to characterize strength but it is somewhat redundant to report a measurethat contains direction information. It also will not extend to multiple
regression models where we have more than one predictor variable in the same
model. 

</span>


Inregression models, we use the  ***coefficientof determination***  (symbol:  **R<sup>
2</sup>
** ) to accompany ourregression line and describe the strength of the relationship. It can either be
scaled between 0 and 1 or 0 to 100% and has "units" of the proportion or
percentage of the variation in *Y* thatis explained by the model that includes *x*corresponds to explaining 0% of the variation in the response with our modelwas explained by the model. In between, it provides a nice summary of how much ofthe total variability in the response we can account for with our model
including x (and, in Chapter 7, including multiple predictor variables). 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1031" type="#_x0000_t75" style='width:466. 15pt;height:291. 4pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image082. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=622 height=389src="chapter5_files/image083. jpg" v:shapes="_x0000_i1031">
<![endif]>
</span>
style='font-size:12. 0pt;line-height:107%;mso-bidi-font-weight:bold;mso-no-proof:yes'>


</span>


*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-19: Three
scatterplots with the same estimated regression line. 

</span>
*


&nbsp;
</span>


the model fit, here we call it </span>

SS<sub>
regression</sub>
</span>
, and the residualvariability, still </span>

SS<sub>
error</sub>
 = </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:nary>
<m:naryPr>
<m:chr  m:val="&#8721;"/>
<m:limLoc m:val="undOvr"/>
<m:subHide m:val="on"/>
<m:supHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:dPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
</m:d>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:60. 75pt;height:18pt'>

 <v:imagedata src="chapter5_files/image084. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=81 height=24src="chapter5_files/image085. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. The </span>

SS<sub>
regression</sub>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 is most easily calculated as </span>

SS<sub>
regression</sub>
 = SS<sub>
Total</sub>
</span>
style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 ? SS<sub>
error</sub>
</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
, the difference between the total
variability and the variability not explained by the model under consideration. 
Using these quantities, we calculate the portion of the total variability that
the model explains as</span>

 

</span>

 **R</span>
** <sup>
2</span>
</sup>
<span  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SS</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SS</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
Total</m:r>
</span>
*</m:sub>
</m:sSub>
</m:den>
</m:f>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SS</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SS</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-10. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:133. 9pt;height:28. 15pt'>

 <v:imagedata src="chapter5_files/image086. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=179 height=38src="chapter5_files/image087. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. 

</span>


????????????? </span>
It also ends up that the
coefficient of determination for models with one predictor is the correlation
coefficient ( ***r*** ) squared ( **** <sup>
2</sup>
=  ***r*** <sup>
2</sup>
). So we canquickly find coefficients of determination if we know correlations in simple
linear regression models. In the real *Beers*and *BAC* data,  ***r*** =0. 8943. So  **** <sup>
2</sup>
 = 0. 79998 or approximately0. 80. So 80% of the variation in *BAC*That leaves 20% of the variation in the responses to be unexplained by ourmodel. In this case much of the unexplained variation is likely attributable to
differences in physical characteristics (that were not measured) but the
statistical model places that unexplained variation into the category of
"random errors". We don?t actually have to find  ***r***  to get coefficients of determination ? the result is part ofthe regular summary of a regression model that we have not been reproducing. 
The full </span>

``lm``  model summary follows with``Multiple R-squared`` 
" that is bolded below. Itis reported as a proportion and it is your choice whether you want to report
and interpret it as a proportion or percentage, just make that clear in how you
discuss it. 

</span>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
lm</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(BAC~Beers, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
data=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
BBfake5)</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
summary</span>
</span>
<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(m1)</span>
</span>



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Call:</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## lm(formula = BAC ~ Beers, data = BBfake5)</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Residuals:</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
??????</span>
Min<span style='mso-spacerun:yes'>
??????? </span>
1Q<spanstyle='mso-spacerun:yes'>
??? </span>
Median<spanstyle='mso-spacerun:yes'>
??????? </span>
3Q<span style='mso-spacerun:yes'>
??????</span>
Max </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## -0. 033807 -0. 023807 -0. 002045<spanstyle='mso-spacerun:yes'>
? </span>
0. 004432<span style='mso-spacerun:yes'>
?</span>
0. 106193 </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Coefficients:</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
?????????????</span>
Estimate Std. Error t value Pr(&gt;|t|)<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## (Intercept) -0. 007386<spanstyle='mso-spacerun:yes'>
?? </span>
0. 021254<span style='mso-spacerun:yes'>
?</span>
-0. 348 0. 733018<span style='mso-spacerun:yes'>
??? </span>
</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Beers<span style='mso-spacerun:yes'>
????</span>
???</span>
0. 018239<spanstyle='mso-spacerun:yes'>
?? </span>
0. 004051<span style='mso-spacerun:yes'>
??</span>
4. 503 0. 000421 ***</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## ---</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Signif. codes:<spanstyle='mso-spacerun:yes'>
? </span>
0 '***' 0. 001 '**' 0. 01 '*' 0. 05 '. ' 0. 1 ' '1</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Residual standard error: 0. 03448 on 15
degrees of freedom</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##  **MultipleR-squared:<span style='mso-spacerun:yes'>
? </span>
** , AdjustedR-squared:<span style='mso-spacerun:yes'>
? </span>
0. 5464 </span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## F-statistic: 20. 27 on 1 and 15 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: 0. 0004211</span>
</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


In thisoutput, be careful because there is another related quantity called  ***Adjusted***  that we willdiscuss later. This other quantity is not a measure of the strength of the
relationship but will be useful. 

</span>


????????????? </span>
We could also revisit the ANOVA
table for this model to verify the source of the  **** <sup>
2</sup>
 of 0. 80based on </span>
SS<sub>
regression</sub>
 </span>
= 0. 02337 and</span>
 </span>
12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
SS<sub>
Total</sub>
 </span>

line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
= 0. 02337+0. 00585</span>

. </span>

line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 This provides 0. 80 from 0. 02337/0. 02922. </span>



</span>

```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(m1)</span>
</span>




<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##Analysis of Variance Table</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Response: BAC</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##<span style='mso-spacerun:yes'>
??????????</span>
Df<span style='mso-spacerun:yes'>
??? </span>
Sum Sq<spanstyle='mso-spacerun:yes'>
?? </span>
Mean Sq F value<spanstyle='mso-spacerun:yes'>
??? </span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
???</span>
</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Beers<span style='mso-spacerun:yes'>
?????</span>
1 0. 0233753 0. 0233753<span style='mso-spacerun:yes'>
? </span>
55. 9442. 969e-06 ***</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
## Residuals 14 0. 0058497 0. 0004178<spanstyle='mso-spacerun:yes'>
????????????????????? </span>
</span>
</span>
 **
<![if !supportLineBreakNewLine]>
 **<![endif]>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>


</span>
</span>


<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
SStotal&lt;-</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
0. 0233753+0. 0058497</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
SSregression&lt;-</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
0. 0233753</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
SSregression/SStotal</span>
</span>




<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##[1] 0. 7998392</span>
</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


????????????? </span>
In Figure 5-19, we saw three
examples with the same regression model, but different strengths of
relationships. In the real data set  **** <sup>
2</sup>
 = 80%. For the firstfake data set (middle panel), the  **** <sup>
2</sup>
 drops to 13. 8% and forthe second fake data set (right panel),  **** <sup>
2</sup>
 is 97. 3%. As asummary,  **** <sup>
2</sup>
 provides a natural scale to understand "how good"each model is at explaining the responses. We can revisit some of our previous
models to get a little more practice with using this summary of strength or quality
of regression models. 

</span>


leaves quite a bit of unexplained variation in the responses. There is a long
list of reasons why this explanatory variable leaves a lot of variation in the
response unexplained. Note that we were careful about using the scaling of the
response variable (log(area burned)) in the interpretation ? this is because we
would get a much different answer if area burned vs temperature was considered. 


</span>


```{r}



```


data=</span>

mtfires)</span>
 **summary</span>
## Residualstandard error: 1. 476 on 21 degrees of freedom</span>
 **##  **MultipleR-squared:<span style='mso-spacerun:yes'>
? </span>
** , AdjustedR-squared:<span style='mso-spacerun:yes'>
? </span>
0. 6458 </span>
 **## F-statistic: 41. 12 on 1 and 21 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: 2. 347e-06</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


 **????????????? </span>
</span>
** For the
model for female Australian athletes that used *Body fat* to explain *Hematocrit*,   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
   mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
Hc</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
   mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
</m:acc>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  "Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
t</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:192pt;height:18. 4pt'>

 <v:imagedata src="chapter5_files/image050. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=256 height=25src="chapter5_files/image051. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and  ***r*** =-0. 168. The coefficient of determination is R<sup>
2</sup>
 = (-0. 168)<sup>
2</sup>
 =
0. 0282. So *body fat* explains 2. 8% ofthe variation in *Hematocrit* in thesewomen. That is not a very good regression model with over 97% of the variation
model. The scatterplot showed a fairly weak relationship but this providesnumerical and interpretable information that drives that point home. 

</span>


```{r}



```


data=</span>

1</span>

#Results for Females </span>
 **summary</span>
## Residualstandard error: 2. 598 on 97 degrees of freedom</span>
 **##  **MultipleR-squared:<span style='mso-spacerun:yes'>
? </span>
** , <spanstyle='mso-spacerun:yes'>
??? </span>
Adjusted R-squared:<spanstyle='mso-spacerun:yes'>
? </span>
0. 0182 </span>
 **## F-statistic: 2. 816 on 1 and 97 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: 0. 09653</span>



</div>


 **
&nbsp;
</span>
** 
<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020930">
5. 8: Outliers: leverage and influence</a>


</h2>


????????????? </span>
In the review of correlation, we
loosely considered the impacts of outliers on the correlation. We removed
unusual points to see both the visual changes (in the scatterplot) as well as
changes in the correlation coefficient in Figures 5-4 and 5-5. In this section, 
we formalize these ideas in the context of impacts of unusual points on our
regression equation. In regression, it is possible for a single point to have a
big impact on the overall regression results but it is also possible to have a
clear outlier that has little impact on the results. We call an observation  ***influential***  if its removal causes a"big" change in the regression line, specifically in terms of impacting the
slope coefficient. Points that are on the edges of the x?s (far from the mean
of the x?s) have the potential for more impact on the line as we will see in
some examples shortly. 

</span>


Youcan think of the regression line being balanced at </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
  bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:6. 75pt;height:15pt'>

 <v:imagedata src="chapter5_files/image088. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=9 height=20src="chapter5_files/image089. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
and the further from that location a point is, the more a single point can move the line. We can measure the distance of
points from </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
<m:chr
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:
  bold;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:6. 75pt;height:15pt'>

 <v:imagedata src="chapter5_files/image088. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=9 height=20src="chapter5_files/image089. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
to quantify each observation?s potential forimpact on the line using what is called the  ***leverage*** of a point. Leverage is a positive numerical measurewith larger values corresponding to more leverage. The scale changes depending
complexity of the model so all that matters is which observations have more orless relative leverage in a particular data set. The observations with x-values
that provide higher leverage have increased potential to influence the
estimated regression line. Along with measuring the leverage, we can also
measure the influence that each point has on the regression line using  ***Cook?s Distance***  or  ***Cook?s D*** . It also is a positive
measure with higher values suggesting more influence. The rule of thumb is that
Cook?s D values over 1. 0 correspond to clearly influential points, values over
0. 5 have some influence and values lower than 0. 5 indicate points that are not
influential on the regression model slope coefficients. One part of the regular
diagnostic plots we will use for regression models displays the leverages on
the x-axis, the standardized residuals on the y-axis, and adds contour lines
for Cook?s Distances in a panel that is labeled "Residuals vs Leverage". This
allows us to see the potential for impact of a point (leverage), how far it?s
observation was from the regression line (residual), and to see a measure of
that point?s influence (Cook?s D). 

</span>


Toextract the level of Cook?s D on the "Residuals vs Leverage" plot, look for contours
to show up on the upper and lower right of the plot. They show increasing
levels of influence going to the upper and lower right corners as you combine
higher leverage (x-axis) and larger residuals (y-axis) ? the two ingredients
required to be influential on the line. The contours are displayed for Cook?s D
values of 0. 5 and 1. 0 if there are points near or over those levels. The Cook?s
D values come from a topographical surface of values that is a sort of U-shaped
valley in the middle of the plot centered at y=0 with the lowest contour
corresponding to Cook?s D values below 0. 5 (no influence). As you move to the
upper right or lower right corners, the influence increases and the edges of
the valley get steeper. If you do not see any contours in the plot, then no
points were even close to being influential based on Cook?s D. 

</span>


Toillustrate these concepts, the original Beers and BAC data are used again. In
the scatter plot in Figure 5-20, two points are plotted with different characters. 
The point for 1 Beer and BAC of 0. 010 is displayed as a "</span>

style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
" and the 9 Beer and BAC 0. 19 observation isdisplayed with a "o". These two points are the furthest from the mean of the
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:31. 15pt;height:15. 4pt'>

 <v:imagedata src="chapter5_files/image090. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=42 height=21src="chapter5_files/image091. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
= 4. 8) but show two different levels ofinfluence on the line. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1030" type="#_x0000_t75" style='width:394. 5pt;height:246. 75pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image092. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=526 height=329src="chapter5_files/image093. jpg" v:shapes="_x0000_i1030">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-20: Scatterplot and
Residuals vs Leverage plot for the real BAC data. 

</span>
*


&nbsp;
</span>


The "</span>
style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
Symbol;mso-bidi-font-weight:bold'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
" point has a leverage of 0. 27 and the 9 Beerobservation ("o") had a leverage of 0. 30. The 1 beer observation was close to
the pattern defined by the other points, had a small residual, and a Cook?s D
value below 0. 5 (it did not exceed the first of the contours). So even though
it had high leverage, it was not an influential point. The 9 beer observation
had the highest leverage in the data set and was quite a bit above the pattern
defined by the other points and ends up being an influential point with a
Cook?s D over 1. We might want to consider fitting this model without that
observation to get a better estimate of the effects of beer consumption on BAC
or revisit our assumption that the relationship is really linear here. 

</span>


Tofurther explore influence, we will add a point to the original data set and
move it around so you can see how those changes impact the results. For each
scatterplot in Figure 5-21, the Residuals vs Leverage plot is displayed to its
right. The original data are "<span style='color:#AEAAAA;mso-themecolor:background2;mso-themeshade:191'>
?</span>
" and the original regression line is the dashedline in Figure 5-21. First, a fake observation at 11 Beers and 0. 1 BAC is added, 
at (11, 0. 1), in the top panels of the figure. This observation is clearly an
outlier and heavily impacts the slope of the regression line (so is clearly influential). 
to 0. 24. The accompanying Residuals vs Leverage plot shows that this point hasextremely high leverage and a Cook?s D over 1 ? it is a clearly influential
Consider the second row of plots with an added point of (11, 0. 19). TheThis point has the same leverage as in the first example since it is the sameset of x?s and the distance to the mean is unchanged. But it is not influential
since its Cook?s D is less than 0. 5. This occurred because it followed the
overall pattern of observations even though it was "far away" from the other
observations in the x-direction. The last two rows of plots show what happens
when low leverage outliers are encountered. Placing observations near the
center of the x?s means that to be influential the points have to be very far
from the pattern of the other observations. The (5, 0. 19) example almost
attains a Cook?s D of 0. 5 but has little impact on the regression line, 
especially the slope coefficient. It does impact the y-intercept and drops the
R-squared value to 0. 57. The same result occurs if the observation is
noticeably lower than the other points. 

</span>


Whenwe are doing regressions, we get very worried about points "at the edges"
having an undue influence on the results. When we start using multiple
predictors, say if we had body weight data on these subjects as well as beer
consumption, it becomes harder to "see" if the points are "far away" from the
other observations and we will trust the Residuals vs Leverage plots to help us
identify the influential points. These techniques work the same in the multiple
regression models in Chapter 7 as they do in these simpler, single predictor
regression models. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1029" type="#_x0000_t75" style='width:7in;height:630pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image094. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=672 height=840src="chapter5_files/image095. jpg" v:shapes="_x0000_i1029">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-21: Plots exploring
the impacts of moving a single additional observation. </span>
*

</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020931">
5. 9: Residual diagnostics ? setting the stage for
inference</a>


</h2>


????????????? </span>
Influential points are not the
only potential issue that can cause us to have concerns about our regression
model. There are two levels to these considerations. The first is related to
issues that directly impact the least squares regression line and cause
concerns about whether a line is a reasonable representation of the
relationship between the two variables. These issues for regression model
estimation have been discussed previously (the same concerns in estimating
correlation apply to regression models). The second level is whether the line
we have will be useful for making inferences for the population that our data
were collected from and whether the data follow our assumed model. Our window
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
e</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
   italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:69. 4pt;height:17. 65pt'>

 <v:imagedata src="chapter5_files/image096. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=93 height=24src="chapter5_files/image097. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
)</span>

. Byexploring patterns in how the line "misses" the responses we can gain much
information about the reasonableness of using the estimated regression line and
sometimes information about how we might fix problems. The validity conditions
for doing inference in a regression setting (Chapter 6) involve two sets of considerations, 
those that are assessed based on the data collection and measurement process
and those that can be assessed using diagnostic plots. The first set is:

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **Quantitative variables condition

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
We?ll discuss using   categorical predictor variables later ? to use simple linear regression
   both the explanatory and response variables need to quantitative. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **Independence of observations

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
As in the ANOVA models,    linear regression models assume that the observations are collected in a
   fashion that makes them independent. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
This will be based on the   "story" of the data. Consult a statistician if your data violate this
   assumption as there are more advanced methods that adjust for dependency
   in observations but they are beyond the scope of this material. 

</span>
</li>

 </ul>

</ul>


Theremaining assumptions for getting valid inferences from regression models can
be assessed using diagnostic plots. 

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **Linearity of relationship

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
We should not report a linear   regression model if the data show a curve (curvilinear relationship
   between x and y). 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
Examine the Residuals vs Fitted   plot:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
If the model missed a    curve in the relationship, the residuals often will highlight that
    missed pattern and a curve will show up in this plot. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Try to explain or understand    the pattern in what is left over. If we have a good model, there
    shouldn?t be much left to "explain" in the residuals (i. e. , no patterns
    left over after accounting for x). 

</span>
</li>

 </ul>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **Equal (constant) variance

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
We assume that the variation   is the same for all the observations and especially that the variability
   does not change in the responses as a function of our predictor variables
   or the fitted values. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
There are two plots to help   with this:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Examine the Residuals    vs Fitted plot and look for evidence of changing spread in the
    residuals, being careful to try to separate curving patterns from
    non-constant variance (and look for situations where both are present as
    you can violate both conditions simultaneously). 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Examine the "Scale-Location"    plot and look for changing spread as a function of the fitted values. 

</span>
</li>

  <ul style='margin-top:0in' type=disc>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level4 lfo12'>
The y-axis in this    plot is the square-root of the absolute value of the standardized
    residual. This scale flips the negative residuals on top of the
    positive ones to help you better assess changing variability without
    being distracted by whether the residuals are above or below 0. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level4 lfo12'>
Because of the    absolute value, curves in the Residuals vs Fitted plot can present as
    sort of looking like non-constant variance ? check for nonlinearity in
    the residuals vs fitted before using this plot. If nonlinearity is
    present, just use the Residuals vs Fitted for assessing constant
    variance around curving pattern. 

</span>
</li>

  </ul>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
If there are patterns of   increasing or decreasing variation (often described as funnel or cone shapes), 
   then it might be possible to use a transformation to fix this problem
   (more later). It is possible to have decreasing and then increasing
   variability in one plot and this also is a violation of this condition. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **Normality of residuals

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
Examine the Normal QQ plot   for violations of the normality assumption as in Chapters 2 and 3. 

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Specifically review the    discussion of identifying skews in different directions and heavy vs
    light tailed distributions. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Skewed and    heavy-tailed distributions are the main problems for our inferences, 
    especially since both kinds of distributions can contain outliers that
    can wreak havoc on the estimated regression line. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Light-tailed    distributions cause us no real inference issues except that the results
    are conservative so you should note when you observe these situations
    but feel free to proceed with using your model results. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l4 level3 lfo12'>
Remember that clear    outliers are an example of a violation of the normality assumption but
    some outliers may just influence the regression line and make it fit
    poorly for the results of the observations and this issue will be more
    clearly observed in the residuals vs fitted. 

</span>
</li>

 </ul>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level1 lfo12'>
 **No influential points:

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
Examine the Residuals vs   Leverage plot as discussed in the previous section. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l4 level2 lfo12'>
Consider removing influential   points and focusing on results without that point in the data set. 

</span>
</li>

 </ul>

</ul>


To assessthese later assumptions, we will use the four residual diagnostic plots that R
provides from </span>

``lm``  fitted models. They aresimilar to the results from ANOVA models but the Residuals vs Leverage plot is
now interesting as was discussed in Section 5. 8. Now we can fully assess the
potential for trusting the estimated regression models in a couple of our
examples:

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
 **Beers vs BAC:

</span>
** </li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Quantitative variables   condition:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
Both variables are    quantitative. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Independence of observations:

</span>
</li>
 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpLast style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
We can assume that    all the subjects are independent of each other. There is only one
    measurement per student and it is unlikely that one subject?s beer
    consumption would impact another?s BAC. Unless the students were trading
    blood it is isn?t possible for one person?s beer consumption to change
    someone else?s BAC. 

</span>
</li>

 </ul>

 </ul>

</ul>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
lm</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(BAC~Beers, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
data=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
BB)</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
par</span>
</span>
<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
mfrow=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
c</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
2</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
2</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
))</span>
</span>
 **<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
plot</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(m1, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
add. smooth=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
F, </span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
main=</span>
</span>

<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
"Beersvs BAC"</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
)</span>
</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1028" type="#_x0000_t75" style='width:400. 9pt;height:250. 5pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image098. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=535 height=334src="chapter5_files/image099. jpg" v:shapes="_x0000_i1028">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-22: Full suite of
diagnostics plots for Beer vs BAC data. </span>
*

</span>

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<ul style='margin-top:0in' type=disc>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Linearity, constant variance   from Residuals vs Fitted:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
We previously have    identified a potentially influential outlying point in these data. 
    Consulting the Residuals vs Fitted plot in Figure 5-22, if you trust
    that influential point, shows some curvature with a pattern of
    decreasing residuals and then an increase at the right. Or, if you do
    not trust that highest BAC observation, then there is a mostly linear
    relationship with an outlier identified. We would probably suggest that
    it is an outlier, should be removed from the analysis, and inferences
    constrained to the region of beer consumption from 1 to 8 beers since we
    don?t know what might happen at higher values. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Constant variance from   Scale-Location:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
There is some    evidence of increasing variability in this plot as the spread of the
    results increases from left to right, however this is just an artifact
    of the pattern in the original residuals and not real evidence of
    non-constant variance. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Normality from Normal QQ   Plot: 

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
The left tail is a    little short and the right tail is a little long, suggesting a right
    skewed distribution in the residuals. This corresponds to having a large
    positive outlying value. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Influential points from   Residuals vs Leverage:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
Previously discussed,     this plot shows one influential point with a Cook?s D value over 1 that
    is distorting the fitted model. 

</span>
</li>

 </ul>

 </ul>

</ul>


``"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>

&nbsp;
`` 

<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
 **Tree height and tree diameter</span>
**  (suspicious
   observation already removed):

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Quantitative variables: Met

</span>
</li>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Independence of observations:

</span>
</li>
 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
There are multiple    trees that were measured in each plot. One problem might be that once a
    tree is established in an area, the other trees might not grow as tall. 
    The other problem is that some sites might have better soil conditions
    than others. Then, all the trees in those rich soil areas might be
    systematically taller than the trees in other areas. Again, there are
    statistical methods to account for this sort of "clustering" of
    measurements but this technically violates the assumption that all the
    trees are independent of each other. So this assumption is violated, but
    we will proceed with that caveat on our results ? the precision of our
    inferences might be slightly over-stated due to some potential
    dependency in the measurements. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Linearity, constant variance   from Residuals vs Fitted in Figure 5-23. 

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
There is evidence of    a curve that was missed by the linear model. 

</span>
</li>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
There is also    evidence of increasing variability AROUND the curve in the residuals. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Constant variance from   Scale-Location:

</span>
</li>

 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
This plot actually shows    relatively constant variance but this plot is misleading when curves are
    present in the data set.  **** 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Normality in Normal QQ plot:

</span>
</li>
 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
There is no indication    of any problem with the normality assumption. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Influential points?

</span>
</li>
 <ul style='margin-top:0in' type=square>

  <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
    mso-add-space:auto;mso-list:l2 level3 lfo13'>
The Cook?s D contours    do not show up in this plot so none of the points are influential. 

</span>
</li>

 </ul>

 </ul>

</ul>


So the mainissues with this model are the curving relationship and non-constant variance. 
We?ll revisit this example later to see if we can find a model on transformed
variables that has better diagnostics. 

</span>



&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1027" type="#_x0000_t75" style='width:396. 75pt;height:248. 25pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image100. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=529 height=331src="chapter5_files/image101. jpg" v:shapes="_x0000_i1027">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-23: Diagnostics
plots for tree height and diameter simple linear regression model. </span>
*

</span>


&nbsp;
</span>


Reportingthe following regression model that has a decent R<sup>
2</sup>
 of 62. 6% would
be misleading since it does not accurately represent the relationship between
tree diameter and tree height. 

</span>


```{r}



```


data=</span>

168</span>

, ])</span>
 **summary</span>
##Coefficients:</span>
 **##<span style='mso-spacerun:yes'>
????????????</span>
Estimate Std. Error t value Pr(&gt;|t|)<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept) 11. 98364<spanstyle='mso-spacerun:yes'>
??? </span>
0. 57422<span style='mso-spacerun:yes'>
??</span>
20. 87<span style='mso-spacerun:yes'>
?? </span>
&lt;2e-16 ***</span>
 **## dbh. cm<span style='mso-spacerun:yes'>
??????</span>
0. 32939<span style='mso-spacerun:yes'>
??? </span>
0. 01395<spanstyle='mso-spacerun:yes'>
?? </span>
23. 61<span style='mso-spacerun:yes'>
??</span>
&lt;2e-16 ***</span>
 **## </span>
 **## Residual standard error: 4. 413 on 333 degrees offreedom</span>
 **## Multiple R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 626, <span style='mso-spacerun:yes'>
? </span>
Adjusted R-squared:<spanstyle='mso-spacerun:yes'>
? </span>
0. 6249 </span>
 **## F-statistic: 557. 4 on 1 and 333 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: &lt; 2. 2e-16</span>



(</span>

c</span>

(</span>
2</span>
, </span>
2</span>
))</span>
 **plot</span>
(tree1, </span>
add. smooth=</span>
</div>


style='font-size:8. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<aname="_Toc459020932">
5. 10: Old Faithful discharge and waiting times</span>
</span>
</a>
</span>
 **

</span>
** 

????????????? </span>
A study in August 1985 considered
time for Old Faithful and how that might relate to *waiting time* for the next eruption (Azzalini and Bowman, 1990). 
This sort of research provides the park staff a way to show tourists a predicted
time to next eruption so they can quickly see it and then get back in their
cars. Both variables are measured in minutes and the scatterplot in Figure 5-24
shows a moderate to strong positive and relatively linear relationship. We
to this plot. Smoothing lines provide regression-like fits but are performed onlocal areas of the relationship and so can highlight where the relationships
change and can highlight curvilinear relationships. They can also return
straight lines just like the regression line if that is reasonable. The
technical details of regression smoothing are not covered here but they are a
useful graphical addition to help visualize nonlinearity in relationships. 

</span>


Inthese data, there appear to be two groups of eruptions (shorter length, shorter
wait and longer length, longer wait) ? but we don?t know enough about these
data to assume that there are two groups. The smoothing line does help us to
see if the relationship appears to change or stay the same across different
The smoothing line suggests that the upper group might have a less steep slopethan the lower group as it sort of levels off for observations with *Durations* over 4 minutes. It alsoindicates that there is one point for an eruption under 1 minute in *Duration* that might be causing someproblems. The story of these data involve some measurements during the night
that were just noted as being short, medium, and long ? and they were re-coded
as 2, 3, or 4 minute duration eruptions. We?ll see if our diagnostics detect
some of these issues when we fit a simple linear regression to try to explain
waiting time based on duration of prior eruption. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1026" type="#_x0000_t75" style='width:6in;height:270pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image102. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=576 height=360src="chapter5_files/image103. jpg" v:shapes="_x0000_i1026">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-24: Scatterplot of
Old Faithful waiting times to next eruption (minutes) and duration of prior
eruption (minutes) with smoothing line (bold, red) and regression line (thin, 
green). </span>
*

</span>

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


????????????? </span>
The first concern with these data
is that the observations are likely not independent. Since they were taken
consecutively, one waiting time might be related to the next waiting time ?
violating the independence assumption. As noted above, there might be two
groups (types) of eruptions ? short ones and long ones. The Normal QQ-Plot in Figure
5-25 also suggests a few observations creating a slightly long right tail. 
Those observations might warrant further exploration as they also show up as
unusual in the Residuals vs Fitted plot. There are no highly influential points
in the data set with all points having Cook?s D smaller than 0. 5, so these
outliers are not necessarily moving the regression line around. There are two
distinct groups of observations but the variability is not clearly changing so
we do not have to worry about non-constant variance here. So these results might
be relatively trustworthy if we assume that the same relationship holds for all
levels of duration of eruptions. 

</span>


```{r}



```

 **data</span>
(geyser)</span>
 **G2&lt;-</span>
(</span>

geyser$waiting[-</span>

], </span>

geyser$duration[-</span>

])</span>
 **scatterplot</span>
data=</span>

spread=</span>

lm</span>

data=</span>

G2)</span>
 **summary</span>
##Coefficients:</span>
 **##<span style='mso-spacerun:yes'>
????????????</span>
Estimate Std. Error t value Pr(&gt;|t|)<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept)<span style='mso-spacerun:yes'>
?</span>
34. 9452<span style='mso-spacerun:yes'>
???? </span>
1. 1807<spanstyle='mso-spacerun:yes'>
?? </span>
29. 60<span style='mso-spacerun:yes'>
??</span>
&lt;2e-16 ***</span>
 **## Duration<span style='mso-spacerun:yes'>
????</span>
10. 7751<span style='mso-spacerun:yes'>
???? </span>
0. 3235<spanstyle='mso-spacerun:yes'>
?? </span>
33. 31<span style='mso-spacerun:yes'>
??</span>
&lt;2e-16 ***</span>
 **## </span>
 **## Residual standard error: 6. 392 on 296 degrees offreedom</span>
 **## Multiple R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 7894, Adjusted R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 7887 </span>
 **## F-statistic:<span style='mso-spacerun:yes'>
?</span>
1110 on 1 and 296 DF, <span style='mso-spacerun:yes'>
? </span>
p-value:&lt; 2. 2e-16</span>



(</span>

c</span>

(</span>
2</span>
, </span>
2</span>
))</span>
 **plot</span>
(OF1)</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75" style='width:383. 25pt;
 height:239. 25pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter5_files/image104. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=511 height=319src="chapter5_files/image105. jpg" v:shapes="Picture">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 5-25: Diagnostic plots
for Old Faithful waiting time model. </span>
*

</span>

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


The  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";mso-bidi-font-weight:bold;font-style:
  italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
WaitingTime</m:r>
</span>
*</m:e>
</m:acc>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:209. 65pt;height:15. 75pt'>

 <v:imagedata src="chapter5_files/image106. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=280 height=21src="chapter5_files/image107. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, suggesting that for a 1 minute increase
in eruption *Duration* we would expect, on average, a 10. 77 minute change in the *waitingtime*. This equation might provide a useful tool for the YNP staff to
predict waiting times The </span>
 **R</span>
** <sup>
2</span>
</sup>
 is fairly large: 78. 9% of thevariation in *waiting time* isexplained by the *duration* of theprevious eruption. But maybe this is more about two types of eruptions/waiting
times? We could consider the relationship within the shorter and longer
eruptions but since there are observations residing between the two groups, it
is difficult to know where to split the explanatory variable into two groups. 
Maybe we really need to measure additional information that might explain why
there are two groups in the responses?

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020933">
5. 11: Chapter summary</a>


</h2>


ThePearson?s Product Moment Correlation Coefficient) measures the strength anddirection of the linear relationship between two quantitative variables. 
Regression models estimate the impacts of changes in x on the mean of the
response variable y. Direction of the assumed relationship (which variable
explains or causes the other) matters for regression models but does not matter
for correlation. Regression lines only describe the pattern of the
relationship; in regression, we use the coefficient of determination to
describe the strength of the relationship between the variables as a percentage
of the response variable that is explained by the model. If we are choosing
between models, we prefer them to have higher R<sup>
2</sup>
 values for obvious
reasons, but we will discover in Chapter 7 that maximizing the coefficient of
determination is not a good way to pick a model when we have multiple candidate
options. 

</span>


Inthis chapter, a wide variety of potential problems were explored when using
regression models. This included a discussion of the conditions that will be
required for using the models to perform trustworthy inferences in the next two
chapters. It is important to remember that correlation and regression models
misleading if a nonlinear relationship is present. Similarly, influentialobservations can completely distort the apparent relationship between variables
and should be checked for before trusting any regression output. It is also
important to remember that regression lines should not be used outside the
scope of the original observations ? extrapolation should be checked for and
avoided whenever possible or at least acknowledged. 

</span>


????????????? </span>
Regression models look like they
estimate the changes in y that are caused by changes in x. This is not true
unless the levels of x are randomly assigned and only then we can make causal
inferences. Since this is not generally true, you should initially always
assume that any regression equation describes the relationship ? if you observe
two subjects that are 1 unit of x apart, you can expect their mean to differ by
</span>
*b<sub>
1$</span>
*? </span>
- you should not, however, say that changingx causes a change in the mean of the responses. Despite all these cautions, 
regression models are very popular statistical methods. They provide detailed
descriptions of relationships between variables and can be extended to
situations where we are interested in multiple predictor variables. They also
share ties to the ANOVA models from earlier this semester. When you are running
R code, you will note that all the ANOVAs and the regression models are all fit
using </span>

``lm`` . The assumptions and diagnostic plots arequite similar. And in the next chapter, we will see that inference techniques
look similar. People still like to distinguish among the different types of
situations, but the underlying  ***linear models*** are actually exactly the same? 

</span>


&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020934">
5. 12: Important R code</a>


</h2>


The maincomponents of the R code used in this chapter follow with the components to
modify in red where y is a response variable, x is an explanatory variable, and
the data are in DATASETNAME. 

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``pairs. panels(<span   style='color:red'>
DATASETNAME`` , ellipses=F, scale=T, smooth=F, col=0)

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Makes a scatterplot matrix   that also displays the correlation coefficient. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Requires the </span>
   style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";   mso-bidi-font-weight:bold'>
psych </span>

package. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``cor(<span   style='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<span   style='color:red'>
DATASETNAME</span>
)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Provides the estimated   correlation coefficient between x and y. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``plot(<span   style='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<span   style='color:red'>
DATASETNAME</span>
)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Provides a scatter plot. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``scatterplot(<span   style='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<span   style='color:red'>
DATASETNAME</span>
, smooth=F)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Provides a scatter plot with   a regression line. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Requires the </span>
   style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";   mso-bidi-font-weight:bold'>
car</span>

 package. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   line-height:107%;font-family:"Lucida Console";color:red;mso-bidi-font-weight:   bold'>
Modelname</span>

``&lt;-lm(<span   style='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<span   style='color:red'>
DATASETNAME</span>
)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Estimates a regression model   using least squares. 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``summary(<span   style='color:red'>
Modelname`` )

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Provides parameter estimates   and R-squared (used heavily in Chapter 6 and 7 as well). 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level1 lfo13'>
``par(mfrow=c(2, 2));   plot(<span style='color:red'>
Modelname`` )

</span>
</li>
 <ul style='margin-top:0in' type=circle>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l2 level2 lfo13'>
Provides four regression   diagnostic plots in one plot. 

</span>
</li>

 </ul>

</ul>


style='font-size:6. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020935">
5. 13: Practice problems</a>


</h2>


Thesequestions revisit the treadmill data set from Chapter 0. Researchers were
interested in whether the run test variable could be used to replace the treadmill
oxygen consumption variable that is expensive to measure. The following code loads
the data set and provides a scatterplot matrix using </span>

style='mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
pairs. panel</span>

. 

</span>

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
treadmill&lt;-read. csv("http://www. math. montana. edu/courses/s217/documents/treadmill. csv")

</span>


style='mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
require(psych)

</span>


style='mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-bidi-font-weight:bold'>
pairs. panels(treadmill, ellipses=F, smooth=F, col=0)

</span>


style='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-weight:bold'>

&nbsp;
</span>


5. 1. First, 
we should get a sense of the strength of the correlation between the variable
of primary interest, </span>

``TreadMillOx`` , and theother variables and consider whether outliers or nonlinearity are going to be
major issues here. Which variable is it most strongly correlated with? Which
variables are next most strongly correlated with this variable?

</span>


5. 2. Fit
the SLR using *RunTime* as explanatoryvariable for *TreadMillOx*. Report theestimated model. 

</span>


5. 3. 
Predict the treadmill oxygen value for a subject with a run time of 14 minutes. 
Repeat for a subject with a run time of 16 minutes. Is there something
different about these two predictions?

</span>


5. 4. 
Interpret the slope coefficient from the estimated model, remembering the units
on the variables. 

</span>


5. 5. Report
and interpret the y-intercept from the SLR. 

</span>


5. 6. Report
the R<sup>
2</sup>
 value from the output. Show how you can find this value from
the original correlation matrix result. 

</span>


5. 7. Produce the diagnostic plots and discuss any potential issues. Whatis the approximate leverage of the highest leverage observation and how large
is its Cook?s D? What does that tell you about its potential influence in this
model?</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:major-fareast'>
<![if !supportFootnotes]>
<spanstyle='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;
[4]</span>
</span>
<![endif]>
</span>
</span>
</span>
</a>

 


</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:major-fareast'>
<![if !supportFootnotes]>
<spanstyle='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;
[5]</span>
</span>
<![endif]>
</span>
</span>
</span>
</a>

 


</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:major-fareast'>
<![if !supportFootnotes]>
<spanstyle='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;
[6]</span>
</span>
<![endif]>
</span>
</span>
</span>
</a>

 



</div>


</div>


** 

</html>

