---
title: "Chapter2 Scratch"
author: "Greta Linse Peterson"
date: "June 4, 2017"
output: html_document
---






Here is a full write-up of the results using all 6+ hypothesis testing steps, using the permutation results:

</span>


????????????? </span>
*Isolate the claim to be proved and method to
use (define a test statistic T)

*</span>


We want to test for a difference in the means between males and females and will use the equal-variance two-sample
t-test statistic to compare them, making a decision at the 5% significance
level. 

</span>


margin-bottom:0in;margin-left:. 75in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;mso-bidi-font-family:Calibri;mso-bidi-theme-font:minor-latin'>
1)<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Write the null and alternative
hypotheses

</span>


margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_0$: </span>
*$\mu$</span>
*<sub>
Male</span>
</sub>
 = </span>
*$\mu$</span>
*<sub>
Female</span>
</sub>
 

</span>

margin-bottom:0in;margin-left:1. 75in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
where</span>
*$\mu$</span>
*<sub>
Male</span>
</sub>
 is the true mean GPA for males and </span>
*$\mu$</span>
*<sub>
Female</span>
</sub>
 is true mean GPA for females. 

</span>

margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
$H_A$: </span>
*$\mu$</span>
*<sub>
Male</span>
</sub>
 &#8800; </span>
*$\mu$</span>
*<sub>
Female</span>
</sub>


</span>


&nbsp;
</span>

????????????? </span>
2)Check conditions for the procedure being used

</span>


0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;text-indent:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
 **Independent observations condition:</span>
**  It appears that this assumption is met because there
is no reason to assume any clustering or grouping of responses that might
create dependence in the observations. The only possible consideration is that
the observations were taken from different sections and there could be some
differences between the sections. However, for overall GPA this not likely to
be a big issue. The only way this could create a violation here is if certain
sections tended to attract students with different GPA levels (such as the 9 am
section had the best/worst GPA students?). 

</span>



&nbsp;
</span>

0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;text-indent:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
 **Equal variance condition</span>
** : There is a small difference in the range of the
observations in the two groups but the standard deviations are very similar so
there is no evidence that this condition is violated. 

</span>



&nbsp;
</span>

0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;text-indent:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
 **Similar distribution condition: </span>
** Based on the side-by-side boxplots and beanplots, it appears
that both groups have slightly left-skewed distributions, which could be
problematic for the parametric approach, but the permutation approach condition
is not violated since the distributions look to have fairly similar shapes. 

</span>



&nbsp;
</span>

????????????? </span>
3)Find the value of the appropriate test statistic

</span>


0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;text-indent:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
*T</span>
*=2. 69 from the previous R output. 

</span>



&nbsp;
</span>

????????????? </span>
4)Find the p-value

</span>


margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
p-value=0. 008 from the permutation
distribution results. 

</span>


margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
This means that there is about a 0. 8%
chance we would observe a difference in mean GPA (female-male or male-female)
of 0. 25 points or more if there in fact no difference in true mean GPA between
females and males in STAT 217 in a particular semester. 

</span>



&nbsp;
</span>

????????????? </span>
5)Decision

</span>


0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;text-indent:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Since the p-value is "small" (*a priori* 5% significance level selected), we can reject the null hypothesis. 

</span>



&nbsp;
</span>

????????????? </span>
6)Conclusion and scope of inference, specific to the problem

</span>


margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
There is strong evidence against the
null hypothesis of no difference in the true mean GPA between males and females
for the STAT 217 students in this semester and so we conclude that there is
evidence of a difference in the mean GPAs between males and females in STAT 217
students. 

</span>


margin-bottom:0in;margin-left:1. 25in;margin-bottom:. 0001pt;mso-add-space:auto;style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-fareast-font-family:Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
Because this was not a randomized
experiment, we can?t say that the difference in sex causes the difference in
mean GPA and because it was not a random sample from a larger population, our
inferences only pertain the STAT 217 students that responded to the survey in
that semester. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo7'>
<a
name="_Toc459020880">
1. 7: Confidence intervals and bootstrapping</a>
</h2>


Randomly shuffling the treatments between the observations is like randomly sampling the treatments
without replacement. In other words, we randomly sample one observation at a
observations. This provides us with a technique for testing hypotheses because it provides new splits of the observations into groups that are as interesting
as what we observed if the null hypothesis is assumed true. In most situations, 
we also want to estimate parameters of interest and provide  ***confidenceintervals***  for those parameters (an interval where we are __%  ***confident*** that the true parameter lies). As before, there are two options we will
consider ? a parametric and a nonparametric approach. The nonparametric
approach will be using what is called  ***bootstrapping***  and draws its name from "pull yourself up by your bootstraps" where you improve your situation
based on your own efforts. In statistics, we make our situation or inferences
better by re-using the observations we have by assuming that the sample
represents the population. Since each observation represents other similar
observations in the population that we didn't get to measure, if we  ***samplewith replacement***  to generate a new data set of size *n* from our data set (also of size *n*) it mimics the process of taking
our population of interest. This process also ends up giving us useful sampling distributions of statistics even when our standard normality assumption is
violated, similar to what we encountered in the permutation tests. 
Bootstrapping is especially useful in situations where we are interested in
statistics other than the mean (say we want a confidence interval for a median
or a standard deviation) or when we consider functions of more than one
parameter and don?t want to derive the distribution of the statistic (say the
difference in two medians). In this text, bootstrapping is used to provide more
trustworthy inferences when some of our assumptions (especially normality)
might be violated for our parametric procedure. 

</span>


????????????? </span>
To perform bootstrapping, we will use the </span>

``resample``  function from the </span>
``mosaic`` to a data set and get a new version of the data set by sampling new observations *with replacement* from the original one. The new, bootstrapped version of the data set (called </span>
``MockJury_BTS``  below) contains a new variable called </span>

``orig. id``  which is the number of the subject from the original data set. By summarizing how often each of these id?s
occurred in a bootstrapped data set, we can see how the re-sampling works. The </span>

``table``  function will count up how many times each observation was used in the bootstrap sample, providing a row with
the id followed by a row with the count<a style='mso-footnote-id:ftn17'
href="#_ftn17" name="_ftnref17" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[17]</span>
</span>
<![endif]>
</span>
</span>
</a>
. 
In the first bootstrap sample shown, the 2<sup>
nd</sup>
, 7<sup>
th</sup>
, and 9<sup>
th</sup>

observations were sampled one time each, the 4<sup>
th</sup>
 observation was
sampled three times, and the 1<sup>
st</sup>
, 3<sup>
rd</sup>
, 5<sup>
th</sup>
, 
and many others were not sampled at all. Bootstrap sampling thus picks some
observations multiple times and to do that it has to ignore some observations. 

</span>


```{r}



```

resample</span>
(MockJury2)</span>
 **table</span>
(</span>
as. numeric</span>
##<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
4<span style='mso-spacerun:yes'>
? </span>
7<spanstyle='mso-spacerun:yes'>
? </span>
9 10 13 14 15 16 17 18 19 25 26 27 29 31 3233 35 38 39 40 41 43 </span>
 **##<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
3<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
? </span>
3<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
3<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
3<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1 </span>
 **## 45 46 49 50 52 54 56 57 58 59 63 64 65 66 68 69 7072 73 74 </span>
 **##<span style='mso-spacerun:yes'>
? </span>
3<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
3<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
2<span</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Like in permutations, one randomization isn't enough. A second bootstrap sample is also provided to help
you get a sense of what it is doing to generate a data set. It did not select
subject 7 but did select 2, 4, 6, and 8 two times. You can see other variations
in the resulting re-sampling of subjects with the most sampled subject being
the chance of selecting any observation for any slot in the new data set is1/75 and the expected or mean number of appearances we expect to see for an
observation is the number of tries times the probably of selection on each so
75*1/75=1. 

</span>


```{r}



```

resample</span>
(MockJury2)</span>
 **table</span>
(</span>
as. numeric</span>
##<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
3<spanstyle='mso-spacerun:yes'>
? </span>
4<span style='mso-spacerun:yes'>
?</span>
5<span style='mso-spacerun:yes'>
? </span>
6<spanstyle='mso-spacerun:yes'>
? </span>
8 10 11 13 14 15 16 17 19 20 23 24 26 27 2930 31 32 33 </span>
 **##<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
3<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
3<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
3<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1 </span>
 **## 34 35 37 38 39 47 48 49 50 51 53 54 55 56 60 62 6366 68 69 70 71 75 </span>
 **##<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
3<span style='mso-spacerun:yes'>
?</span>
1<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
1<span style='mso-spacerun:yes'>
?</span>
5<span style='mso-spacerun:yes'>
? </span>
1<span</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


We can use the two results to get an idea of distribution of results in terms of number of times observations might
be re-sampled when sampling with replacement and the variation in those
results, as shown in Figure 1-17. We could also derive the expected counts for
each number of times of re-sampling when we start with all observations having
an equal chance and sampling with replacement but this isn't important for
using bootstrapping methods. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1029" type="#_x0000_t75" style='width:346. 5pt;height:242. 65pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter1_files/image052. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=462 height=324src="chapter1_files/image052. png" v:shapes="_x0000_i1029">
<![endif]>
</span>
*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 1-17: Counts of number
of times of observation (or not observed for times re-sampled of 0) for two
bootstrap samples. 

</span>
*

*<spanstyle='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>

&nbsp;
</span>
*

The main point of this exploration was to see that each run of the </span>

``resample``  function provides a new version of the data set. Repeating this *B* times using another </span>

``for`` loop, we will track our quantity of interest, say *T*, in all these new "data sets" and call those results *T**. The distribution of the bootstrappedrange of results to expect for the statistic and the middle __% of the *T**?s provides a  ***bootstrap confidence interval<astyle='mso-footnote-id:ftn18' href="#_ftn18" name="_ftnref18" title="">
<![if !supportFootnotes]>
<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
[18]</span>
** </span>
<![endif]>
</span>
</span>
</a>
*** 
for the true parameter ? here the *differencein the two population means*. 

</span>

To make this concrete, we can revisit our previous examples, starting with the </span>

``MockJury2``  data created before and our interest in comparing the mean sentences for the *Average*and *Unattractive* picture groups. The bootstrapping code is very similar to the permutation code except that we apply
the </span>

``resample``  function to the entire data set as opposed to the </span>

``shuffle``  function being applied to the explanatory variable. 

</span>

```{r}



```

 </span>
diffmean</span>
(Years ~</span>
 </span>
Attr, </span>
data=</span>
MockJury2); Tobs</span>

## diffmean </span>
 **## 1. 837127</span>

?</span>
B&lt;-</span>
 </span>
1000</span>
 **?</span>
Tstar&lt;-</span>
matrix</span>
(</span>
NA</span>
, </span>
nrow=</span>
B)</span>
 **?</span>
for (b in (</span>
1</span>
:B)){</span>
 **?? </span>
Tstar[b]&lt;-</span>
diffmean</span>
(Years ~</span>
 </span>
Attr, </span>
data=</span>
resample</span>
(MockJury2))</span>
 **?? </span>
}</span>
 **?</span>
</span>
hist</span>
(Tstar, </span>
labels=</span>
T)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
Tobs, </span>
col=</span>
"red"</span>
, </span>
lwd=</span>
2</span>
)</span>
 **?</span>
</span>
plot</span>
(</span>
density</span>
(Tstar), </span>
main=</span>
"Density curve ofTstar"</span>

)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
Tobs, </span>
col=</span>
"red"</span>
, </span>
lwd=</span>
2</span>
)

</span>

favstats</span>
(Tstar)</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
min<spanstyle='mso-spacerun:yes'>
?????? </span>
Q1<span style='mso-spacerun:yes'>
??</span>
median<span style='mso-spacerun:yes'>
?????? </span>
Q3<spanstyle='mso-spacerun:yes'>
????? </span>
max<span style='mso-spacerun:yes'>
??</span>
mean<span style='mso-spacerun:yes'>
??????? </span>
sd<spanstyle='mso-spacerun:yes'>
??? </span>
n</span>
 **##<span style='mso-spacerun:yes'>
? </span>
-1. 0555561. 266488 1. 839744 2. 429772 4. 420373 1. 8504 0. 8436095 1000</span>
 **##<span style='mso-spacerun:yes'>
? </span>
missing</span>
 **##<span style='mso-spacerun:yes'>
??? </span>
????</span>
0</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1028" type="#_x0000_t75"
 style='width:375pt;height:242. 65pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter1_files/image053. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=500 height=324src="chapter1_files/image053. png" v:shapes="_x0000_i1028">
<![endif]>
</span>
"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>


</span>


*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 1-18: Histogram and
density curve of bootstrap distributions of difference in sample mean Years
with vertical line for the observed difference in the means of 1. 84 years. 

</span>
*

``color:black;background:#E1E2E5'>

&nbsp;
`` 

In this situation, the observed difference in the mean sentences is 1. 84 years (Unattractive-Average), which is
the vertical line in Figure 1-18. The bootstrap distribution shows the results
for the difference in the sample means when fake data sets are re-constructed
by sampling from the data set with replacement. The bootstrap distribution is
approximately centered at the observed value (difference in the sample means)
and is relatively symmetric. 

</span>


The permutation distribution in the same situation (Figure 1-12) had a similar shape but was
centered at 0. Permutations create sampling distributions based on assuming the
null hypothesis is true, which is useful for hypothesis testing. Bootstrapping
creates distributions centered at the observed result, which is the sampling
distribution "under the alternative" or when no null hypothesis is assumed;
bootstrap distributions are useful for generating confidence intervals for the
true parameter values. 

</span>


To create a 95% bootstrap confidence interval for the difference in the true mean sentences (</span>
*$\mu$</span>
*<sub>
Unattr</span>
</sub>
 - </span>
*$\mu$</span>
*<sub>
Ave</span>
</sub>
), select the middle 95% of results from the bootstrap distribution. Specifically, find the 2. 5<sup>
th</sup>

percentile and the 97. 5<sup>
th</sup>
 percentile (values that put 2. 5 and 97. 5%
of the results to the left) in the bootstrap distribution, which leaves 95% in
the middle for the confidence interval. To find percentiles in a distribution
in R, functions are of the form </span>

``q[Name of distribution]`` , with the function </span>
``qt``  extracting percentiles from a *t*-distribution (examples below). From the bootstrap results, use the </span>

``qdata``  function on the </span>
``Tstar``  results that contain the bootstrap distribution of the statistic of interest. 

</span>

```{r}



```

(Tstar, . </span>
025</span>
)</span>

##<spanstyle='mso-spacerun:yes'>
???????? </span>
p<span style='mso-spacerun:yes'>
?</span>
quantile </span>
 **## 0. 0250000 0. 2561804</span>

qdata</span>
(Tstar, . </span>
975</span>
)</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
p quantile </span>
 **</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


These results tell us that the 2. 5<sup>
th</sup>
percentile of the bootstrap distribution is at 0. 26 years and the 97. 5<sup>
th</sup>

percentile is at 3. 50 years. We can combine these results to provide a 95%
confidence for </span>
*$\mu$</span>
*<sub>
Unattr</span>
</sub>
 - </span>
*$\mu$</span>
*<sub>
Ave</span>
</sub>
 that is between 0. 26 and 3. 50. We
can interpret this as with any confidence interval, that we are 95% confident
that the difference in the true mean suggested sentences (Unattractive minus
Average group) is between 0. 26 and 3. 50 years. We can also obtain both
percentiles in one line of code using:

</span>


```{r}



```

qdata</span>
(Tstar, </span>
c</span>
(. </span>
025</span>
, . </span>
975</span>
))</span>
 **?</span>
quantiles</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
quantile<spanstyle='mso-spacerun:yes'>
???? </span>
p</span>
 **## 2. 5%<span style='mso-spacerun:yes'>
?</span>
0. 2561804 0. 025</span>
 **## 97. 5% 3. 4964777 0. 975</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Figure 1-19 displays those same percentiles on the bootstrap distribution residing in </span>

``Tstar`` . 

</span>

```{r}



```

(Tstar, </span>
labels=</span>
T)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
quantiles$quantile, </span>
col=</span>
"blue"</span>
, </span>
lwd=</span>
3</span>
)</span>
 **?</span>
</span>
plot</span>
(</span>
density</span>
(Tstar), </span>
main=</span>
"Density curve ofTstar"</span>

)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
quantiles$quantile, </span>
col=</span>
"blue"</span>
, </span>
lwd=</span>
3</span>
)</span>

</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1027" type="#_x0000_t75" style='width:301. 15pt;height:210. 75pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter1_files/image054. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=402 height=281src="chapter1_files/image055. jpg" v:shapes="_x0000_i1027">
<![endif]>
</span>
*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 1-19: Histogram and
density curve of bootstrap distribution with 95% bootstrap confidence intervals
displayed (vertical lines). 

</span>
*

*<spanstyle='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>

&nbsp;
</span>
*

Although confidence intervals can exist without referencing hypotheses, we can revisit our previous
$H_0$:
</span>
*$\mu$</span>
*<sub>
Unattr</span>
</sub>
 = </span>
*$\mu$</span>
*<sub>
Ave</span>
</sub>
. This null hypothesis is equivalent
$H_0$: </span>
*$\mu$</span>
*<sub>
Unattr</span>
</sub>
 - </span>
*$\mu$</span>
*<sub>
Ave</span>
</sub>
=0, that the difference in the true
means is equal to 0 years. And the difference in the means was the scale for
our confidence interval, which did not contain 0 years. We will call 0 an interesting
 ***referencevalue***  for the confidence interval, because here it is the value where the true means are equal to each other (have a difference of 0 years). In
general, if our confidence interval does not contain 0, then it is saying that
0 is not one of our likely values for the difference in the true means. This
implies that we should reject a claim that they are equal. This provides the
same inferences for the hypotheses that we considered previously using both a
parametric and permutation approach. The general summary is that we can use
confidence intervals to test hypotheses by assessing whether the reference
$H_0$)
$H_0$). P-values are more
informative about hypotheses but confidence intervals are more information
about the size of differences, so both offer useful information and, as shown
here, can provide consistent conclusions about hypotheses. 

</span>


As in the previous situation, we also want to consider the parametric approach for comparison
purposes and to have that method available, especially to help us understand
some methods where we will only consider parametric inferences in later
chapters. The parametric confidence interval is called the  ***equal variance, ***   ***two-samplet confidence interval***  and assumes that the populations being sampledfrom are normally distributed and leads to using a *t*-distribution to form the interval. The output from the </span>
``t. test``  function provides the parametric 95%confidence interval calculated for you:

</span>


```{r}



```

(Years ~</span>
 </span>
Attr, </span>
data=</span>
MockJury2, </span>
var. equal=</span>
T)</span>

##<spanstyle='mso-spacerun:yes'>
? </span>
Two Sample t-test</span>
 **## data:<span style='mso-spacerun:yes'>
? </span>
Years by Attr</span>
 **## t = -2. 1702, df = 73, p-value = 0. 03324</span>
 **## alternative hypothesis: true difference in means is not equal to 0</span>
 ** **** </span>
 ** **##<span style='mso-spacerun:yes'>
? </span>
-3. 5242237-0. 1500295</span>
**  **## sample estimates:</span>
 **##<span style='mso-spacerun:yes'>
????? </span>
mean in group Average mean in group Unattractive </span>
 **##<span style='mso-spacerun:yes'>
??????????????????</span>
3. 973684<span style='mso-spacerun:yes'>
??????????????????</span>
5. 810811</span>


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The </span>
``t. test`` the groups and provides slightly different end-points than our bootstrap confidence interval (both are made at the 95% confidence level though), which
was slightly narrower. Both intervals have the same interpretation, only the
methods for calculating the intervals and the assumptions differ. Specifically, 
the bootstrap interval can tolerate different distribution shapes other than
normal and still provide intervals that work well<a style='mso-footnote-id:
ftn19' href="#_ftn19" name="_ftnref19" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[19]</span>
</span>
<![endif]>
</span>
</span>
</a>
. 
The other assumptions are all the same as for the hypothesis test, where we
continue to assume that we have independent observations with equal variances
for the two groups. 

</span>


The formula that </span>
``t. test``  is using to calculate the parametric ***equal-variancetwo-sample t confidence interval***  is:

</span>

<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
?</span>
</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;  line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;  line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
t</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;  line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
df</m:r>
<m:r>
 </m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;  line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
*</m:r>
</span>
*</m:sup>
</m:sSubSup>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;  line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
p</m:r>
</span>
*</m:sub>
</m:sSub>
<m:rad>
<m:radPr>
<m:degHide  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:fPr>
<m:num>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:num>
<m:den>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
</m:den>
</m:f>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:fPr>
<m:num>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:num>
<m:den>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
 type="#_x0000_t75" style='width:130. 9pt;height:43. 5pt'>

 <v:imagedata src="chapter1_files/image056. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=175 height=58src="chapter1_files/image057. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>


</span>


In this situation, the *df* is again *n<sub>
1$+n<sub>
2</sub>
-2* and </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
p</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;     line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;     line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSubSup>
*<span style='font-size:12. 0pt;    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;     line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;     line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
<m:r>
2</m:r>
</span>
*</m:den>
</m:f>
</m:e>
</m:rad>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-10. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:115. 9pt;height:30. 4pt'>

 <v:imagedata src="chapter1_files/image042. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=155 height=41src="chapter1_files/image043. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. </span>

 The </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSubSup>
<m:sSubSupPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
t</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
df</m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
*</m:r>
</span>
*</m:sup>
</m:sSubSup>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:15. 75pt;height:16. 9pt'>

 <v:imagedata src="chapter1_files/image058. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=21 height=23src="chapter1_files/image059. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
a multiplier that comes from finding the percentile from the *t*-distribution that puts *C*% in the middle of the distribution with *C* being the confidence level. It is important to note that this *t** hasnothing to do with the previous test statistic *t*. It is confusing and many of you will, at some point, happily take the result from a test statistic calculation and use it for a multiplier
in a *t*-based confidence interval. Figure 1-20 shows the *t*-distribution with 73 degrees of freedom and the cut-offs that put 95% of the area in the
middle. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1026" type="#_x0000_t75" style='width:278. 65pt;height:195pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter1_files/image060. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=372 height=260src="chapter1_files/image061. jpg" v:shapes="_x0000_i1026">
<![endif]>
</span>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
Figure
1-20: Plot of t(73) with cut-offs for putting 95% of distributions in the
middle. 

</span>
*

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
For 95% confidence intervals, the
multiplier is going to be close to 2 and anything else is a sign of a mistake. 
We can use R to get the multipliers for confidence intervals using the </span>

``qt`` <spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 function in a similar fashion to how </span>
``qdata`` <spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 was used in the bootstrap results, except that this new value must be used in the previous confidence interval
formula. This function produces values for requested percentiles, so if we want
to put 95% in the middle, we place 2. 5% in each tail of the distribution and
need to request the 97. 5<sup>
th</sup>
 percentile. Because the t-distribution is
always symmetric around 0, we merely need to look up the value for the 97. 5<sup>
th</sup>

percentile and know that the multiplier for the 2. 5<sup>
th</sup>
 percentile is
just *?t<sup>
*</sup>
*. The *t** multiplier to form the confidenceinterval is 1. 993 for a 95% confidence interval when the *df*=73 based on the results from </span>
line-height:107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
qt</span>

line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
:

</span>


```{r}



```

(. </span>
975</span>
, </span>
df=</span>
73</span>
## [1]1. 992997</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<spanstyle='font-size:12. 0pt;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
minor-fareast'>
Note that the 2. 5<sup>
th</sup>
 percentile is just the negative
of this value due to symmetry and the real source of the minus in the
minus/plus in the formula for the confidence interval. </span>

"Times New Roman";mso-bidi-font-family:"Courier New";color:blue;background:#E1E2E5'>


</span>


```{r}



```

(. </span>
025</span>
, </span>
df=</span>
73</span>
## [1]-1. 992997</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


We can also re-write the confidence interval formula into a slightly more general form as

</span>


<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
t</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
df</m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
*</m:r>
</span>
*</m:sup>
</m:sSubSup>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SE</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:106. 5pt;height:16. 9pt'>

 <v:imagedata src="chapter1_files/image062. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=142 height=23src="chapter1_files/image063. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
OR<span style='mso-spacerun:yes'>
? </span>
</span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:69. 75pt;height:15pt'>

 <v:imagedata src="chapter1_files/image064. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=93 height=20src="chapter1_files/image065. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";


</span>


where </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SE</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
p</m:r>
</span>
*</m:sub>
</m:sSub>
<m:rad>
<m:radPr>
<m:degHide  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:fPr>
<m:num>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:num>
<m:den>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
</m:den>
</m:f>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:f>
<m:fPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:fPr>
<m:num>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:num>
<m:den>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
</m:den>
</m:f>
</m:e>
</m:rad>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-12. 0pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:108. 75pt;height:30. 4pt'>

 <v:imagedata src="chapter1_files/image066. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=145 height=41src="chapter1_files/image067. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSubSup>
<m:sSubSupPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
ME</m:r>
<m:r>
=</m:r>
<m:r>
t</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
df</m:r>
</span>
*</m:sub>
<m:sup>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
*</m:r>
</span>
*</m:sup>
</m:sSubSup>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SE</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
    mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:    "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
    mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:88. 5pt;height:16. 9pt'>

 <v:imagedata src="chapter1_files/image068. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=118 height=23src="chapter1_files/image069. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. In some situations, researchers will
report the  ***standard error***  (SE) or  ***margin of error***  (ME) as a method of
quantifying the uncertainty in a statistic. The SE is an estimate of the
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:37. 5pt;height:15pt'>

 <v:imagedata src="chapter1_files/image070. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=50 height=20src="chapter1_files/image071. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
) and the ME is an estimate of the
precision of a statistic that can be used to directly form a confidence
interval. The ME depends on the choice of confidence level although 95% is
almost always selected. </span>

To finish this example, R can be used to help you do calculations much like a calculator except with much
more power "under the hood". You have to make sure you are careful with using </span>

``( )``  to group items and remember that the asterisk (*) is used for multiplication in R. We need the pertinent information
which is available from the bolded parts of the </span>

``favstats``  output repeated below to calculate the confidence interval "by hand" using R. 

</span>

```{r}



```

(Years~Attr, </span>
data=</span>
MockJury2)</span>

##<spanstyle='mso-spacerun:yes'>
?????????? </span>
Attr min Q1 median Q3 max<spanstyle='mso-spacerun:yes'>
???? </span>
mean<span style='mso-spacerun:yes'>
??????</span>
sd<span style='mso-spacerun:yes'>
? </span>
n missing</span>
 **## 1<span style='mso-spacerun:yes'>
????? </span>
 **** ?? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
????? </span>
3<spanstyle='mso-spacerun:yes'>
? </span>
5<span style='mso-spacerun:yes'>
? </span>
12  ****   **** ?????? </span>
0</span>
 **## 2  **** ?? </span>
1<span style='mso-spacerun:yes'>
?</span>
2<span style='mso-spacerun:yes'>
????? </span>
5 10<spanstyle='mso-spacerun:yes'>
? </span>
15  ****  **** ?????? </span>
0</span>


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Start with typing the following  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
s</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
p</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:10. 9pt;height:16. 5pt'>

 <v:imagedata src="chapter1_files/image072. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=15 height=22src="chapter1_files/image073. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
and store it in a variable named </span>
``sp`` :

</span>

```{r}



```

 </span>
sqrt</span>
(((</span>
38-1</span>
)*(</span>
2. 8235</span>
^</span>
2</span>
)+(</span>
37-1</span>
)*(</span>
4. 364</span>
^</span>
2</span>
))/(</span>
38+37-2</span>
))</span>
 **## [1]3. 665036</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Then calculate the confidence interval that </span>

``t. test`` provided using:

</span>

```{r}



```

+</span>
c</span>
(-</span>
1</span>
, </span>
1</span>
)*</span>
qt</span>
(. </span>
975</span>
, </span>
df=</span>
73</span>
)*sp*</span>
sqrt</span>
(</span>
1</span>
/</span>
38+1</span>
/</span>
37</span>
)</span>

## [1]-3. 5240302 -0. 1499698</span>

</div>


The previous code uses </span>
``c(-1, 1)``  times the margin of error to subtract and add the ME to the difference in the sample means (</span>

``3. 974-5. 811`` ), which generates the lower and then upper bounds of the confidence interval. If desired, we can also use just the
last portion of the previous calculation to find the margin of error, which is
1. 69 here. 

</span>


```{r}



```

(. </span>
975</span>
, </span>
df=</span>
73</span>
)*sp*</span>
sqrt</span>
(</span>
1</span>
/</span>
38+1</span>
/</span>
37</span>
##[1] 1. 68703</span>


</div>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo7'>
<a
name="_Toc459020881">
1. 8: Bootstrap confidence interval for difference in GPAs</a>
</h2>


????????????? </span>
We can now apply the new confidence interval methods on the STAT 217 grade data. 
This time we start with the parametric 95% confidence interval "by hand" in R
and then use </span>

``t. test`` to verify our result. The </span>
``favstats``  output provides us with the required information to calculate the confidence interval:

</span>


```{r}



```

(GPA~Sex, </span>
data=</span>
s217)</span>

##<spanstyle='mso-spacerun:yes'>
?? </span>
Sex<span style='mso-spacerun:yes'>
?</span>
min<span style='mso-spacerun:yes'>
? </span>
Q1 median<spanstyle='mso-spacerun:yes'>
?? </span>
Q3 max<span style='mso-spacerun:yes'>
????</span>
mean<span style='mso-spacerun:yes'>
??????? </span>
sd<spanstyle='mso-spacerun:yes'>
? </span>
n missing</span>
 **## 1<span style='mso-spacerun:yes'>
?? </span>
F 2. 503. 1<span style='mso-spacerun:yes'>
? </span>
3. 400 3. 70<spanstyle='mso-spacerun:yes'>
?? </span>
4 3. 338378 0. 4074549 37<spanstyle='mso-spacerun:yes'>
?????? </span>
0</span>
 **## 2<span style='mso-spacerun:yes'>
?? </span>
M 1. 962. 8<span style='mso-spacerun:yes'>
? </span>
3. 175 3. 46<spanstyle='mso-spacerun:yes'>
?? </span>
4 3. 088571 0. 4151789 42<spanstyle='mso-spacerun:yes'>
?????? </span>
0</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The *df* are 37+42-2 = 77. Using the SDs from the two groups and their
```{r}



```

 </span>
sqrt</span>
(((</span>
37-1</span>
)*(</span>
0. 4075</span>
^</span>
2</span>
)+(</span>
42-1</span>
)*(</span>
0. 41518</span>
^</span>
2</span>
))/(</span>
37+42-2</span>
))</span>
 **## [1]0. 4116072</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The margin of error is:

</span>

```{r}



```

(. </span>
975</span>
, </span>
df=</span>
77</span>
)*sp*</span>
sqrt</span>
(</span>
1</span>
/</span>
37+1</span>
/</span>
42</span>
## [1]0. 1847982</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


All together, the 95% confidence interval is:

</span>


```{r}



```

+</span>
c</span>
(-</span>
1</span>
, </span>
1</span>
)*</span>
qt</span>
(. </span>
975</span>
, </span>
df=</span>
77</span>
)*sp*</span>
sqrt</span>
(</span>
1</span>
/</span>
37+1</span>
/</span>
42</span>
## [1]0. 0646018 0. 4341982</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


So we are 95% confident that the difference in the true mean GPAs between females and males (females minus males)
is between 0. 065 and 0. 434 GPA points. We get a similar<a style='mso-footnote-id:
ftn20' href="#_ftn20" name="_ftnref20" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[20]</span>
</span>
<![endif]>
</span>
</span>
</a>

result from the bolded part of the </span>

``t. test`` ```{r}



```

(GPA~Sex, </span>
data=</span>
s217, </span>
var. equal=</span>
T)</span>

## </span>
 **##<span style='mso-spacerun:yes'>
? </span>
Two Samplet-test</span>
 **## </span>
 **## data:<span style='mso-spacerun:yes'>
? </span>
GPA bySex</span>
 **## t = 2. 6919, df = 77, p-value = 0. 008713</span>
 **## alternative hypothesis: true difference in means isnot equal to 0</span>
 **##  **** </span>
 ** **** ##<span style='mso-spacerun:yes'>
? </span>
 **** </span>
 **## sample estimates:</span>
 **## mean in group F mean in group M </span>
 **##<span style='mso-spacerun:yes'>

&nbsp;


</div>


Note that we can easily switch to 90%or 99% confidence intervals by simply changing the percentile in </span>

``qt``  or changing </span>
``conf. level``  in the </span>
``t. test``  function. In the following two linesof code, we added octothorpes<a style='mso-footnote-id:ftn21' href="#_ftn21"
name="_ftnref21" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[21]</span>
</span>
<![endif]>
</span>
</span>
</a>

(</span>

``#`` ) and then some text after functioncalls to explain what is being calculated. In computer code, octothorpes
provide a way of adding comments that tell the software (here R) to ignore any
text after a "</span>

``#`` a given line. In the color version of the text, comments are also clearlydistinguished. 

</span>


```{r}



```

</span>
qt</span>
(. </span>
95</span>
, </span>
df=</span>
77</span>
) </span>
# For 90% confidence and 77 df</span>

## [1]1. 664885

</span>


&nbsp;


? </span>
</span>
qt</span>
(. </span>
995</span>
, </span>
df=</span>
77</span>
) </span>
#For 99% confidence and 77 df</span>

## [1]2. 641198

</span>

t. test</span>
(GPA~Sex, </span>
data=</span>
s217, </span>
var. equal=</span>
T, </span>
conf. level=</span>
0. 90</span>
##<spanstyle='mso-spacerun:yes'>
? </span>
Two Sample t-test</span>
 **## data:<span style='mso-spacerun:yes'>
? </span>
GPA bySex</span>
 **## t = 2. 6919, df = 77, p-value = 0. 008713</span>
 **## alternative hypothesis: true difference in means isnot equal to 0</span>
 **##  **** </span>
 **##<span style='mso-spacerun:yes'>
? </span>
 **0. 09530553 0. 40430837

** </span>


 **t. test</span>
(GPA~Sex, </span>
data=</span>
s217, </span>
var. equal=</span>
T, </span>
conf. level=</span>
0. 99</span>
)</span>

## t =2. 6919, df = 77, p-value = 0. 008713</span>
 **##  **** </span>
 **##<span style='mso-spacerun:yes'>
? </span>
 **** </span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


As a review of some basicideas with confidence intervals make sure you can answer the following
questions:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-bidi-font-family:Calibri;
1)<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
What is the impact of increasing the
confidence level in this situation? 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-bidi-font-family:Calibri;
2)<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
What happens to the width of the
confidence interval if the size of the SE increases or decreases? 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-bidi-font-family:Calibri;
3)<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
What about increasing the sample size
? should that increase or decrease the width of the interval? 

</span>


All the general results you learnedbefore about impacts to widths of CIs hold in this situation whether we are
considering the parametric or bootstrap methods?

</span>


To finish this example, we will generate the comparable bootstrap 90% confidence interval using the
bootstrap distribution in Figure 1-21. 

</span>


```{r}



```

 </span>
diffmean</span>
(GPA ~</span>
 </span>
Sex, </span>
data=</span>
s217); Tobs</span>

##<spanstyle='mso-spacerun:yes'>
?? </span>
diffmean </span>
 **?</span>
</span>
par</span>
(</span>
mfrow=</span>
c</span>
(</span>
1</span>
, </span>
2</span>
))</span>
 **?</span>
B&lt;-</span>
 </span>
1000</span>
 **?</span>
Tstar&lt;-</span>
matrix</span>
(</span>
NA</span>
, </span>
nrow=</span>
B)</span>
 **?</span>
for (b in (</span>
1</span>
:B)){</span>
 **?? </span>
Tstar[b]&lt;-</span>
diffmean</span>
(GPA ~</span>
 </span>
Sex, </span>
data=</span>
resample</span>
(s217))</span>
 **?? </span>
}</span>
 **?</span>
</span>
 **?</span>
</span>
qdata</span>
(Tstar, . </span>
05</span>
##<spanstyle='mso-spacerun:yes'>
????????? </span>
p<span style='mso-spacerun:yes'>
??</span>
quantile </span>
 **##<span style='mso-spacerun:yes'>
? </span>
0. 0500000 -0. 3932474</span>

?</span>
</span>
qdata</span>
(Tstar, . </span>
95</span>
)</span>

##<spanstyle='mso-spacerun:yes'>
?????????? </span>
p<span style='mso-spacerun:yes'>
???</span>
quantile </span>
 **##<span style='mso-spacerun:yes'>
? </span>
0. 95000000-0. 09363069</span>


?</span>
quantiles&lt;-</span>
qdata</span>
(Tstar, </span>
c</span>
(. </span>
05</span>
, . </span>
95</span>
))</span>
 **##<spanstyle='mso-spacerun:yes'>
??????? </span>
quantile<spanstyle='mso-spacerun:yes'>
??? </span>
p</span>
 **## 5%<span style='mso-spacerun:yes'>
?</span>
-0. 39324743 0. 05</span>
 **## 95% -0. 09363069 0. 95</span>

</div>


normal;tab-stops:45. 8pt 91. 6pt 137. 4pt 183. 2pt 229. 0pt 274. 8pt 320. 6pt 366. 4pt 412. 2pt 458. 0pt 503. 8pt 549. 6pt 595. 4pt 641. 2pt 687. 0pt 732. 8pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>

&nbsp;
</span>


The output tells us that the 90%confidence interval is from -0. 393 to -0. 094 GPA points. The bootstrap
distribution with the observed difference in the sample means and these
cut-offs is displayed in Figure 1-21 using this code:

</span>


```{r}



```

(Tstar, </span>
labels=</span>
T)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
Tobs, </span>
col=</span>
"red"</span>
, </span>
lwd=</span>
2</span>
)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
quantiles$quantile, </span>
col=</span>
"blue"</span>
, </span>
lwd=</span>
3</span>
, </span>
lty=</span>
2</span>
)</span>
 **?</span>
</span>
plot</span>
(</span>
density</span>
(Tstar), </span>
main=</span>
"Density curve ofTstar"</span>

)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
Tobs, </span>
col=</span>
"red"</span>
, </span>
lwd=</span>
2</span>
)</span>
 **?</span>
</span>
abline</span>
(</span>
v=</span>
quantiles$quantile, </span>
col=</span>
"blue"</span>
, </span>
lwd=</span>
3</span>
, </span>
lty=</span>
2</span>
)</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75" style='width:327. 75pt;
 height:229. 9pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter1_files/image074. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=437 height=307src="chapter1_files/image075. jpg" v:shapes="Picture">
<![endif]>
</span>
*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 1-21: Histogram and
density curve of bootstrap distribution of difference in sample mean GPAs (male
minus female) with observed difference (solid vertical line) and quantiles that
delineate the 90% confidence intervals (dashed vertical lines). 

</span>
*

*<spanstyle='font-size:8. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>

&nbsp;
</span>
*

In the previous output, theparametric 90% confidence interval is from 0. 095 to 0. 404, suggesting similar
results again from the two approaches once you account for the two different
orders of differencing of the groups. Based on the bootstrap CI, we can say
that we are 90% confident that the difference in the true mean GPAs for STAT
217 students is between -0. 393 to -0. 094 GPA points (male minus females). 
Because sex cannot be assigned to the subjects, we cannot infer that sex is
causing this difference and because this was a voluntary response sample of
STAT 217 students in a given semester, we cannot infer that a difference of
this size would apply to all STAT 217 students or even students in another
semester. 

</span>


Throughout the semester, pay attention to the distinctions between parameters and statistics, focusing
on the differences between estimates based on the sample and inferences for the
population of interest in the form of the parameters of interest. Remember that
statistics are summaries of the sample information and parameters are
characteristics of populations (which we rarely know). And that our inferences
are limited to the population that we randomly sampled from, if we randomly
sampled. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo7'>
<a
name="_Toc459020882">
1. 9: Chapter summary</a>
</h2>


In this chapter, wereviewed basic statistical inference methods in the context of a two-sample
mean problem. You were introduced to using R to do permutation testing and
generate bootstrap confidence intervals as well as obtaining parametric *t*-test and confidence intervals in thissame situation. You should have learned how to use a </span>

``for``  loop for doing the nonparametricinferences and the </span>

``t. test``  function for generating parametric inferences. In the twoexamples considered, the parametric and nonparametric methods provided similar
results, suggesting that the assumptions were at least close to being met for
the parametric procedures. When parametric and nonparametric approaches
disagree, the nonparametric methods are likely to be more trustworthy since
they have less restrictive assumptions but can still have problems. 

</span>


When the noted conditionsare not met in a hypothesis testing situation, the Type I error rates can be
inflated, meaning that we reject the null hypothesis more often than we have
allowed to occur by chance. Specifically, we could have a situation where our
assumed 5% significance level test might actually reject the null when it is
true 20% of the time. If this is occurring, we call a procedure  ***liberal*** (it rejects too easily) and if the procedure is liberal, how could we trust a
small p-value to be a "real" result and not just an artifact of violating the
assumptions of the procedure? Likewise, for confidence intervals we hope that
our 95% confidence level procedure, when repeated, will contain the true
parameter 95% of the time. If our assumptions are violated, we might actually
have an 80% confidence level procedure and it makes it hard to trust the
reported results for our observed data set. Statistical inference relies on a
belief in the methods underlying our inferences. If we don?t trust our
assumptions, we shouldn?t trust the conclusions to perform the way we want them
to. As sample sizes increase and/or violations of conditions lessen, then the
procedures will perform better. In Chapter 2, we?ll learn some new tools for
doing diagnostics to help us assess how much those conditions are violated. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo7'>
<a
name="_Toc459020883">
1. 10: Summary of important R code</a>
</h2>


The main components of R code used inthis chapter follow with components to modify in red, remembering that any R
packages mentioned need to be installed and loaded for this code to have a
chance of working:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``summary(<spanstyle='color:red'>
DATASETNAME`` )

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesnumerical summaries of all variables in the data set. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``t. test(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
X</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
, conf. level=0. 95)

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providestwo-sample t-test test statistic, df, p-value, and 95% confidence interval. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``2*pt(abs(<spanstyle='color:red'>
Tobs`` ), df=<span style='color:red'>
DF</span>
, lower. tail=F)

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Findsthe two-sided test p-value for an observed 2-sample t-test statistic of </span>

``Tobs`` . 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``hist(<spanstyle='color:red'>
DATASETNAME$Y`` )

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Makesa histogram of a variable named </span>

``Y``  from the data set of interest. 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``boxplot(<spanstyle='color:red'>
Y~X`` , data=<span style='color:red'>
DATASETNAME</span>
)

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Makesa boxplot of a variable named Y for groups in X from the data set. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``beanplot(<spanstyle='color:red'>
Y~X`` , data=<span style='color:red'>
DATASETNAME</span>
)

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Makesa beanplot of a variable named Y for groups in X from the data set. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe beanplot package is loaded. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``mean(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
X</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
); sd(<span style='color:red'>
Y</span>
~<spanstyle='color:red'>
X</span>
, data=<span style='color:red'>
DATASETNAME</span>
) 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesthe mean and sd of responses of Y for each group described in X. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``favstats(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
X</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
) 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesnumerical summaries of Y by groups described in X. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``Tobs &lt;- t. test(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
X</span>
, data=<spanstyle='color:red'>
DATASETNAME</span>
, var. equal=T)$statistic; Tobs

</span>


``B&lt;-1000

`` 

``Tstar&lt;-matrix(NA, nrow=B)

`` 

``for (b in (1:B)){

`` 

``? `` Tstar[b]&lt;-t. test(<spanstyle='color:red'>
Y</span>
~shuffle(<span style='color:red'>
X</span>
), data=<spanstyle='color:red'>
DATASETNAME</span>
, var. equal=T)$statistic

</span>


``? `` }</span>


</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Codeto run a </span>

``for`` loop to generate 1000 permuted versions of the test statistic using the </span>
``shuffle``  function and keep track of theresults in </span>

``Tstar`` <![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``pdata(Tstar, 
margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;`` </span>
</span>
<![endif]>
Findsthe proportion of the permuted test statistics in Tstar that are less than
-|Tobs| or greater than |Tobs|, useful for finding the two-sided test p-value. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;font-family:Symbol;mso-fareast-font-family:Symbol;
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``Tobs &lt;- diffmean(<spanstyle='color:red'>
Y`` ~<span style='color:red'>
X</span>
, data=<spanstyle='color:red'>
 DATASETNAME</span>
); Tobs

</span>


``B&lt;-1000

`` 

``Tstar&lt;-matrix(NA, nrow=B)

`` 

``for (b in (1:B)){

`` 

``? `` Tstar[b]&lt;-diffmean(<spanstyle='color:red'>
Y</span>
~<span style='color:red'>
X</span>
, data=resample(<spanstyle='color:red'>
DATASETNAME</span>
))

</span>


``? `` }</span>


</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Codeto run a </span>

``for`` loop to generate 1000 bootstrapped versions of the data set using the </span>
``resample``  function and keep track of theresults of the statistic in </span>

``Tstar`` . 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``qdata(Tstar, c(0. 025, 0. 975))

`` 

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Providesthe values that delineate the middle 95% of the results in the bootstrap
distribution (</span>

``Tstar`` <h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo7'>
<a
name="_Toc459020884">
1. 11: Practice problems</a>
</h2>


Load the </span>
``HELPrct``  data set from the </span>
``mosaicData``  package (you need to install the </span>
``mosaicData``  package once to be able to load it). The HELP study was a clinical trial for adult inpatients recruited from a
detoxification unit. Patients with no primary care physician were randomly
assigned to receive a multidisciplinary assessment and a brief motivational
intervention or usual care and various outcomes were observed. Two of the
variables in the dataset are </span>

``sex`` , a factor with levels male and female and </span>
``daysanysub``  which is the time (in days) to firstuse of any substance post-detox. We are interested in the difference in mean
number of days to first use of any substance post-detox between males and
females. There are some missing responses and the following code will produce </span>

``favstats``  with the missing values and thenprovide a data set that for complete observations by applying the </span>

``na. omit``  function that removes anyobservations with missing values. 

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
require(mosaicData) #load the dataset

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
data(HELPrct)

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
HELPrct2&lt;-HELPrct[, c("daysanysub", "sex")]
#Just focus on two variables

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
HELPrct3&lt;-na. omit(HELPrct2) #Removes subjects with missing

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
favstats(daysanysub~sex, data = HELPrct2)

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
favstats(daysanysub~sex, data = HELPrct3)

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>

&nbsp;
</span>


1. 1. Based on the results provided, 
how many observations were missing for males and females? Missing values here
likely mean that the subjects didn?t use any substances post-detox in the time
of the study but might have at a later date ? the study just didn?t run long
enough. This is called  ***censoring*** . What is the problem withthe numerical summaries here if the missing responses were all something larger
than the largest observation?

</span>


1. 2. Make a beanplot and a boxplot of
</span>

``daysanysub``  ~ </span>
``sex``  using the </span>
``HELPrct3``  data set created above. Compare thedistributions, recommending parametric or nonparametric inferences. 

</span>

1. 3. Generate the permutation results
and write out the 6+ steps of the hypothesis test, making sure to note the
numerical value of observed test statistic you are using and include a
discussion of the scope of inference. 

</span>


1. 4. Interpret the p-value for these
results. 

</span>


1. 5. Generate the parametric </span>

``t. test``  results, reporting the test-statistic, its distribution under the null hypothesis, and compare the p-value to those
observed using the permutation approach. 

</span>


1. 6. Make and interpret a 95%
bootstrap confidence interval for the difference in the means. 

</span>



&nbsp;
</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>

">
http://fivethirtyeight. com/</a>
. </p>


</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>
 </p>


</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>
 If
you?ve taken calculus, you will know that the curve is being constructed so
that the integral from -&#8734; to &#8734; is 1. If you don?t know calculus, 
think of a rectangle with area of 1 based on its height and width. These cover
the same area but the top of the region wiggles. </p>


</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>

Jittering typically involves adding random variability to each observation that
is uniformly distributed in a range determined based on the spacing of the
function, the results will change. For more details, type ``help(jitter)``  in R. </p>


</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>
 <spanstyle='mso-bidi-font-size:12. 0pt'>
Remember the bell-shaped curve youencountered in introductory statistics? If not, you can see some at </span>
<a
href="https://en. wikipedia. org/wiki/Normal_distribution">
https://en. wikipedia. org/wiki/Normal_distribution</span>
</a>


</span>


</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>
 Well, 
you can use other colors (try "lightblue" for example), but I think bisque
looks nice in these plots. </p>


</div>


<div style='mso-element:footnote' id=ftn7>


<a style='mso-footnote-id:ftn7' href="#_ftnref7"name="_ftn7" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[7]</span>
</span>
<![endif]>
</span>
</span>
</a>

The hypothesis of no difference that is typically generated in the hopes of
being rejected in favor of the alternative hypothesis which contains the sort
of difference that is of interest in the application. </p>


</div>


<div style='mso-element:footnote' id=ftn8>


<a style='mso-footnote-id:ftn8' href="#_ftnref8"name="_ftn8" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[8]</span>
</span>
<![endif]>
</span>
</span>
</a>

The null model is the statistical model that is implied by the chosen null
hypothesis. Here, a null hypothesis of no difference translates to having a
model with the same mean for both groups. </p>


</div>


<div style='mso-element:footnote' id=ftn9>


<a style='mso-footnote-id:ftn9' href="#_ftnref9"name="_ftn9" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[9]</span>
</span>
<![endif]>
</span>
</span>
</a>
 We?ll
more common usage below; while the code to generate ``Perm1``  is provided, it isn?t something to worry about
right now. </p>


</div>


<div style='mso-element:footnote' id=ftn10>


<a style='mso-footnote-id:ftn10' href="#_ftnref10"name="_ftn10" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[10]</span>
</span>
<![endif]>
</span>
</span>
</a>

P-values are the probability of obtaining a result as extreme as or more
extreme than we observed given that the null hypothesis is true. </p>


</div>


<div style='mso-element:footnote' id=ftn11>


<a style='mso-footnote-id:ftn11' href="#_ftnref11"name="_ftn11" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[11]</span>
</span>
<![endif]>
</span>
</span>
</a>

We often say "under" in statistics and we mean "given that the following is
true". </p>


</div>


<div style='mso-element:footnote' id=ftn12>


<a style='mso-footnote-id:ftn12' href="#_ftnref12"name="_ftn12" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[12]</span>
</span>
<![endif]>
</span>
</span>
</a>

This is a fancy way of saying "in advance", here in advance of seeing the
observations. </p>


</div>


<div style='mso-element:footnote' id=ftn13>


<a style='mso-footnote-id:ftn13' href="#_ftnref13"name="_ftn13" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[13]</span>
</span>
<![endif]>
</span>
</span>
</a>

Statistically, a conservative method is one that provides less chance of
rejecting the null hypothesis in comparison to some other method or less than
some pre-defined standard. </p>


</div>


<div style='mso-element:footnote' id=ftn14>


<a style='mso-footnote-id:ftn14' href="#_ftnref14"name="_ftn14" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[14]</span>
</span>
<![endif]>
</span>
</span>
</a>

We?ll leave the discussion of the CLT to your previous stat coursework or an
internet search. Remember that it has something to do with distributions
looking more normal as the sample size increases. </p>


</div>


<div style='mso-element:footnote' id=ftn15>


<a style='mso-footnote-id:ftn15' href="#_ftnref15"name="_ftn15" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[15]</span>
</span>
<![endif]>
</span>
</span>
</a>
 <spanstyle='font-size:11. 0pt;mso-bidi-font-size:12. 0pt'>
On exams, you will be askedto describe the area of interest, sketch a picture of the area of interest, 
and/or note the distribution you would use. </span>


</div>


<div style='mso-element:footnote' id=ftn16>


<a style='mso-footnote-id:ftn16' href="#_ftnref16"name="_ftn16" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[16]</span>
</span>
<![endif]>
</span>
</span>
</a>

In some studies, the same subject might be measured in both conditions and this
violates the assumptions of this procedure. </p>


</div>


<div style='mso-element:footnote' id=ftn17>


<a style='mso-footnote-id:ftn17' href="#_ftnref17"name="_ftn17" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[17]</span>
</span>
<![endif]>
</span>
</span>
</a>

also used here. It really isn?t important but makes sure the output of ``table``  is sorted by observation
number by first converting the *orig. id*variable into a numeric vector. </p>

</div>


<div style='mso-element:footnote' id=ftn18>


<a style='mso-footnote-id:ftn18' href="#_ftnref18"name="_ftn18" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[18]</span>
</span>
<![endif]>
</span>
</span>
</a>

There are actually many ways to use this information to make a confidence
interval. We are using the simplest method called that is called the
"percentile" method. </p>


</div>


<div style='mso-element:footnote' id=ftn19>


<a style='mso-footnote-id:ftn19' href="#_ftnref19"name="_ftn19" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[19]</span>
</span>
<![endif]>
</span>
</span>
</a>

When hypothesis tests "work well" they have high power to detect differences
while having Type I error rates that are close to what we chose a priori. When
confidence intervals "work well", they contain the true parameter value in
repeated random samples at around the selected confidence level. </p>


</div>


<div style='mso-element:footnote' id=ftn20>


<a style='mso-footnote-id:ftn20' href="#_ftnref20"name="_ftn20" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[20]</span>
</span>
<![endif]>
</span>
</span>
</a>

We rounded the means a little and that caused the small difference in results. </p>


</div>


<div style='mso-element:footnote' id=ftn21>


<a style='mso-footnote-id:ftn21' href="#_ftnref21"name="_ftn21" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[21]</span>
</span>
<![endif]>
</span>
</span>
</a>

You can correctly call octothorpes *number*symbols or, in the twitter verse, *hashtags*. For more on this symbol, see <a href="http://blog. dictionary. com/octothorpe/">
http://blog. dictionary. com/octothorpe/</a>
. 
I usually call them number symbols too. </p>


</div>


</div>


** 

</html>
