---
output:
  html_document: default
  pdf_document: default
---

#One-Way ANOVA




</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020892">
2. 6: Pair-wise comparisons for Prisoner Rating data</a>


</h2>


????????????? </span>
Inour previous work with the prisoner rating data, the overall ANOVA test
provided only marginal evidence of some difference in the true means across the
three groups with a p-value=0. 067. Tukey?s HSD does not require you to find a small
employ the methods but if you apply it to situations with p-values larger thanare unlikely to find any pairs that are detected as being different. Somestatisticians suggest that you shouldn?t employ follow-up tests such as Tukey?s
HSD when there is not sufficient evidence to reject the overall null hypothesis
and would be able to reasonably criticize the following results. But for the
sake of completeness, we can find the pair-wise comparison results at our
typical 95% family-wise confidence level in this situation, with the three
confidence intervals displayed in Figure 2-21. 

</span>


```{r}



```


data=</span>

MockJury)</span>
 **require</span>
(multcomp)</span>
 **Tm2 &lt;-</span>
glht</span>

linfct =</span>

mcp</span>

Attr =</span>

"Tukey"</span>

))</span>
 **confint</span>
##<spanstyle='mso-spacerun:yes'>
?? </span>
Simultaneous Confidence Intervals</span>
 **## Multiple Comparisons of Means: Tukey Contrasts</span>
 **## Quantile = 2. 3751</span>
 **## 95% family-wise confidence level</span>
 **## Linear Hypotheses:</span>
 **##<spanstyle='mso-spacerun:yes'>
?????????????????????????????? </span>
Estimate
lwr<span style='mso-spacerun:yes'>
???? </span>
upr<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## Average - Beautiful == 0<spanstyle='mso-spacerun:yes'>
????? </span>
-0. 3596<span style='mso-spacerun:yes'>
?</span>
-2. 2969<span style='mso-spacerun:yes'>
? </span>
1. 5776</span>
 **## Unattractive - Beautiful == 0<spanstyle='mso-spacerun:yes'>
? </span>
1. 4775<span style='mso-spacerun:yes'>
?</span>
-0. 4730<span style='mso-spacerun:yes'>
? </span>
3. 4280</span>
 **## Unattractive - Average == 0<spanstyle='mso-spacerun:yes'>
??? </span>
1. 8371<span style='mso-spacerun:yes'>
old. par&lt;-</span>

 </span>
par</span>
(</span>
c</span>

(</span>
1. 5</span>
, </span>
2. 5</span>
, </span>
1</span>
, </span>
1</span>
)) </span>
#Makes room on the plot forthe group names</span>
 **plot</span>
(Tm2)

</span>

cld</span>
(Tm2)</span>



##<spanstyle='mso-spacerun:yes'>
??? </span>
Beautiful<spanstyle='mso-spacerun:yes'>
????? </span>
Average Unattractive </span>
 **##<span style='mso-spacerun:yes'>
?????????</span>
"a"<span style='mso-spacerun:yes'>
????????? </span>
"a"<spanstyle='mso-spacerun:yes'>
????????? </span>
"a"</span>



</div>


 id="_x0000_i1027" type="#_x0000_t75" style='width:393pt;height:262. 15pt; visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter2_files/image099. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=524 height=350src="chapter2_files/image099. png" v:shapes="_x0000_i1027">
<![endif]>
</span>
``

`` 
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";
Figure 2-21: Tukey?s HSD confidence interval
results at the 95% family-wise confidence level. 

</span>
*


&nbsp;
</span>

At the family-wise 5%significance level, there are no pairs that are detectably different ? they all
get the same letter of "a". Now we will produce results for the reader that
thought a 10% significance was suitable for this application before seeing any
of the results. We just need to change the confidence level or significance
level that the CIs or tests are produced with inside the functions. For the </span>

``confint``  function, the </span>
``level``  option is the confidence level andfor the </span>

 it is the family-wise significance
level. Note that 90% confidence corresponds to a 10% significance level. 

</span>


```{r}



```


0. 9</span>

##<spanstyle='mso-spacerun:yes'>
?? </span>
Simultaneous Confidence Intervals</span>
 **## Multiple Comparisons of Means: Tukey Contrasts</span>
 **## Quantile = 2. 0739</span>
 **## 90% family-wise confidence level</span>
 **##<spanstyle='mso-spacerun:yes'>
?????????????????????????????? </span>
Estimatelwr<span style='mso-spacerun:yes'>
???? </span>
upr<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## Average - Beautiful == 0<spanstyle='mso-spacerun:yes'>
????? </span>
-0. 3596<span style='mso-spacerun:yes'>
?</span>
-2. 0513<span style='mso-spacerun:yes'>
? </span>
1. 3320</span>
 **## Unattractive - Beautiful == 0<spanstyle='mso-spacerun:yes'>
? </span>
1. 4775<span style='mso-spacerun:yes'>
?</span>
-0. 2257<span style='mso-spacerun:yes'>
? </span>
3. 1806</span>
 **## Unattractive - Average == 0<spanstyle='mso-spacerun:yes'>
??? </span>
1. 8371<span style='mso-spacerun:yes'>
old. par&lt;-</span>

 </span>
par</span>
(</span>
c</span>

(</span>
1. 5</span>
, </span>
2. 5</span>
, </span>
1</span>
, </span>
1</span>
)) </span>
#Makes room on the plot forthe group names</span>
 **plot</span>
(</span>
confint</span>
level=</span>

9</span>

cld</span>
(Tm2, </span>
0. 1</span>

##<spanstyle='mso-spacerun:yes'>
??? </span>
Beautiful<spanstyle='mso-spacerun:yes'>
????? </span>
Average Unattractive </span>
 **##<span style='mso-spacerun:yes'>
????????</span>
"ab"<span style='mso-spacerun:yes'>
?????????</span>
"a"<span style='mso-spacerun:yes'>
?????????</span>
"b"</span>




</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1026" type="#_x0000_t75" style='width:421. 15pt;height:280. 9pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter2_files/image100. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=562 height=375src="chapter2_files/image101. jpg" v:shapes="_x0000_i1026">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 2-22: Tukey?s HSD 90%
family-wise confidence intervals. 

</span>
*


&nbsp;
</span>

With family-wise 10% significance and90% confidence levels, the *Unattractive*and *Average* picture groups aredetected as being different but the *Average*group is not detected as different from *Beautiful*and *Beautiful* is not detected to bedifferent from *Unattractive*. Thisleaves the "overlap" of groups across the sets of groups that was noted earlier. 
as being dissimilar from levels in two different sets and so gets two differentletters. 

</span>


The beanplot (Figure 2-23)helps to clarify some of the reasons for this set of results. The detection of
a difference between *Average* and *Unattractive* just barely occurs and theother two so it ends up not being detectably different from either one. Thissort of overlap is actually a fairly common occurrence in these sorts of
situations so be prepared a mixed set of letters for some levels. 

</span>


```{r}



```


MockJury, </span>

""</span>

col=</span>

, </span>

"jitter"</span>

)</span>
 **text</span>
c</span>

(</span>
1</span>
), </span>
c</span>
(</span>
5</span>
), </span>
"ab"</span>
, </span>
"blue"</span>

cex=</span>

)</span>
 **text</span>
c</span>

(</span>
2</span>
), </span>
c</span>
(</span>
4. 8</span>
), </span>
"a"</span>
, </span>
"green"</span>

cex=</span>

)</span>
 **text</span>
c</span>

(</span>
3</span>
), </span>
c</span>
(</span>
6. 5</span>
), </span>
"b"</span>
, </span>
"red"</span>

cex=</span>

)

</span>



&nbsp;
</span>

</div>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75" style='width:504. 4pt;
 height:336. 4pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter2_files/image102. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=673 height=449src="chapter2_files/image103. jpg" v:shapes="Picture">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;mso-bidi-font-size:11. 0pt;line-height:107%;font-family:
"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:
major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure
2-23: Beanplot of sentences with compact letter display results from 10%
family-wise significance level Tukey?s HSD. Average and Unattractive picture
groups are detected as being different and are displayed as belonging to
different groups. Beautiful picture responses are not detected as different
from the other two groups. 

</span>
*


&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020893">
2. 7: Chapter Summary</a>


</h2>


????????????? </span>
Inthis chapter, we explored methods for comparing a quantitative response across *J* groups (*J* </span>

Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;mso-char-type:
symbol;mso-symbol-font-family:Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
 2), with what is called the One-Way ANOVA procedure. Theinitial test is based on assessing evidence against a null hypothesis of no
groups. There are two different methods for estimating these One-Way ANOVAmodels: the cell-means model and the reference-coded versions of the model. 
There are times when either model will be preferred, but for the rest of the text, 
the reference coding is used (sorry!). The ANOVA *F*-statistic, often presented with underlying information in theANOVA table, provides a method of assessing evidence against the null
hypothesis either using permutations or via the *F*-distribution. Pair-wise comparisons using Tukey?s HSD provide amethod for comparing all the groups and are a nice complement to the overall
ANOVA results. A compact letter display was shown that enhanced the
interpretation of Tukey?s HSD result. 

</span>


In the guinea pigexample, we are left with some lingering questions based on these results. It
appears that the effect of *dosage*changes as a function of the *deliverymethod* (OJ, VC) because the size of the differences between OJ and VC
change for different dosages. These methods can?t directly assess the question
of whether the effect of delivery method is the same or not across the
different dosages. In chapter 3, the two variables, *Dosage* and *Delivery method*are modeled as two separate variables so we can consider their effects both
separately and together. This allows more refined hypotheses, such as *Is the effect of delivery method the samefor all dosages?*, to be tested. This will introduce new models and methods
for analyzing data where there are two factors as explanatory variables in a
model for a quantitative response variable in what is called the Two-Way ANOVA. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020894">
2. 8: Summary of important R code</a>


</h2>


The main components of R code used inthis chapter follow with components to modify in red, remembering that any R
packages mentioned need to be installed and loaded for this code to have a
chance of working:

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
MODELNAME`` 
``&lt;-lm(<span style='color:red'>
Y`` ~<spanmargin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Probablythe most frequently used command in R. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Hereit is used to fit the reference-coded One-Way ANOVA model with Y as the response
variable and X as the grouping variable, storing the estimated model object in
MODELNAME. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``color:red'>
MODELNAME`` 
``&lt;-lm(<span style='color:red'>
Y`` ~<spanmargin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Fitsthe cell means version of the One-Way ANOVA model. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``summary`` 
(</span>
``MODELNAME)`` 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesmodel summary information including the estimated model coefficients, SEs, 
t-tests, and p-values. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``anova`` 
(</span>
``MODELNAME)`` 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesthe ANOVA table but  **** . 

</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Resultsare incorrect if run on the cell-means model since the reduced model under the
null is that the mean of all the observations is 0!

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``pf(<spanstyle='color:red'>
FSTATISTIC`` , df1=<span style='color:red'>
NUMDF</span>
, df2=<spanstyle='color:red'>
DENDF</span>
, lower. tail=F)

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Findsthe p-value for an observed *F*-statisticwith NUMDF and DENDF degrees of freedom. 

</span>

<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
?</span>
</span>
``par(mfrow=c(2, 2)); plot(<spanstyle='color:red'>
MODELNAME`` )

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesfour diagnostic plots including the Residuals vs Fitted and Normal Q-Q plot. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``plot(allEffects(<spanstyle='color:red'>
MODELNAME`` ))</span>


</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Plotsthe estimated model component. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe </span>

 package be loaded. 

</span>


<![if !supportLists]>
<spanstyle='font-size:12. 0pt;line-height:115%;font-family:Symbol;mso-fareast-font-family:
Symbol;mso-bidi-font-family:Symbol'>
?<spanstyle='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; </span>
</span>
</span>
<![endif]>
``Tm2&lt;-glht(<spanstyle='color:red'>
MODELNAME`` , linfct=mcp(<span style='color:red'>
X</span>
="Tukey");confint(Tm2); plot(Tm2); cld(Tm2)</span>



</span>

margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Requiresthe </span>

 package to be installed and loaded. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Canonly be run on the reference-coded version of the model. 

</span>


margin-bottom:6. 0pt;margin-left:1. 0in;mso-add-space:auto;text-indent:-. 25in;style='font-size:12. 0pt;line-height:115%;font-family:"Courier New";mso-fareast-font-family:"Courier New"'>
o<span style='font:7. 0pt "Times New Roman"'>
&nbsp;&nbsp;</span>
</span>
</span>
<![endif]>
Generatesthe text output and plot for Tukey?s HSD as well as the compact letter display. 

</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l2 level2 lfo1'>
<a
name="_Toc459020895">
2. 9: Practice problems</a>


</h2>


For thesepractice problems, you will work with the cholesterol data set from the </span>

style='font-size:12. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;font-family:"Lucida Console"'>
multcomp</span>

 package that was used to generate the Tukey?s HSDresults. To load the data set and learn more about the study, use the following
code:

</span>


style='font-size:12. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;font-family:"Lucida Console"'>
require(multcomp)

</span>


style='font-size:12. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;font-family:"Lucida Console"'>
data(cholesterol)

</span>


style='font-size:12. 0pt;mso-bidi-font-size:10. 0pt;line-height:107%;font-family:"Lucida Console"'>
help(cholesterol)

</span>


2. 1. 
Graphically explore the differences in the changes in Cholesterol levels for
the five levels using boxplots and beanplots. 

</span>


2. 2. Is the
design balanced? 

</span>


2. 3. 
Complete all 6+ steps of the hypothesis test using the parametric *F*-test, reporting the ANOVA table andthe distribution of the test statistic under the null. 

</span>


2. 4. 
Discuss the scope of inference using the information that the treatment levels
were randomly assigned to volunteers in the study. 

</span>


2. 5. Generate
the permutation distribution and find the p-value. Compare the parametric
p-value to the permutation test results. 

</span>


2. 6. 
Perform Tukey?s HSD on the data set. Discuss the results ? which pairs were
detected as different and which were not? Bigger reductions in cholesterol are
good, so are there any levels you would recommend or that might provide similar
reductions?

</span>


2. 7. Find and
interpret the CLD and compare that to your interpretation of results from 2. 6. 

</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>




</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn7>


<a style='mso-footnote-id:ftn7' href="#_ftnref7"name="_ftn7" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[7]</span>
</span>
<![endif]>
</span>
</span>
</a>
 


</div>


<div style='mso-element:footnote' id=ftn8>


<a style='mso-footnote-id:ftn8' href="#_ftnref8"name="_ftn8" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[8]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn9>


<a style='mso-footnote-id:ftn9' href="#_ftnref9"name="_ftn9" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[9]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn10>


<a style='mso-footnote-id:ftn10' href="#_ftnref10"name="_ftn10" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[10]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn11>


<a style='mso-footnote-id:ftn11' href="#_ftnref11"name="_ftn11" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[11]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn12>


<a style='mso-footnote-id:ftn12' href="#_ftnref12"name="_ftn12" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[12]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


</div>


** 

</html>

