 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l0 level2 lfo1'>
<a
name="_Toc459020967">
8. 1: The impact of simulated chronic nitrogen deposition
on the biomass and N2-fixation activity of two boreal feather
moss?cyanobacteria associations</a>


</h2>


????????????? </span>
Ina 16-year experiment, Gundale, Bach, and Nordin (2013) studied the impacts of
Nitrogen (N) additions on the mass of two feather moss species (*Pleurozium schreberi *(</span>
``PS`` )*``HS`` 
)) in the Svartberget ExperimentalForest in Sweden. They used a randomized block design: here this means that
within each of 6 blocks (pre-specified areas that were divided into three
experimental units or plots of area 0. 1 hectacre), one of the three treatments
were randomly applied.  ***Randomized block designs***  involverandomization of levels within blocks or groups as opposed to  ***completelyrandomized designs *** where each  ***experimental unit***  (the subject or plotthat will be measured) could be randomly assigned to any treatment. This is
done in agricultural studies to control for systematic differences across the
fields by making sure each treatment level is used in each area or  ***block*** of the field. 

</span>


????????????? </span>
Thethree treatments involved different levels of N applied immediately after
snowmelt, *Control* (no additional N ?just the naturally deposited amount), 12. 5 kg N ha<sup>
-1</sup>
 yr<sup>
-1</sup>
(*N12. 5*), and 50 kg N ha<sup>
-1</sup>
yr<sup>
-1</sup>
 (*N50*). Theresearchers were interested in whether the treatments would have differential
impacts on the two species of moss growth. They measured a variety of other
variables, but here we focus on the estimated *biomass* per hectare (mg/ha) of the *specie*s (*PS* or *HS*), both measured for each plot withineach block, considering differences across the *treatments* (*Control*, *N12. 5*, or *N50*). The beanplot in Figure 8-2 provides some initial informationabout the responses. Initially there seem to be some differences in the
combinations of groups and some differences in variability in the different
groups, especially with much more variability in the *control* treatment level and more variability in the *PS* responses than for the *HS* responses. 

</span>

<!--[if gte vml 1]>
<v:shape id="_x0000_i1034" type="#_x0000_t75" style='width:446. 65pt;height:279pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image029. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=596 height=372src="chapter8_files/image030. jpg" v:shapes="_x0000_i1034">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-2: Beanplot of
biomass responses by treatment and species. 

</span>
*

<span style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:
major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";

&nbsp;
</span>


```{r}



```


"http://www. math. montana. edu/courses/s217/documents/gundalebachnordin_2. csv"</span>

)</span>



(beanplot)</span>
 **beanplot</span>
data=</span>

side =</span>

"b"</span>

col =</span>

list</span>

"white"</span>

"lightgreen"</span>

xlab=</span>

, </span>

"Biomass"</span>

)</span>
 **legend</span>
"topright"</span>

bty=</span>

, </span>

c</span>
(</span>
"HS"</span>
"PS"</span>

fill =</span>

c</span>

(</span>
"white"</span>
, </span>
))

</span>


(</span>

)</span>
 **intplot</span>
data=</span>

col=</span>

(</span>

1</span>
, </span>
2</span>
), </span>
lwd=</span>
2</span>

&nbsp;
</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The Two-WAY ANOVA modelthat contains a *species* by *treatmen*t interaction is of interest (sothis has a quantitative response variable of *biomass* and two categorical predictors of *species* and *treatment*)<astyle='mso-footnote-id:ftn1' href="#_ftn1" name="_ftnref1" title="">
<sup>
<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</sup>
<![endif]>
</span>
</sup>
</a>
. We
can make an interaction plot to focus on the observed patterns of the means
across the combinations of levels as provided in Figure 8-3. The interaction
plot suggests a relatively additive pattern of differences between PS and HS
across the three treatment levels. However, the variability seems to be quite
different based on this plot as well. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1033" type="#_x0000_t75" style='width:430. 9pt;height:269. 25pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image031. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=575 height=359src="chapter8_files/image032. jpg" v:shapes="_x0000_i1033">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-3: Interaction plot
of biomass responses by treatment and species. 

</span>
*


&nbsp;
</span>

Based on the initial plots, we aregoing to be concerned about the equal variance assumption initially. We can fit
the interaction model and explore the diagnostic plots to verify that we have a
problem. 

</span>


```{r}



```


data=</span>

gdn)</span>
 **summary</span>
##<spanstyle='mso-spacerun:yes'>
????????????????????????? </span>
Estimate Std. Error t
 **## (Intercept)<spanstyle='mso-spacerun:yes'>
?????????????? </span>
1694. 80<spanstyle='mso-spacerun:yes'>
???? </span>
211. 86<span style='mso-spacerun:yes'>
??</span>
8. 000 6. 27e-09 ***</span>
 **## SpeciesPS<spanstyle='mso-spacerun:yes'>
????????????????? </span>
859. 88<spanstyle='mso-spacerun:yes'>
???? </span>
299. 62<span style='mso-spacerun:yes'>
??</span>
2. 870<span style='mso-spacerun:yes'>
? </span>
0. 00745 ** </span>
 **## TreatmentN12. 5<spanstyle='mso-spacerun:yes'>
??????????? </span>
-588. 26<spanstyle='mso-spacerun:yes'>
???? </span>
299. 62<span style='mso-spacerun:yes'>
?</span>
-1. 963<span style='mso-spacerun:yes'>
? </span>
0. 05893 . <spanstyle='mso-spacerun:yes'>
? </span>
</span>
 **## TreatmentN50<spanstyle='mso-spacerun:yes'>
???????????? </span>
-1182. 91<spanstyle='mso-spacerun:yes'>
???? </span>
299. 62<span style='mso-spacerun:yes'>
?</span>
-3. 948<span style='mso-spacerun:yes'>
? </span>
0. 00044 ***</span>
 **## SpeciesPS:TreatmentN12. 5<spanstyle='mso-spacerun:yes'>
?? </span>
199. 42<span style='mso-spacerun:yes'>
????</span>
423. 72<span style='mso-spacerun:yes'>
?? </span>
0. 471<spanstyle='mso-spacerun:yes'>
? </span>
0. 64130<span style='mso-spacerun:yes'>
???</span>
</span>
 **## SpeciesPS:TreatmentN50<spanstyle='mso-spacerun:yes'>
????? </span>
88. 29<span style='mso-spacerun:yes'>
????</span>
423. 72<span style='mso-spacerun:yes'>
?? </span>
0. 208<spanstyle='mso-spacerun:yes'>
? </span>
0. 83636<span style='mso-spacerun:yes'>
???</span>
</span>
 **## </span>
 **## Residual standard error: 519 on 30 degrees offreedom</span>
 **## Multiple R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 6661, Adjusted R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 6104 </span>
 **## F-statistic: 11. 97 on 5 and 30 DF, <spanstyle='mso-spacerun:yes'>
?</span>
p-value: 2. 009e-06</span>


</span>

(</span>

c</span>

(</span>
2</span>
, </span>
2</span>
), </span>
oma=</span>
(</span>

0</span>
, </span>
0</span>
, </span>
2</span>
, </span>
0</span>
))</span>
 **plot</span>
(m1, </span>
sub. caption=</span>
"InitialMassperha 2-WAY model"</span>
</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


There is a clear problem withnon-constant variance showing up in a fanning shape<a style='mso-footnote-id:
ftn2' href="#_ftn2" name="_ftnref2" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>

in the Residuals versus Fitted and Scale-Location plots in Figure 8-4. 
Interestingly, the normality assumption is not an issue so hopefully we will
not worsen this result by using a transformation to try to address the
non-constant variance issue. The independence assumption is violated in two
ways for this model by this study design ? the blocks create clusters or groups
of observations and the block should be accounted for (they did this in their
categorical variable to their models). Using blocked designs and accounting forthe blocks in the model will typically give more precise inferences for the
effects of interest, the treatments randomized within the blocks. Additionally, 
 **** within block, one for *SP* and one for *HS* and these might be related (for
example, high *HS* biomass might beassociated with high *SP*) so puttingboth observations into a model  ****  at a second level. It takes more advancedstatistical models (called linear mixed models) to see how to fully deal with
this, for now it is important to recognize the issues. The more complicated
models provide similar results here and include the *treatment* by *species*interaction we are going to explore, they just add to this basic model to
account for these other issues. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1032" type="#_x0000_t75" style='width:376. 9pt;height:235. 5pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image033. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=503 height=314src="chapter8_files/image034. jpg" v:shapes="_x0000_i1032">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-4: Diagnostic plots
of treatment by species interaction model for Biomass. </span>
**

</span>
*

``color:black;background:#E1E2E5'>

&nbsp;
`` 

Remember that  **before using a *log*** ,  **** :

</span>


```{r}



```





##<spanstyle='mso-spacerun:yes'>
??? </span>
Min. 1st Qu. <spanstyle='mso-spacerun:yes'>
? </span>
Median<span style='mso-spacerun:yes'>
???</span>
Mean 3rd Qu. <span style='mso-spacerun:yes'>
??? </span>
Max. </span>
 **##<span style='mso-spacerun:yes'>
?? </span>
319. 1<spanstyle='mso-spacerun:yes'>
? </span>
1015. 0<span style='mso-spacerun:yes'>
?</span>
1522. 0<span style='mso-spacerun:yes'>
? </span>
1582. 0<spanstyle='mso-spacerun:yes'>
? </span>
2027. 0<span style='mso-spacerun:yes'>
?</span>
3808. 0

</span>



&nbsp;
</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The minimum is 319. 1 so it is safe toapply the natural log-transformation to the response variable (*Biomass*) and repeat the previous plots:

</span>

```{r}



```


(gdn$Massperha)</span>
 **par</span>
mfrow=</span>

(</span>

1</span>
, </span>
2</span>
))</span>
 **beanplot</span>
data=</span>

side =</span>

"b"</span>

col =</span>

list</span>

"white"</span>

"lightgreen"</span>

xlab=</span>

, </span>

"log-Biomass"</span>

main=</span>

)</span>
 **legend</span>
"topright"</span>

bty=</span>

, </span>

c</span>
(</span>
"HS"</span>
"PS"</span>

fill =</span>

c</span>

(</span>
"white"</span>
, </span>
))</span>
 **intplot</span>
data=</span>

col=</span>

(</span>

1</span>
, </span>
2</span>
), </span>
lwd=</span>
2</span>
, </span>
main=</span>
"(b)"</span>
</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The variability in the beanplot inFigure 8-5(a) appears to be more consistent across the groups but the lines
appear to be a little less parallel in the interaction plot Figure 8-5(b) for
the log-scale response. That is not problematic but suggests that we may now
have an interaction present ? it is hard to tell visually sometimes. Again, 
fitting the interaction model and exploring the diagnostics is the best way to
assess the success of the transformation applied. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1031" type="#_x0000_t75"
 style='width:482. 65pt;height:301. 9pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image035. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=644 height=403src="chapter8_files/image036. jpg" v:shapes="_x0000_i1031">
<![endif]>
</span>
style='font-size:10. 0pt;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:
#E1E2E5'>


</span>


<i"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;mso-hansi-theme-font:major-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure
8-5: Beanplot and interaction plot of the log-Biomass responses by treatment


</span>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

????????????? </span>
Thelog(Mass per ha) version of the response variable has little issue with
changing variability present in the residuals in Figure 8-6 with much more
similar variation in the residuals across the fitted values. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1030" type="#_x0000_t75" style='width:447. 75pt;height:279. 75pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image037. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=597 height=373src="chapter8_files/image038. jpg" v:shapes="_x0000_i1030">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-6: Diagnostic plots
of treatment by species interaction model for log-Biomass. </span>
*

</span>


&nbsp;
</span>

The normality assumption is leaningtowards a slight violation with too little variability in the right tail and so
maybe a little bit of a left skew. This is only a minor issue and fixes the
other big issue (clear non-constant variance), so this model is at least closer
to giving us trustworthy inferences than the original model. The model presents
some evidence of a *Species* by *Treatment* interaction (*F*(2, 30)=4. 2, p-value=0. 026). This
suggests that the effects on the log-biomass of the treatments differ between
the two species. The mean log-biomass is lower for *HS* than *PS* with themean log-biomass to decrease more rapidly than for *PS*. In other words, increasing nitrogen has more of an impact on
the resulting log-biomass for *HS* thanfor *PS*. The highest mean log-biomassrates were observed under the control conditions for both species making
nitrogen appear to inhibit growth of these species. 

</span>


```{r}



```


data=</span>

gdn)</span>
 **summary</span>
##<spanstyle='mso-spacerun:yes'>
????????????????????????? </span>
Estimate Std. Error tvalue Pr(&gt;|t|)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept)<spanstyle='mso-spacerun:yes'>
??????????????? </span>
7. 4108<spanstyle='mso-spacerun:yes'>
???? </span>
0. 1160<span style='mso-spacerun:yes'>
?</span>
63. 902<span style='mso-spacerun:yes'>
? </span>
&lt; 2e-16 ***</span>
 **## SpeciesPS<spanstyle='mso-spacerun:yes'>
????????????????? </span>
0. 3921<spanstyle='mso-spacerun:yes'>
???? </span>
0. 1640<span style='mso-spacerun:yes'>
??</span>
2. 391<span style='mso-spacerun:yes'>
? </span>
0. 02329 *<spanstyle='mso-spacerun:yes'>
? </span>
</span>
 **## TreatmentN12. 5<spanstyle='mso-spacerun:yes'>
??????????? </span>
-0. 4228<spanstyle='mso-spacerun:yes'>
???? </span>
0. 1640<span style='mso-spacerun:yes'>
?</span>
-2. 578<span style='mso-spacerun:yes'>
? </span>
0. 01510 *<spanstyle='mso-spacerun:yes'>
? </span>
</span>
 **## TreatmentN50<spanstyle='mso-spacerun:yes'>
????????????? </span>
-1. 1999<spanstyle='mso-spacerun:yes'>
???? </span>
0. 1640<span style='mso-spacerun:yes'>
?</span>
-7. 316 3. 79e-08 ***</span>
 **## SpeciesPS:TreatmentN12. 5<spanstyle='mso-spacerun:yes'>
?? </span>
0. 2413<span style='mso-spacerun:yes'>
????</span>
0. 2319<span style='mso-spacerun:yes'>
? </span>
?</span>
1. 040<span style='mso-spacerun:yes'>
?</span>
0. 30645<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## SpeciesPS:TreatmentN50<spanstyle='mso-spacerun:yes'>
???? </span>
0. 6616<span style='mso-spacerun:yes'>
????</span>
0. 2319<span style='mso-spacerun:yes'>
?? </span>
2. 853<spanstyle='mso-spacerun:yes'>
? </span>
0. 00778 ** </span>
 **## </span>
 **## Residual standard error: 0. 2841 on 30 degrees offreedom</span>
 **## Multiple R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 7998, Adjusted R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 7664 </span>
 **## F-statistic: 23. 96 on 5 and 30 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: 1. 204e-09</span>




(car)</span>
 **Anova</span>
## AnovaTable (Type II tests)</span>
 **## </span>
 **## Response: logMassperha</span>
 **##<span style='mso-spacerun:yes'>
??????????????????</span>
Sum Sq Df F value<span style='mso-spacerun:yes'>
???</span>
Pr(&gt;F)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## Species<span style='mso-spacerun:yes'>
??????????</span>
4. 3233<span style='mso-spacerun:yes'>
? </span>
1<spanstyle='mso-spacerun:yes'>
? </span>
53. 577 3. 755e-08 ***</span>
 **## Treatment<span style='mso-spacerun:yes'>
????????</span>
4. 6725<span style='mso-spacerun:yes'>
? </span>
2<spanstyle='mso-spacerun:yes'>
? </span>
28. 952 9. 923e-08 ***</span>
 **## Species:Treatment 0. 6727<spanstyle='mso-spacerun:yes'>
? </span>
2<span style='mso-spacerun:yes'>
??</span>
4. 168<span style='mso-spacerun:yes'>
?? </span>
0. 02528 *<spanstyle='mso-spacerun:yes'>
? </span>
</span>
 **## Residuals<span style='mso-spacerun:yes'>
????????</span>
2. 4208 30<span style='mso-spacerun:yes'>
????????????????????? </span>
</span>
 **par</span>
mfrow=</span>

(</span>

2</span>
, </span>
2</span>
), </span>
oma=</span>
c</span>
(</span>
0</span>
, </span>
0</span>
, </span>
2</span>
, </span>
0</span>
))</span>
 **plot</span>
(m2, </span>
sub. caption=</span>
"log-Massperha2-WAY model"</span>
</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The researchers actuallyapplied a log(y+1) transformation to all the variables. This was used because
one of their many variables had a value of 0 and so they added 1 to avoid
analyzing a ?&#8734; response. This was not needed for most of their variables
because most did not attain the value of 0. Adding a small value to
observations and then log-transforming is a common but completely arbitrary
practice and the choice of the added value can impact the results. Sometimes
considering a square-root transformation can accomplish similar benefits as the
log-transform and be applied safely to responses that include 0s. Or more
complicated statistical models can be used that allow 0s in responses and still
account for the violations of the linear model assumptions ? see a statistician
or continue exploring more advanced statistical methods for ideas in this
direction. 

</span>


The term-plot in Figure8-7 provides another display of the results with some information on the
results for each combination of the species and treatments. Finding evidencethat the treatments caused different results for the different species is a
good first start. And it appears that there are some clear differences among
certain combinations such as the mean for *PS-Control*is clearly larger than for *HS*-*N50*. The researchers were probably
really interested in whether the *N12. 5*results differed from *Control* for *HS* and whether the *species* differed at *Control*sites. As part of performing all pair-wise comparisons, we can assess those
sorts of detailed questions. This sort of follow-up could be considered in any
Two-Way ANOVA model but will be most interesting in situations where there are
important interactions. 

</span>


```{r}



```


(m2), </span>

T, </span>

"bars"</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1029" type="#_x0000_t75" style='width:263. 25pt;height:164. 65pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image039. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=351 height=220src="chapter8_files/image040. jpg" v:shapes="_x0000_i1029">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-7: Term-plot of the
interaction model for log-biomass. </span>
*

</span>

 **Follow-up Pairwise Comparisons:

</span>
** 

????????????? </span>
Givenstrong evidence of an interaction, many researchers would like more details
about the source of the differences. We can re-fit the model with a unique mean
for each combination of the two predictor variables, fitting a One-Way ANOVA
model (here with six levels) and using Tukey's HSD to provide safe inferences
for differences among pairs of the true means. There are six groups
corresponding to all combinations of *Species*(*HS*, *PS*) and treatment levels (*Control*, *N12. 5*, and *N50*) provided in the new variable </span>

``SpTrt``  by the </span>
``interaction``  function with new levels of *HS. Control*, *PS. Control*, *HS. N12. 5*, *PS. N12. 5*, *HS. N50*, and *PS. N50*. The One-WayANOVA *F*-test (*F*(5, 30)=23. 96, p-value&lt;0. 0001) suggests that there is evidenceof some difference in the true mean log-biomass among the six treatment combinations. 
Note that the One-Way ANOVA table contains the test for at least one of those
means being different from the others; the interaction test above was testing a
more refined hypothesis ? does the effect of treatment differ between the two
species? With a small p-value from the overall One-Way ANOVA test, the
pair-wise comparisons should be of interest. 

</span>


```{r}



```

 **gdn$SpTrt&lt;-</span>
interaction</span>
(gdn$Species, gdn$Treatment)</span>
 **levels</span>
(gdn$SpTrt)</span>



?? </span>
"PS. N12. 5"<spanstyle='mso-spacerun:yes'>
? </span>
?</span>
"HS. N50"<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## [6] "PS. N50"</span>



lm</span>

data=</span>

gdn)</span>
 **require</span>
(car)</span>
 **Anova</span>
## AnovaTable (Type II tests)</span>
 **## </span>
 **## Response: logMassperha</span>
 **##<span style='mso-spacerun:yes'>
?????????? </span>
SumSq Df F value<span style='mso-spacerun:yes'>
??? </span>
Pr(&gt;F)<spanstyle='mso-spacerun:yes'>
??? </span>
</span>
 **## SpTrt<span style='mso-spacerun:yes'>
????</span>
9. 6685<span style='mso-spacerun:yes'>
? </span>
5<spanstyle='mso-spacerun:yes'>
? </span>
23. 963 1. 204e-09 ***</span>
 **## Residuals 2. 4208 30<spanstyle='mso-spacerun:yes'>
????????????????????? </span>
</span>
 **
<![if !supportLineBreakNewLine]>
 **<![endif]>



(multcomp)</span>
 **PWnewm2 &lt;-</span>
glht</span>

linfct =</span>

mcp</span>

SpTrt =</span>

"Tukey"</span>

))</span>
 **?</span>
</span>
confint</span>
##<spanstyle='mso-spacerun:yes'>
?? </span>
Simultaneous Confidence Intervals</span>
 **## Multiple Comparisons of Means: Tukey Contrasts</span>
 **## Quantile = 3. 0418</span>
 **## 95% family-wise confidence level</span>
 **## Linear Hypotheses:</span>
 **##<spanstyle='mso-spacerun:yes'>
????????????????????????????? </span>
Estimate lwr<spanstyle='mso-spacerun:yes'>
????? </span>
upr<span style='mso-spacerun:yes'>
????</span>
</span>
 **## PS. Control - HS. Control == 0<spanstyle='mso-spacerun:yes'>
? </span>
0. 39210 -0. 10677<spanstyle='mso-spacerun:yes'>
? </span>
0. 89098</span>
 **## HS. N12. 5 - HS. Control == 0<spanstyle='mso-spacerun:yes'>
?? </span>
-0. 42277 -0. 92165<spanstyle='mso-spacerun:yes'>
? </span>
0. 07610</span>
 **## PS. N12. 5 - HS. Control == 0<spanstyle='mso-spacerun:yes'>
? </span>
??</span>
0. 21064 -0. 28823<spanstyle='mso-spacerun:yes'>
? </span>
0. 70952</span>
 **## HS. N50 - HS. Control == 0<spanstyle='mso-spacerun:yes'>
???? </span>
-1. 19994 -1. 69881 -0. 70106</span>
 **## PS. N50 - HS. Control == 0<spanstyle='mso-spacerun:yes'>
???? </span>
-0. 14620 -0. 64508<spanstyle='mso-spacerun:yes'>
? </span>
0. 35267</span>
 **## HS. N12. 5 - PS. Control == 0<spanstyle='mso-spacerun:yes'>
?? </span>
-0. 81487 -1. 31375 -0. 31600</span>
 **## PS. N12. 5 - PS. Control == 0<spanstyle='mso-spacerun:yes'>
?? </span>
-0. 18146 -0. 68033<spanstyle='mso-spacerun:yes'>
? </span>
0. 31742</span>
 **## HS. N50 - PS. Control == 0<spanstyle='mso-spacerun:yes'>
???? </span>
-1. 59204 -2. 09091 -1. 09316</span>
 **## PS. N50 - PS. Control == 0<spanstyle='mso-spacerun:yes'>
???? </span>
-0. 53830 -1. 03718 -0. 03943</span>
 **## PS. N12. 5 - HS. N12. 5 == 0<spanstyle='mso-spacerun:yes'>
????? </span>
0. 63342<span style='mso-spacerun:yes'>
?</span>
0. 13454<span style='mso-spacerun:yes'>
? </span>
1. 13229</span>
 **## HS. N50 - HS. N12. 5 == 0<spanstyle='mso-spacerun:yes'>
?????? </span>
-0. 77717 -1. 27604 -0. 27829</span>
 **## PS. N50 - HS. N12. 5 == 0<spanstyle='mso-spacerun:yes'>
??????? </span>
0. 27657 -0. 22230<spanstyle='mso-spacerun:yes'>
? </span>
0. 77545</span>
 **## HS. N50 - PS. N12. 5 == 0<spanstyle='mso-spacerun:yes'>
?????? </span>
-1. 41058 -1. 90946 -0. 91171</span>
 **## PS. N50 - PS. N12. 5 == 0<spanstyle='mso-spacerun:yes'>
?????? </span>
-0. 35685 -0. 85572<spanstyle='mso-spacerun:yes'>
? </span>
0. 14203</span>
 **## PS. N50 - HS. N50 == 0<spanstyle='mso-spacerun:yes'>
????????? </span>
1. 05374<spanstyle='mso-spacerun:yes'>
? </span>
0. 55486<span style='mso-spacerun:yes'>
?</span>
1. 55261</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


We can also generate the CompactLetter Display (CLD) to help us group up the results. 

</span>


```{r}



```





## HS. ControlPS. Control<span style='mso-spacerun:yes'>
?? </span>
HS. N12. 5<spanstyle='mso-spacerun:yes'>
?? </span>
PS. N12. 5<span style='mso-spacerun:yes'>
????</span>
HS. N50<span style='mso-spacerun:yes'>
???? </span>
PS. N50 </span>
 **##<span style='mso-spacerun:yes'>
??????</span>
"bd"<span style='mso-spacerun:yes'>
???????</span>
"d"<span style='mso-spacerun:yes'>
???????</span>
"b"<span style='mso-spacerun:yes'>
?????? </span>
"cd"<spanstyle='mso-spacerun:yes'>
??????? </span>
"a"<spanstyle='mso-spacerun:yes'>
?????? </span>
"bc"</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


We can also add it to an interaction plotto create Figure 8-8. Researchers often use displays like this to simplify the
presentation of pair-wise comparisons. Sometimes researchers add bars or stars
to provide the same information about pairs that are or are not detectably
different. The following code creates the plot of these results using our </span>

``intplot``  function and the </span>
``cld=T``  option. 

</span>

```{r}



```


T, </span>

. </span>

15</span>
, </span>
cldcol=</span>
(</span>

2</span>
, </span>
3</span>
, </span>
4</span>
, </span>
5</span>
, </span>
6</span>
, </span>
8</span>
), </span>
gdn, </span>

2</span>

, </span>
main=</span>
"Interaction with CLD from Tukey's HSD on One-WayANOVA"</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


These results suggestfrom all the other groups (letter "a"). The rest of the story is morecomplicated since many of the sets contain overlapping groups in terms of
detectable differences. Some specific aspects of those results are most
interesting. The mean log-biomasses were not detectably different between the
(they share a "d"). In other words, without treatment, there is no evidence ofa difference in how much of the two species are present in the sites. For *N12. 5* and *N50* treatments, there are detectable differences between the S*pecies*. These comparisons are probablyof the most interest initially and suggest that the treatments have a different
impact on the two species, remembering that in the control treatments, the
results for the two species were not detectably different. Further explorations
of the sizes of the differences that can be extracted from selected confidence
intervals in the Tukey's HSD results printed above. 

</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1028" type="#_x0000_t75" style='width:333pt;height:208. 15pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image041. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=444 height=278src="chapter8_files/image042. jpg" v:shapes="_x0000_i1028">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-8: Interaction plot
for log-biomass with CLD from Tukey's HSD for all pairwise comparisons. </span>
*

</span>


&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l0 level2 lfo1'>
<a
name="_Toc459020968">
8. 2: Ants learn to rely on more informative attributes
during decision-making</a>


</h2>


????????????? </span>
InSasaki and Pratt (2013), a set of ant colonies were randomly assigned to one of
two treatments to study whether the ants could be "trained" to have a
preference for or against certain attributes for potential nest sites. The
colonies were either randomly assigned to experience the repeated choice of two
identical colony sites except for having an inferior light or entrance size
attribute. Then the ants were allowed to choose between two nests, one that had
a large entrance but was dark and the other that had a small entrance but was
bright. 54 of the 60 colonies that were randomly assigned to one of the two
treatments completed the experiment by making a choice between the two types of
sites. The data set and some processing code follows. 

</span>


????????????? </span>
Thefirst question of interest is what type of analysis is appropriate here. Once
we recognize that there are two categorical variables being considered (*Treatment* group with two levels and *After* treatment choice with two levels *SmallBright* or *LargeDark*), then this is recognized as being within our Chi-square
testing framework. The random assignment of colonies (the subjects here) to
treatment levels tells us that the  ***Chi-square Homogeneity test***  isappropriate here and that we can make causal statements if we find evidence of
differences in the patterns of responses. 

</span>


```{r}



```


"http://www. math. montana. edu/courses/s217/documents/sasakipratt. csv"</span>

)</span>
 **sasakipratt$group&lt;-</span>
factor</span>
(sasakipratt$group)</span>
 **levels</span>
(sasakipratt$group)&lt;-</span>
c</span>
(</span>
"Light"</span>
, </span>
)</span>
 **sasakipratt$after&lt;-</span>
(sasakipratt$after)</span>
 **levels</span>
c</span>

(</span>
"SmallBright"</span>
, </span>
)</span>
 **sasakipratt$before&lt;-</span>
(sasakipratt$before)</span>
 **levels</span>
c</span>

(</span>
"SmallBright"</span>
, </span>
)</span>
 **plot</span>
data=</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1027" type="#_x0000_t75" style='width:361. 15pt;height:225. 75pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image043. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=482 height=301src="chapter8_files/image044. jpg" v:shapes="_x0000_i1027">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-9: Stacked bar chart
for Ant Colony results. </span>
*

</span>

```{r}



```

 **tally</span>
data=</span>

##<spanstyle='mso-spacerun:yes'>
?????????? </span>
after</span>
 **## group<span style='mso-spacerun:yes'>
?????</span>
SmallBright LargeDark</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
Light<spanstyle='mso-spacerun:yes'>
???????????? </span>
19<spanstyle='mso-spacerun:yes'>
???????? </span>
9</span>
 **##<span style='mso-spacerun:yes'>
??</span>
Entrance<span style='mso-spacerun:yes'>
?????????? </span>
9<spanstyle='mso-spacerun:yes'>
??????? </span>
17

</span>

tally</span>

data=</span>

margins=</span>



</span>


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The null hypothesis of interest hereis that there is no difference in the distribution of responses on *After* ? the rates of their choice of denin the population of all ant colonies like those studied. The alternative isthat there is some difference in the distributions of *After* between the *groups*in the population. 

</span>


????????????? </span>
Touse the Chi-square distribution to find a p-value for the X<sup>
2</sup>

statistic, we need all the expected cell counts to be larger than 5, so we
should check that. Note that in the following, the </span>

``correct=F``  option is used to keep the functionfrom slightly modifying the statistic used that occurs when overall sample
sizes are small. 

</span>


```{r}



```


F)$expected</span>




##<spanstyle='mso-spacerun:yes'>
?????????? </span>
after</span>
 **## group<span style='mso-spacerun:yes'>
?????</span>
SmallBright LargeDark</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
Light<spanstyle='mso-spacerun:yes'>
?????? </span>
14. 51852<span style='mso-spacerun:yes'>
?</span>
13. 48148</span>
 **##<span style='mso-spacerun:yes'>
??</span>
Entrance<span style='mso-spacerun:yes'>
??? </span>
13. 48148<spanstyle='mso-spacerun:yes'>
? </span>
12. 51852

</span>


&nbsp;


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Our expected cell count condition ismet, so we can proceed to explore the results of the test:

</span>


```{r}



```


F)</span>




## </span>
 **##<span style='mso-spacerun:yes'>
? </span>
Pearson'sChi-squared test</span>
 **## </span>
 **## data:<span style='mso-spacerun:yes'>
? </span>
table1</span>
 **## X-squared = 5. 9671, df = 1, p-value = 0. 01458</span>



</div>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

<spanstyle='font-size:12. 0pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
"Courier New"'>
The X<sup>
2</sup>
 statistic is 5. 97 which, if our assumptions
are met, should approximately follow a Chi-squared distribution with (*R*-1)*(*C*-1)=1 degrees of freedom under the null hypothesis. The p-value is
0. 015, suggesting that there is good evidence against the null hypothesis. We
can conclude that there is a difference in the distribution of the responses
between the two treated groups in the population of all ant colonies that could
have been treated. Because of the random assignment, we can say that the
treatments caused differences in the colony choices. These results cannot be
extended to ants beyond those being studied by these researchers because they
were not randomly selected. 

</span>


<spanstyle='font-size:12. 0pt;mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:
exploration of the standardized residuals can provide more insights in some situations, although here they are similar for all the cells:

</span>


```{r}



```


correct=</span>

##<spanstyle='mso-spacerun:yes'>
?????????? </span>
after</span>
 **## group<span style='mso-spacerun:yes'>
?????</span>
SmallBright LargeDark</span>
 **##<span style='mso-spacerun:yes'>
?? </span>
Light<spanstyle='mso-spacerun:yes'>
?????? </span>
1. 176144 -1. 220542</span>
 **##<span style='mso-spacerun:yes'>
??</span>
Entrance<span style='mso-spacerun:yes'>
?? </span>
-1. 220542<spanstyle='mso-spacerun:yes'>
? </span>
1. 266616</span>




</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


When all the standardized residualcontributions are similar, that suggests that there are differences in all the
cells from what we would expect if the null hypothesis were true. Basically, 
that means that what we observed is a bit larger than expected for the *Light* treatment group in the *SmallBright* choice and lower than
treated ants preferred the small and bright den. And for the *Entrance* treated group, they preferred
the large entrance, dark den at a higher rate than expected if the null is true
and lower than expected in the small entrance, bright location. 

</span>


????????????? </span>
Theresearchers extended this basic result a little further using a statistical
model called  ***logistic regression*** , which involves using something like alinear model but with a categorical response variable (well ? it actually only works
for a two-category response variable). They also had measured which of the two
types of dens that each colony chose before treatment and controlled for that
choice. So the actual model used in their paper contained two predictor
variables ? the randomized treatment received that we explored here and the
prior choice of den type. The interpretation of their results related to the
same treatment effect, but they were able to discuss it after adjusting for the
colonies previous selection. Their conclusions were similar to those found with
our simpler analysis. Logistic regression models are a special case of what are
called generalized linear models and are a topic for the next level of statistics
if you continue exploring. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l0 level2 lfo1'>
<a
name="_Toc459020969">
8. 3: Multi-variate models are essential for understanding
vertebrate diversification in deep time</a>


</h2>


????????????? </span>
Bensonand Mannion (2012) published a paleontology study that considered modeling the
diversity of *Sauropodomorphs* across *n*=26 "stage-level" time bins. Diversityis measured by the count of the number of different species that have been
found in a particular level of fossils. Specifically, the counts in the *Sauropodomorphs* group were obtained fora stage between *Carnian* and *Maastrichtian*, with the first threestages in the *Triassic*, the next tenin the *Jurassic*, and the last elevenconcerned about variation in sampling efforts and the ability of paleontologiststo find fossils across different stages creating a false impression of the
changes in biodiversity over time. They first wanted to see if the species
counts were related to factors such as the count of dinosaur-bearing-formations
(*DBF*) and the count ofdinosaur-bearing-collections (*DBC*)that have been identified for each period. The thought is that if there are
more formations or collections of fossils from certain stages, the diversity
might be better counted (more found of those available to find) and those
stages with less information available might be under-counted. They also
but did not consider it in their models since they want to reflect thediversity and longer stages would likely have higher diversity. 

</span>


????????????? </span>
Theirmain goal was to develop a model that would  ****  the effects of sampling efforts and allow them toperform inferences for whether the diversity was different between the *Triassic/Jurassic* (grouped together) andconsidered models that included two different versions of sampling effortvariables and one for the comparisons of periods (an indicator variable *TJK*: 0 if the observation is in *Triassic* or *Jurassic* or 1 if in *Cretaceous*). quantitative variables because the untransformed variables created diagnosticissues including influential points. They explored a model just based on the *DBC* predictor<a style='mso-footnote-id:ftn3' href="#_ftn3" name="_ftnref3" title="">
<sup>
<![if !supportFootnotes]>
<sup>
<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;
minor-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
[3]</span>
</sup>
<![endif]>
</span>
</sup>
</a>

and they analyzed the residuals from that model to see if the biodiversity was
different in the *Cretaceous* orbefore, finding a "p-value **** 0. 0001"(I think they meant &lt;0. 0001<a style='mso-footnote-id:ftn4' href="#_ftn4"
name="_ftnref4" title="">
<![if !supportFootnotes]>
<spanfont-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:minor-latin;mso-hansi-theme-font:minor-latin;
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>
). 
They were comparing the MLR models you learned to some extended regression
models that incorporated a correction for correlation in the responses over
time, but we can proceed with fitting some of their MLR models and using an AIC
comparison similar to what they used. There are some obvious flaws in their
analysis and results that we will avoid<a style='mso-footnote-id:ftn5'
href="#_ftn5" name="_ftnref5" title="">
<sup>
<![if !supportFootnotes]>
<sup>
<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;
minor-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
[5]</span>
</sup>
<![endif]>
</span>
</sup>
</a>
. 


</span>


????????????? </span>
First, we start with a plot of the log-diversity vs the log-dinosaur bearing
collections by period. We can see fairly strong positive relationships between
the log amounts of collections and species found with potentially similar
slopes for the two periods but what look like different intercepts. Especially
for *TJK* level 1 (*Cretaceous* period) observations, we might need to worry about acurving relationship. Note that a similar plot can also be made using the
formations version of the quantitative predictor variable and that the research
questions involve whether *DBF* or *DBC* are better predictor variables. 

</span>

```{r}



```


"http://www. math. montana. edu/courses/s217/documents/bensonmanion. csv"</span>

)</span>
 ** **bm$logDBCs&lt;-</span>
(bm$DBCs)</span>
 **bm$logDBFs&lt;-</span>
(bm$DBFs)</span>
 **bm$TJK&lt;-</span>
(bm$TJK)

</span>
</span>


require</span>
</span>
(car)</span>
</span>
 **scatterplot</span>
(logSpecies~logDBCs|TJK, </span>
data=</span>
</span>
bm, </span>
smooth=</span>
</span>
"Scatterplot oflog-diversity vs log-DBCs by period"</span>
</span>
2</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1026" type="#_x0000_t75" style='width:391. 15pt;height:244. 5pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image045. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=522 height=326src="chapter8_files/image046. jpg" v:shapes="_x0000_i1026">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-10: Scatterplot of
log-biodiversity vs log-DBCs by TJK. </span>
*

</span>

????????????? </span>
Thefollowing results will allow us to explore models similar to theirs. One "full"
 style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image047. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image048. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image049. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image050. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
aswell as the simpler models that each suggests (</span>
<!--[if gte msEquation 12]>
<m:oMath>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
 "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:228pt;height:17. 65pt'>

 <v:imagedata src="chapter8_files/image051. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=304 height=24src="chapter8_files/image052. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
 style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:228pt;height:17. 65pt'>

 <v:imagedata src="chapter8_files/image053. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=304 height=24src="chapter8_files/image054. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
 style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:192. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image055. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=257 height=26src="chapter8_files/image056. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
 style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:132. 4pt;height:17. 65pt'>

 <v:imagedata src="chapter8_files/image057. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=177 height=24src="chapter8_files/image058. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
). Both versions of the models (based on *DBF* or *DBC*) start with an MLR model with a quantitative variable and two
slopes. We can obtain some of the needed model selection results from the first
full model using:

</span>


```{r}



```


data=</span>

bm)</span>
 **require</span>
(MuMIn)</span>
 **options</span>
na. action =</span>

"na. fail"</span>

)</span>
 **dredge</span>
</span>
(bd1, </span>
</span>
rank=</span>
</span>
"AIC"</span>
</span>
, </span>
</span>
extra=</span>
</span>

 </span>
</span>
c</span>
</span>
(</span>
</span>
"R^2"</span>
</span>
, </span>
</span>
adjRsq=</span>
</span>
function(x)</span>
</span>

summary</span>
</span>
(x)$adj. r. squared))</span>
</span>


</span>

## Globalmodel call: lm(formula = logSpecies ~ logDBCs + TJK, data = bm)</span>
 **## ---</span>
 **## Model selection table </span>
 **##<span style='mso-spacerun:yes'>
??</span>
(Intrc)<span style='mso-spacerun:yes'>
? </span>
lgDBC TJK<spanstyle='mso-spacerun:yes'>
????? </span>
R^2<span style='mso-spacerun:yes'>
??</span>
adjRsq df<span style='mso-spacerun:yes'>
? </span>
logLik<spanstyle='mso-spacerun:yes'>
? </span>
AIC delta weight</span>
 **## 4 -1. 0890 0. 7243<span style='mso-spacerun:yes'>
??</span>
+ 0. 580900<span style='mso-spacerun:yes'>
? </span>
0. 54440<spanstyle='mso-spacerun:yes'>
? </span>
4 -12. 652 33. 3 <spanstyle='mso-spacerun:yes'>
?</span>
0. 00<span style='mso-spacerun:yes'>
?</span>
0. 987</span>
 **## 2<span style='mso-spacerun:yes'>
? </span>
0. 19880. 4283<span style='mso-spacerun:yes'>
???? </span>
0. 369100<spanstyle='mso-spacerun:yes'>
? </span>
0. 34280<span style='mso-spacerun:yes'>
?</span>
3 -17. 969 41. 9<span style='mso-spacerun:yes'>
? </span>
8. 63<spanstyle='mso-spacerun:yes'>
? </span>
0. 013</span>
 **## 1<span style='mso-spacerun:yes'>
?</span>
2. 5690<span style='mso-spacerun:yes'>
??????????? </span>
0. 000000<spanstyle='mso-spacerun:yes'>
? </span>
0. 00000<span style='mso-spacerun:yes'>
?</span>
2 -23. 956 51. 9 18. 61<span style='mso-spacerun:yes'>
? </span>
0. 000</span>
 **## 3<span style='mso-spacerun:yes'>
?</span>
2. 5300<span style='mso-spacerun:yes'>
????????? </span>
+ 0. 004823-0. 03664<span style='mso-spacerun:yes'>
? </span>
3 -23. 893 53. 8 20. 48<spanstyle='mso-spacerun:yes'>
? </span>
0. 000</span>
 **## Models ranked by AIC(x)</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


And fromthe second model:

</span>


```{r}



```


data=</span>

bm)</span>
 **dredge</span>
</span>
(bd2, </span>
</span>
rank=</span>
</span>
"AIC"</span>
</span>
, </span>
</span>
extra=</span>
</span>

 </span>
</span>
c</span>
</span>
(</span>
</span>
"R^2"</span>
</span>
, </span>
</span>
adjRsq=</span>
</span>
function(x)</span>
</span>

summary</span>
</span>
(x)$adj. r. squared))</span>
</span>


</span>

## Globalmodel call: lm(formula = logSpecies ~ logDBFs + TJK, data = bm)</span>
 **## ---</span>
 **## Model selection table </span>
 **##<span style='mso-spacerun:yes'>
??</span>
(Intrc)<span style='mso-spacerun:yes'>
? </span>
lgDBF TJK<spanstyle='mso-spacerun:yes'>
????? </span>
R^2<span style='mso-spacerun:yes'>
??</span>
adjRsq df<span style='mso-spacerun:yes'>
? </span>
logLik<spanstyle='mso-spacerun:yes'>
? </span>
AIC delta weight</span>
 **## 4 -2. 4100 1. 3710<span style='mso-spacerun:yes'>
??</span>
+ 0. 519900<span style='mso-spacerun:yes'>
? </span>
0. 47810<spanstyle='mso-spacerun:yes'>
? </span>
4 -14. 418 36. 8<spanstyle='mso-spacerun:yes'>
? </span>
0. 00<span style='mso-spacerun:yes'>
?</span>
0. 995</span>
 **## 2<span style='mso-spacerun:yes'>
? </span>
0. 59640. 4882<span style='mso-spacerun:yes'>
???? </span>
0. 209800<spanstyle='mso-spacerun:yes'>
? </span>
0. 17690<span style='mso-spacerun:yes'>
?</span>
3 -20. 895 47. 8 10. 95<span style='mso-spacerun:yes'>
? </span>
0. 004</span>
 **## 1<span style='mso-spacerun:yes'>
?</span>
2. 5690<span style='mso-spacerun:yes'>
??????????? </span>
0. 000000<spanstyle='mso-spacerun:yes'>
? </span>
0. 00000 <spanstyle='mso-spacerun:yes'>
?</span>
2 -23. 956 51. 9 15. 08<spanstyle='mso-spacerun:yes'>
? </span>
0. 001</span>
 **## 3<span style='mso-spacerun:yes'>
?</span>
2. 5300<span style='mso-spacerun:yes'>
????????? </span>
+ 0. 004823-0. 03664<span style='mso-spacerun:yes'>
? </span>
3 -23. 893 53. 8 16. 95<spanstyle='mso-spacerun:yes'>
? </span>
0. 000</span>
 **## Models ranked by AIC(x)

</span>



style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The top AIC model is </span>
<!--[if gte msEquation 12]>
<m:oMath>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
 "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image047. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image048. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
with an AIC of 33. 3. The next best model was </span>
<!--[if gte msEquation 12]>
<m:oMath>
<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
 "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image049. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image050. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
with an AIC of 36. 8, so 3. 5 AIC units worsethan the top model. We put these two runs of results together in Table 8-1, re-computing
all the </span>

AICs based on the top model from the
first full model considered. </span>

<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>


</span>

style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


style='font-size:9. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>

&nbsp;
</span>


*<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;mso-ascii-theme-font:major-latin;
Table 8-1: Model comparison table. 

</span>
*

<table class=MsoNormalTable border=1 cellspacing=0 cellpadding=0
 mso-yfti-tbllook:1184;mso-padding-alt:0in 5. 4pt 0in 5. 4pt;mso-border-insideh:

 <tr style='mso-yfti-irow:0;mso-yfti-firstrow:yes;height:15. 0pt'>

 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
Model

</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
R<sup>
2</sup>


</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
adjR<sup>
2</sup>


</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
df

</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
logLik

</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri; mso-fareast-font-family:"Times New Roman";mso-hansi-font-family:Calibri;
 mso-bidi-font-family:"Times New Roman";color:black'>
AIC

</span>

 </td>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;line-height: normal'>
<span style='font-size:12. 0pt;font-family:Symbol;mso-ascii-font-family: Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
 mso-hansi-theme-font:minor-latin;mso-char-type:symbol;mso-symbol-font-family:
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
AIC

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:1;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;   mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
   font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:247. 15pt;height:16. 5pt'>

  <v:imagedata src="chapter8_files/image059. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=330 height=22 src="chapter8_files/image060. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 5809

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 5444

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
4

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-12. 652

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
33. 3

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:2;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;   mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
   font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sub>
</m:sSub>
<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:247. 15pt;height:16. 5pt'>

  <v:imagedata src="chapter8_files/image061. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=330 height=22 src="chapter8_files/image062. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 5199

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 4781

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
4

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-14. 418

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
36. 8

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
3. 5

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:3;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;   mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
   font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:195. 4pt;height:15pt'>

  <v:imagedata src="chapter8_files/image063. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=261 height=20 src="chapter8_files/image064. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 3691

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 3428

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
3

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-17. 969

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
41. 9

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
8. 6

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:4;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;   mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
   font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:195. 4pt;height:15pt'>

  <v:imagedata src="chapter8_files/image065. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=261 height=20 src="chapter8_files/image066. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 2098

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 1769

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
3

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-20. 895

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
47. 8

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
14. 5

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:5;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:113. 65pt;height:15pt'>

  <v:imagedata src="chapter8_files/image067. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=152 height=20 src="chapter8_files/image068. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
2

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-23. 956

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
51. 9

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
18. 6

</span>

 </td>

 </tr>

 <tr style='mso-yfti-irow:6;mso-yfti-lastrow:yes;height:15. 0pt'>

 padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal style='margin-bottom:0in;margin-bottom:. 0001pt;text-align: justify;line-height:normal'>
<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<span  style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;  font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSub>
<m:sSubPr>
   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
</m:oMathPara>
<![endif]-->
<![if !msEquation]>
<span style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
 mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
 minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
 mso-bidi-theme-font:minor-bidi;mso-ansi-language:EN-US;mso-fareast-language:
 EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
  type="#_x0000_t75" style='width:165pt;height:16. 5pt'>

  <v:imagedata src="chapter8_files/image069. png" o:title="" chromakey="white"/>

 </v:shape>
<![endif]-->
<![if !vml]>
<img width=220 height=22 src="chapter8_files/image070. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<span style='font-size:12. 0pt;mso-ascii-font-family:Calibri;mso-fareast-font-family:
 "Times New Roman";mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>


</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
0. 0048

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-0. 03664

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
3

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
-23. 893

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
53. 8

</span>

 </td>

 <td nowrap valign=bottom style='border-top:none;border-left:none;border-bottom:
 solid windowtext . 5pt;padding:0in 5. 4pt 0in 5. 4pt;height:15. 0pt'>

 <p class=MsoNormal align=right style='margin-bottom:0in;margin-bottom:. 0001pt; text-align:right;line-height:normal'>
<span style='font-size:12. 0pt; mso-ascii-font-family:Calibri;mso-fareast-font-family:"Times New Roman";
 mso-hansi-font-family:Calibri;mso-bidi-font-family:"Times New Roman";
 color:black'>
20. 5

</span>

 </td>

 </tr>

</table>



&nbsp;
</span>

Table 8-1 suggests some interestingleads to the worst performing model on the AIC measure, ranking below a modelwith nothing in it (mean-only) and 20. 5 AIC units worse than the top model. But
the two top models distinctly benefit from the inclusion of *TJK*. This suggests that aftercontrolling for the sampling effort, either through *DBC* or *DBF*, the differencesin the stages captured by *TJK* can be moreclearly observed. 

</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";the top model in our (correct) results<a style='mso-footnote-id:ftn6'href="#_ftn6" name="_ftnref6" title="">
<sup>
<![if !supportFootnotes]>
<sup>
<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</sup>
<![endif]>
</span>
</sup>
</a>

different intercepts for the two periods. We can interrogate this model furtherbut probably should do something we should have done first, check the
diagnostics (Figure 8-11) and consider our model assumptions as AICs are not
valid if model assumptions are not met. 

</span>


```{r}



```


c</span>

(</span>
2</span>
, </span>
2</span>
), </span>
oma=</span>
(</span>

0</span>
, </span>
0</span>
, </span>
2</span>
, </span>
0</span>
))</span>
 **plot</span>
(bd1)</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75" style='width:391. 9pt;height:244. 9pt;visibility:visible;
 mso-wrap-style:square'>

 <v:imagedata src="chapter8_files/image071. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=523 height=327src="chapter8_files/image072. jpg" v:shapes="Picture">
<![endif]>
</span>



*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 8-11: Diagnostic plots
for the top AIC model. </span>
*

</span>


&nbsp;
</span>

????????????? </span>
Theconstant variance and assessment of influence do not suggest any real problems. 
The normality assumption is possibly violated but shows lighter tails than
expected from a normal distribution and so should cause few problems with
inferences (we would be looking for an answer of "yes, there is a violation of
the normality assumption but, for a bonus point, that problem is minor because
the pattern is not the problematic type of violation because?"). The other
assumption that  ****  is that the observations are independent. Between neighboring stagesin time, there would likely be some sort of relationship in the biodiversity so
we should not assume that the observations are independent (this is a  ***timeseries***  of observations). The authors acknowledged this issue but
unskillfully attempted to deal with it. Because a</span>

12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
n interaction was not considered in any of the models, there
also is an assumption that the results are parallel enough for the two groups. 
The scatterplot in Figure 8-10 suggests that using parallel lines for the two
groups is probably reasonable but a full assessment really should also explore
that fully to verify that there is no support for an interaction. 

</span>


????????????? </span>
Ignoringthe violation of the independence assumption, we are otherwise ok to explore
the model more and see what it tells us about biodiversity of *Sauropodomorphs*. The top model is  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math"'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
</m:acc>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=-1. 089+0. 724</m:r>
</span>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr  m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height:  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:277. 15pt;height:17. 65pt'>

 <v:imagedata src="chapter8_files/image073. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=370 height=24src="chapter8_files/image074. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. This suggests that for the early
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math"'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
</m:acc>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=-1. 089+0. 724</m:r>
</span>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr  m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:215. 25pt;height:16. 5pt'>

 <v:imagedata src="chapter8_files/image075. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=287 height=22src="chapter8_files/image076. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and for the Cretaceous period (<i  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math"'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
</m:acc>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=-1. 089-0. 75+0. 724</m:r>
</span>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr  m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:252pt;height:16. 5pt'>

 <v:imagedata src="chapter8_files/image077. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=336 height=22src="chapter8_files/image078. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
which simplifies to </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math"'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr   m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:   12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:   "Cambria Math", serif;mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:
   "Cambria Math";font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
</m:e>
</m:acc>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=-1. 84+0. 724</m:r>
</span>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr  m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:208. 5pt;height:16. 5pt'>

 <v:imagedata src="chapter8_files/image079. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=278 height=22src="chapter8_files/image080. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. This suggests that the sampling efforts
have the same effect on all observations and having an increase in *logDBC*s is associated with increases inSpecifically, for a 1 log-count increase in the *log-DBCs*, we expect, on average, to have a 0. 724 log-count change
in the mean log-biodiversity, after accounting for different intercepts for the
two periods considered. We could also translate this to the original count
scale but will leave it as is, because their real question of interest involves
the differences between the periods. The change in the y-intercepts of -0. 76
suggests that the Cretaceous has a lower average log-biodiversity by 0. 75
log-count, after controlling for the log-sampling effort. This suggests that
the *Cretaceous* had a lower correctedmean log-Sauropodomorph biodiversity (*t*<sub>
23</sub>
=-3. 41;<s>
p-value=0. 0024</s>
) than the combined results for the Triassic and Jurassic. 
On the original count scale, this suggests exp(-0. 76)=0. 47 times (53% drop in)
the median biodiversity count per stage for Cretaceous versus the prior time
period, after correcting for log-sampling effort in each stage. </span>



</span>

```{r}



```





##<spanstyle='mso-spacerun:yes'>
???????????? </span>
Estimate Std. Error t valuePr(&gt;|t|)<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## (Intercept)<span style='mso-spacerun:yes'>
?</span>
-1. 0887<span style='mso-spacerun:yes'>
???? </span>
0. 6533<spanstyle='mso-spacerun:yes'>
? </span>
-1. 666<span style='mso-spacerun:yes'>
??</span>
0. 1092<span style='mso-spacerun:yes'>
??? </span>
</span>
 **## logDBCs<span style='mso-spacerun:yes'>
??????</span>
0. 7243<span style='mso-spacerun:yes'>
???? </span>
0. 1288<spanstyle='mso-spacerun:yes'>
?? </span>
5. 622 1. 01e-05 ***</span>
 **## TJK1<span style='mso-spacerun:yes'>
????????</span>
-0. 7598<span style='mso-spacerun:yes'>
???? </span>
0. 2229<spanstyle='mso-spacerun:yes'>
? </span>
-3. 409<span style='mso-spacerun:yes'>
??</span>
0. 0024 ** </span>
 **## </span>
 **## Residual standard error: 0. 4185 on 23 degrees offreedom</span>
 **## Multiple R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 5809, Adjusted R-squared:<span style='mso-spacerun:yes'>
?</span>
0. 5444 </span>
 **## F-statistic: 15. 94 on 2 and 23 DF, <spanstyle='mso-spacerun:yes'>
? </span>
p-value: 4. 54e-05</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


????????????? </span>
Theirstudy shows some interesting contrasts between methods. They tried to use
AIC-based model selection methods across all the models but then used p-values
to really make their final conclusions. This presents a philosophical
inconsistency that bothers some more than others but should bother everyone. 
One thought is whether they needed to use AICs at all since they wanted to use
p-values? The one reason they might have preferred to use AICs is that it
 style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image047. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image048. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
to</span>
<!--[if gte msEquation 12]>
<m:oMath>
<span style='font-size:14. 0pt; mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
  m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family: "Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="p"/>
</m:rPr>
log&#8289;</m:r>
<m:r>
*(*</m:r>
<m:r>
<i style='mso-bidi-font-style:  mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#946;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
I</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
TJK</m:r>
<m:r>
, </m:r>
<m:r>
i</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
&#949;</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:288. 75pt;height:19. 15pt'>

 <v:imagedata src="chapter8_files/image049. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=385 height=26src="chapter8_files/image050. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, exploring whether *DBC* or *DBF* is the "better"There is no hypothesis test to compare these two models because one is not  ***nested*** in the other ?  **it is not possible to getfrom one model to the other by setting one or more slope coefficients to 0 so
** . The AICs suggest
that the model with *DBC* and *TJK* is better than the model with *DBF* and *TJK*, so that helps us make that decision. After that step, we couldrely on *t*-tests or ANOVA *F*-tests to decide whether furtherrefinement is suggested/possible for the model with *DBC* and *TJK*. This wouldprovide the direct inferences that they probably want and are trying to obtain
from AICs along with p-values in their paper. </span>



</span>

????????????? </span>
Finally, their results would actually be more valid if they had used a set of
statistical methods designed for modeling responses that are counts of events
or things, especially those whose measurements change as a function of sampling
effort; models called  ***Poisson rate models***  would be idealfor their application. The other aspect of the biodiversity that they measured
for each stage was the duration of the stage. They never incorporated that
information and it makes sense given their interests in comparing biodiversity
across stages, not understanding why more or less biodiversity might occur. But
other researchers might want to estimate the biodiversity after also
controlling for the length of time that the stage lasted and the sampling
efforts involved in detecting the biodiversity of each stage, models that are
only a few steps away from those considered here. In general, this study
presents some of the pitfalls of attempting to use advanced statistical methods
as well as hinting at the benefits. The statistical models are the only way to
access the results of interest; inaccurate usage of statistical models can
provide inaccurate conclusions. They seemed to mostly get the right answers
despite a suite of errors in their work. 

</span>


EN-US;mso-bidi-language:AR-SA'>
 **</span>


<h2 style='margin-left:0in;text-indent:0in;mso-list:l0 level2 lfo1'>
<a
name="_Toc459020970">
8. 4: General summary</a>


</h2>


????????????? </span>
Aswe wrap up, it is important to remember that these tools are limited by the
quality of the data collected. If you are ever involved in applying these
statistical models, whether in a research or industrial setting, make sure that
the research questions are discussed before data collection. And before data
collection is started, make sure that the methods will provide results that can
address the research questions. And, finally, make sure someone involved in the
project knows how to perform the appropriate statistical analysis. One way to
make sure you know how to analyze a data set and, often, clarify the research
questions and data collection needs, is to  **** . With this sort of preparation, many issuescan be avoided. Remember to think about reasons why assumptions of your
proposed method might be violated. 

</span>


????????????? </span>
Youare now  ****  and a bit  ****  with statistical methods. Ifyou go to use them, remember the fundamentals and find the story in the data. 
After deciding on any research questions of interest, graph the data and make
sure that the statistical methods will give you results that make some sense
based on the graphical results. In the MLR results, it is possible that graphs
will not be able to completely tell you the story, but all the other methods
should follow the pictures you see. Even when (or especially when) you use
sophisticated statistical methods, graphical presentations are critical to
helping others understand the results. We have discussed examples that involve
displaying categorical and quantitative variables and even some displays that
bridge both types of variables. We hope you have enjoyed this material and been
able to continue to develop your interests in statistics. You will see it in
many future situations both in courses in your area of study and in real
problems that need answers. You are also prepared to take more advanced
statistics courses - if you want to discuss the next options, we are happy to
provide additional information about the best next steps in learning
statistics. 

</span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>

The researchers did not do this analysis so never directly addressed this
research question although they did discuss it in general ways. 



</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>
 <spanstyle='font-size:11. 0pt;mso-bidi-font-size:12. 0pt'>
Instructors in this classoften get asked what a problem with non-constant variance actually looks like -
this is it!</span>

</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>

This was not even close to their top AIC model so they made an odd choice. 



</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>
 I
had students read this paper in a class and one decided that this was a
reasonable way to report small p-values ? it is WRONG. We are interested in how
small a p-value might be and saying it is over a value is never useful. 



</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>

All too often, I read journal articles that have under-utilized, 
under-reported, mis-applied, or mis-interpreted statistical methods and
results. One of the reasons that I wanted to write this book is to help more
people move from basic statistical knowledge to correct use of intermediate
statistical methods and beginning to see the potential in more advanced
statistical methods. It has taken me many years of being a statistician just to
feel armed for battle when confronted with new applications and two stat
courses are not enough to get you there, but you have to start somewhere. You
are only a couple of hundred hours into your 10, 000 hours required for mastery. 
This presentation is intended get you some solid fundamentals to build on or a
few intermediate tools to use if this is your last statistics training
experience. 



</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>
 <spanstyle='mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
Theyalso had an error in their AIC results that is difficult to explain here but
was due to an un-careful usage of the results from the more advanced models
that account for autocorrelation, which seems to provide the proper ranking of
models (*that they ignored*) but didnot provide the correct differences among models. </span>



</div>


</div>


** 

</html>

