Figure6-6 provided a term-plot of the estimated regression line and a shaded area surrounding
the estimated regression equation. Those shaded areas are based on connecting
the dots on 95% confidence intervals constructed for the true mean *y* value across all the x-values. Toformalize this idea, consider a specific value of *x*, and call it </span>
 ***x<sub>
&#957;</sub>
</span>
***  (pronounced  **** <a style='mso-footnote-id:ftn8' href="#_ftn8" name="_ftnref8"style='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
[8]</span>
</sup>
<![endif]>
</span>
</sup>
</a>
). 
Then the true mean response for this  **s*ubpopulation***  (a subpopulation is allobservations we could obtain at </span>

x</span>
= </span>
x<sub>
&#957;</sub>
</span>
) is given by </span>
 **E(Y) = </span>
**  ***<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;
symbol;mso-symbol-font-family:Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
m</span>
</span>
***  ***<sub>
 </span>
</sub>
***  **<sub>
&#957;</span>
</sub>
**  **= </span>
**  **<i12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:
minor-latin;mso-char-type:symbol;mso-symbol-font-family:Symbol'>
<spanstyle='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
b</span>
</span>
</i>
**  ***<sub>
0</span>
</sub>
***  ** + </span>
**  **<i12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:
minor-latin;mso-char-type:symbol;mso-symbol-font-family:Symbol'>
<spanstyle='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
b</span>
</span>
</i>
**  ***<sub>
1</span>
</sub>
***  ***x<sub>
&#957;</sub>
</span>
*** . To estimate the mean response at </span>
 ***x<sub>
&#957;</sub>
</span>
*** , we plug </span>
 ***x<sub>
&#957;</sub>
</span>
***  into the estimated regression
equation:

</span>


<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  "Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
0</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:89. 65pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image153. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=120 height=24src="chapter6_files/image154. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
. 

</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
To form the confidence interval, we
appeal to our standard formula of </span>
<!--[if gte msEquation 12]>
<m:oMath>
 ***<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
 mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
  m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
estimate</m:r>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
t</m:r>
</span>
*** </m:e>
<m:sub>
</m:sub>
<m:sup>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
*</m:r>
</span>
*** </m:sup>
</m:sSubSup>
 ***<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
 mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
E</m:r>
</span>
*** </m:e>
<m:sub>
 ***<span  style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif;
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:132pt;height:15. 75pt'>

 <v:imagedata src="chapter6_files/image155. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=176 height=21src="chapter6_files/image156. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
. </span>
** <span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
The  ***standarderror for the estimated mean at any x-value*** , denoted </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 4pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image157. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image158. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, can be calculated as

</span>


<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:14. 0pt;    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
<m:sup>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSubSup>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;  mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#963;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
</m:num>
<m:den>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
<m:r>

style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
 type="#_x0000_t75" style='width:171. 75pt;height:51pt'>

 <v:imagedata src="chapter6_files/image159. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=229 height=68src="chapter6_files/image160. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>


</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#963;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:15. 4pt;height:18pt'>

 <v:imagedata src="chapter6_files/image161. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=21 height=24src="chapter6_files/image162. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
is the squared residual standard error. This formula combines
  style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
SE</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
<m:sup>
</m:sup>
</m:sSubSup>
</m:oMath>
<![endif]-->
<![if !msEquation]>
<spanstyle='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:22. 9pt;height:18. 75pt'>

 <v:imagedata src="chapter6_files/image163. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=31 height=25src="chapter6_files/image164. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 9pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image165. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image166. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
from </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:acc>
<m:accPr>
<m:chr  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:7. 9pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image167. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=11 height=24src="chapter6_files/image168. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and the variability around the regression  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#963;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:15. 4pt;height:18pt'>

 <v:imagedata src="chapter6_files/image161. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=21 height=24src="chapter6_files/image162. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. Fortunately, we?ll use R?s </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
predict</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 function to provide these results for us
and avoid doing this calculation by hand most of the time. The  ***confidenceinterval for*** </span>
 ***<span style='font-size:12. 0pt;line-height:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
$\mu$</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:15pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image169. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=20 height=24src="chapter6_files/image170. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 9pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image165. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image166. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
, </span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
is

</span>


<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
$\mu$</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
&#957;</m:r>
</span>
*** </m:sub>
</m:sSub>
 ***<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
 "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
t</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
n</m:r>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
-</m:r>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
2</m:r>
</span>
*** </m:sub>
<m:sup>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
*</m:r>
</span>
*** </m:sup>
</m:sSubSup>
 ***<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:
 "Cambria Math", serif;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***    font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
     m:val="bi"/>
</m:rPr>
$\mu$</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sub>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:87pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image171. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=116 height=26src="chapter6_files/image172. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
. 

</span>
** 
style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
In application, these intervals get wider
the further we go from the mean of the x?s. These have interpretations that are
exactly like those for the y-intercept: 

</span>


<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
x</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:14. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image173. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image174. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, we are  **** % confident that the true mean of  **** is between  ****  and  ****   **** . 

</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
It is also useful to remember that this
interpretation applies individually to every *x* displayed in term-plots. </span>


</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";second type of interval in this situation takes on a more challenging task ? toplace an interval on where we think a new observation will fall, called a  ***predictioninterval*** (PI). This PI will need to be much wider than the CI for the
mean since we need to account for both the uncertainty in the mean and the
randomness in sampling a new observation from the normal distribution centered
at the true mean for </span>

x<sub>
&#957;</sub>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
. The interval is centered at the estimated regression line
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 15pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image175. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=18 height=24src="chapter6_files/image176. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
tohelp us see that this interval is for a  **** * y* at this x-value. The </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 4pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image177. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image178. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
incorporates the core of the previous SE calculation and adds in
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
&#963;</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sup>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:15. 4pt;height:18. 4pt'>

 <v:imagedata src="chapter6_files/image179. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=21 height=25src="chapter6_files/image180. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
:</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>


</span>


<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
*<span style='font-size:14. 0pt;    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
b</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
1</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
<m:sup>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSubSup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
-</m:r>
</span>
*<m:acc>
<m:accPr>
<m:chr
   mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:   "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
   mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#963;</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sup>
*<span style='font-size:14. 0pt;    mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
</m:num>
<m:den>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
n</m:r>
<m:r>

  style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***    font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
     m:val="bi"/>
</m:rPr>
&#963;</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sup>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
2</m:r>
</span>
*** </m:sup>
</m:sSup>
</m:e>
</m:rad>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:radPr>
<m:deg>
</m:deg>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;     mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:14. 0pt;   mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:14. 0pt;mso-bidi-font-size:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***    font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
     m:val="bi"/>
</m:rPr>
&#963;</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sup>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-13. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:306. 75pt;height:37. 15pt'>

 <v:imagedata src="chapter6_files/image181. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=409 height=50src="chapter6_files/image182. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
. 

</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
The __% PI is calculated as

</span>


<!--[if gte msEquation 12]>
<m:oMathPara>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
     m:val="bi"/>
</m:rPr>
y</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
&#957;</m:r>
</span>
*** </m:sub>
</m:sSub>
 *** font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
 mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubSupPr>
<m:e>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
t</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
n</m:r>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
-</m:r>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
2</m:r>
</span>
*** </m:sub>
<m:sup>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
    m:val="bi"/>
</m:rPr>
*</m:r>
</span>
*** </m:sup>
</m:sSubSup>
 *** font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
 mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
  mso-fareast-theme-font:minor-fareast'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:   12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***    font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
      m:val="bi"/>
</m:rPr>
y</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sub>
 ***   font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
EN-US;mso-bidi-language:AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025"
 type="#_x0000_t75" style='width:86. 25pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image183. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=115 height=26src="chapter6_files/image184. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>


</span>


style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
and interpreted as: 

</span>


<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
We are __% sure that a
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
x</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:14. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image173. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image174. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
will be between  ****  and  ****   **** . 

</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
Since </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 4pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image177. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image178. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
&gt;</span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif;mso-fareast-font-family:
  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif;mso-fareast-font-family:"Times New Roman";
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:30. 75pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image185. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=41 height=26src="chapter6_files/image186. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
, the  **** . Additionally, if either the SEfor the slope or the residual variance increases, both intervals would get
wider. 

</span>


<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
As in confidence
intervals, we assume that a 95% PI "succeeds" ? contains the new observation -
in 95% of applications of the methods and fails the other 5% of the time. 
Remember that for any interval estimate, the true value is either in the
interval or it is not and our confidence level essentially sets our failure
rate! Because PIs push into the tails of the assumed distribution of the
responses these methods are very sensitive to violations of assumptions so we
need to be especially certain that we have met the assumptions to trust these
results will work as advertised (at the  ***nominal***  (specified) level). 

</span>

????????????? </span>
Thereare two ways to explore CIs for the mean and PIs for a new observation. The
first is to focus on a specific x-value of interest. The second is to plot the
results for all x?s. To do both of these, but especially to make plots, we want
to learn to use the </span>

``predict `` function. It can either produce the estimate for a particular  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
x</m:r>
</span>
*</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:13. 9pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image165. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image166. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and the </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 75pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image187. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image188. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-fareast-font-family:
"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
</span>
line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
or we can get it to directly calculate the CI and PI. The first
way to use it is </span>

font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
predict(modelname, se. fit=T)</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 which will provide fitted values and </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size:  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 75pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image187. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image188. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
for all observed x?s. We can then use the </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 75pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image187. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image188. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
tocalculate </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 4pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image177. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image178. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
?</span>
and form our own PIs. If you want CIs, run </span>
style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast;
predict(modelname, interval= "confidence")</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
; if you want PIs, run </span>

style='font-size:12. 0pt;line-height:107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
predict(modelname, 
interval="prediction")</span>

<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
. If you want to do prediction at an x-value that was not in the original
observations, add the option </span>

107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
newdata=data. frame(XVARIABLENAMEFROMORIGINALMODEL=Xnu)</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 to the </span>

12. 0pt;line-height:107%;font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
predict</span>

style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
 function call. </span>



</span>

????????????? </span>
Someexamples of using the predict function follow<a style='mso-footnote-id:ftn9'
href="#_ftn9" name="_ftnref9" title="">
<sup>
<![if !supportFootnotes]>
<sup>
<span style='font-size:12. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;
minor-latin;mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi;
[9]</span>
</sup>
<![endif]>
</span>
</sup>
</a>
. 
For example, it might be interesting to use the regression model to find a 95%
CI and PI for the *Beers* vs *BAC* study for a student who wouldconsume 8 beers. Four different applications of the predict function follow. 
Note that </span>

``lwr`` and </span>
 in the output depend on what we
requested. The first use of </span>

``predict``  just returns the estimated mean for 8 beers:

</span>

```{r}



```


data=</span>

BB)</span>
 **?</span>
</span>
predict</span>
newdata=</span>

(</span>

8</span>

##<spanstyle='mso-spacerun:yes'>
???????? </span>
1 </span>
 **
&nbsp;
</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


By turning on the </span>

``se. fit=T`` option, we also get the SE for the confidence interval and degrees of freedom. Note that elements returned are labelled as </span>

``$fit`` ``$se. fit`` 
, etc. </span>
"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5'>


</span>


```{r}



```


data. frame</span>

Beers=</span>

), </span>

T)</span>




## $fit</span>
 **##<span style='mso-spacerun:yes'>
???????? </span>
1 </span>
 **## 0. 1310095 </span>
 **## </span>
 **## $se. fit</span>
 **## [1] 0. 009204354</span>
 **## </span>
 **## $df</span>
 **## [1] 14</span>
 **## </span>
 **## $residual. scale</span>
 **## [1] 0. 02044095

</span>


&nbsp;
</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Instead of using the components of the intervals to
make them, we can also directly request the CI or PI using the </span>

``interval=?``  option, as in the following two lines of code. 

</span>

```{r}



```


data. frame</span>

Beers=</span>

), </span>

"confidence"</span>

##<spanstyle='mso-spacerun:yes'>
???????? </span>
fit<spanstyle='mso-spacerun:yes'>
?????? </span>
lwr<span style='mso-spacerun:yes'>
??????</span>
upr</span>
 **?</span>
</span>
(m1, </span>

data. frame</span>

Beers=</span>

), </span>

"prediction"</span>

##<spanstyle='mso-spacerun:yes'>
???????? </span>
fit<spanstyle='mso-spacerun:yes'>
??????? </span>
lwr<spanstyle='mso-spacerun:yes'>
?????? </span>
upr</span>
 **</div>


``color:black;background:#E1E2E5'>

&nbsp;
`` 

Based on these results, we are 95% confident that the true mean *BAC*for 8 beers consumed is between 0. 111 and 0. 15 grams of alcohol per dL ofblood. For a new student drinking 8 beers, we are 95% sure that the observed
BAC will be between 0. 083 and 0. 179 g/dL. You can see from these results that
the PI is much wider than the CI ? it has to capture a new individual?s results
95% of the time which is much harder than trying to capture the mean. For
completeness, we should do these same calculations by hand. The </span>

``predict(?. , se. fit=T)``  output provides almost all of thepieces we need to calculate the CI and PI. The </span>

``$fit``  is the estimate = </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;mso-bidi-font-size:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:14. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image189. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image190. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
?</span>
</span>
= 0. 131, the </span>
``$se. fit``  is the SE for the estimate of the mean = </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:14. 0pt;mso-bidi-font-size: 12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:   style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;    line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:14. 0pt;style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:27. 75pt;height:19. 5pt'>

 <v:imagedata src="chapter6_files/image187. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=37 height=26src="chapter6_files/image188. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
=0. 0092</span>

, </span>
``$df``  is *n*-2 = 16-2=14, and </span>

<span  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;mso-bidi-font-size:12. 0pt;  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:14. 0pt;mso-bidi-font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:8. 65pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image191. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=12 height=24src="chapter6_files/image192. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
=0. 02044. So we just need the *t** multiplier for 95% confidence and 14 *df*:

</span>


```{r}



```


975</span>
, </span>
df=</span>
) </span>

#t* multiplier for 95% CI or95% PI</span>



## [1]2. 144787</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The 95% CI for the true mean at </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
  normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
   m:val="bi"/>
</m:rPr>
x</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:14. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image173. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image174. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **<span style='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
?</span>
= 8</span>
**  is then:

</span>

```{r}



```


c</span>
(-</span>
1</span>
, </span>
1</span>
)*</span>
*</span>

## [1]0. 1112678 0. 1507322</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Which matches the previous outputquite well. 

</span>


The 95% PI requires the calculationof </span>
<!--[if gte msEquation 12]>
<m:oMath>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
S</m:r>
</span>
*<m:sSub>
<m:sSubPr>
  font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
  mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
E</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
    normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
y</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
&#957;</m:r>
</span>
*</m:sub>
</m:sSub>
</m:sub>
</m:sSub>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:radPr>
<m:deg>
</m:deg>
<m:e>
*<span style='font-size:12. 0pt;line-height:   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
E</m:r>
</span>
*</m:e>
<m:sub>
<m:sSub>
<m:sSubPr>
    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:     12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
     mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
     normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
*<span style='font-size:12. 0pt;     line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
$\mu$</m:r>
</span>
*</m:e>
</m:acc>
</m:e>
<m:sub>
*<span style='font-size:12. 0pt;   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSup>
<m:sSupPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:    12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
    mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
    normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:accPr>
<m:e>
 ***<span    style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
     m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
&#963;</m:r>
</span>
*** </m:e>
</m:acc>
</m:e>
<m:sup>
 ***<span   style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
2</m:r>
</span>
*** </m:sup>
</m:sSup>
</m:e>
</m:rad>
*<span style='font-size:12. 0pt;line-height: 107%;font-family:"Cambria Math", serif'>
<m:r>
=</m:r>
</span>
*<m:rad>
<m:radPr>
<m:degHide
  mso-bidi-font-size:12. 0pt;font-family:"Cambria Math", serif;mso-ascii-font-family:  "Cambria Math";mso-hansi-font-family:"Cambria Math";font-style:italic;
  mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:radPr>
<m:deg>
</m:deg>
<m:e>
*<span style='font-size:12. 0pt;line-height:   style='font-size:12. 0pt;mso-ansi-font-size:12. 0pt;mso-bidi-font-size:12. 0pt;   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-style:italic;mso-bidi-font-style:
   normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
)</m:r>
</span>
*</m:e>
<m:sup>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
2</m:r>
</span>
*</m:sup>
</m:sSup>
*<span style='font-size:12. 0pt;line-height:  107%;font-family:"Cambria Math", serif'>
<m:r>
+</m:r>
</span>
*<m:sSup>
<m:sSupPr>
   font-family:"Cambria Math", serif;mso-ascii-font-family:"Cambria Math";
   mso-hansi-font-family:"Cambria Math";font-weight:bold;mso-bidi-font-weight:
   normal;font-style:italic;mso-bidi-font-style:normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSupPr>
<m:e>
*<span style='font-size:12. 0pt;   line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
0. 02044</m:r>
</span>
*</m:e>
<m:sup>
 ***<span   style='font-size:12. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
    m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
2</m:r>
</span>
*** </m:sup>
</m:sSup>
</m:e>
</m:rad>
*<span style='font-size:12. 0pt;line-height:style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
-11. 5pt;mso-ansi-language:EN-US;mso-fareast-language:EN-US;mso-bidi-language:
AR-SA'>
<!--[if gte vml 1]>
<v:shape id="_x0000_i1025" type="#_x0000_t75"
 style='width:264. 75pt;height:30. 4pt'>

 <v:imagedata src="chapter6_files/image193. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=353 height=41src="chapter6_files/image194. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
<spanstyle='font-size:12. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";
0. 0224. 

</span>


```{r}



```


^</span>

2</span>
+0. 02044</span>
^</span>
2</span>
)</span>



## [1]0. 02241503</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


style='font-size:12. 0pt;mso-bidi-font-size:11. 0pt;line-height:107%;mso-fareast-font-family:"Times New Roman";mso-fareast-theme-font:minor-fareast'>
The 95% PI </span>

at </span>
<!--[if gte msEquation 12]>
<m:oMath>
<m:sSub>
<m:sSubPr>
<span  style='font-size:14. 0pt;mso-ansi-font-size:14. 0pt;font-family:"Cambria Math", serif;
  mso-ascii-font-family:"Cambria Math";mso-hansi-font-family:"Cambria Math";
  font-weight:bold;mso-bidi-font-weight:normal;font-style:italic;mso-bidi-font-style:
  normal'>
<m:ctrlPr>
</m:ctrlPr>
</span>
</m:sSubPr>
<m:e>
 ***<span style='font-size:14. 0pt;  mso-bidi-font-size:11. 0pt;line-height:107%;font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr
   m:val="roman"/>
<m:sty m:val="bi"/>
</m:rPr>
x</m:r>
</span>
*** </m:e>
<m:sub>
 ***  font-family:"Cambria Math", serif'>
<m:r>
<m:rPr>
<m:scr m:val="roman"/>
<m:sty
style='font-size:11. 0pt;line-height:107%;font-family:"Calibri", sans-serif;mso-ascii-theme-font:minor-latin;mso-fareast-font-family:Calibri;mso-fareast-theme-font:
minor-latin;mso-hansi-theme-font:minor-latin;mso-bidi-font-family:"Times New Roman";
<!--[if gte vml 1]>
<v:shape
 id="_x0000_i1025" type="#_x0000_t75" style='width:14. 25pt;height:17. 65pt'>

 <v:imagedata src="chapter6_files/image173. png" o:title="" chromakey="white"/>

</v:shape>
<![endif]-->
<![if !vml]>
<img width=19 height=24src="chapter6_files/image174. png" v:shapes="_x0000_i1025">
<![endif]>
</span>
<![endif]>
 **?</span>
= 8</span>
**  is

</span>

```{r}



```


c</span>
(-</span>
1</span>
, </span>
1</span>
)*</span>
*</span>

## [1]0. 08295648 0. 17904352</span>



</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


These calculations are "fun"and informative but displaying these results for all x-values is a bit more
informative about the performance of the two types of intervals and for results
we might expect in this application. The calculations we just performed provide
To make this plot, we need to create a sequence of *Beers* values to get other results for, say from 0 to 10, using the </span>

``seq``  function. The </span>
``from``  and </span>
``to``  define the endpoints and the </span>
``length. out``  defines the resolution of the gridof equally spaced points to create. Here, 30 points is more than enough to make
the confidence and prediction invervals from 0 to 10 *Beers*. 

</span>

```{r}



```

</span>
seq</span>
</span>
(</span>
</span>
from=</span>
</span>
0</span>
</span>
, </span>
</span>
to=</span>
</span>
10</span>
</span>
, </span>
</span>
length. out=</span>
</span>
30</span>
</span>
)</span>
</span>


</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


Now we can call the </span>
``predict``  function at all these values storedin </span>

 to get the CIs for all these *Beers* values:

</span>

```{r}



```

</span>
data. frame</span>
</span>
(</span>
</span>
predict</span>
</span>
(m1, </span>
</span>
newdata=</span>
</span>
 </span>
</span>
data. frame</span>
</span>
(</span>
</span>
Beers=</span>
</span>
beerf), </span>
</span>
interval=</span>
</span>
"confidence"</span>
</span>
))</span>
</span>
 **head</span>
(BBCI)</span>


</span>

##<spanstyle='mso-spacerun:yes'>
???????????? </span>
fit<spanstyle='mso-spacerun:yes'>
????????? </span>
lwr<spanstyle='mso-spacerun:yes'>
??????? </span>
upr</span>
</span>

 **## 1 -0. 0127006040 -0. 039805351 0. 01440414</span>
 **## 2 -0. 0065062033 -0. 031996517 0. 01898411</span>
 **## 3 -0. 0003118027 -0. 024210653 0. 02358705</span>
 **## 4<span style='mso-spacerun:yes'>
?</span>
0. 0058825980 -0. 016452670 0. 02821787</span>
 **## 5<span style='mso-spacerun:yes'>
?</span>
0. 0120769986 -0. 008728854 0. 03288285</span>
 **## 6<span style='mso-spacerun:yes'>
?</span>
0. 0182713992 -0. 001047321 0. 03759012</span>


</span>


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


And the PIs:

</span>

```{r}



```

</span>
data. frame</span>
</span>
(</span>
</span>
predict</span>
</span>
(m1, </span>
</span>
newdata=</span>
</span>
data. frame</span>
</span>
(</span>
</span>
Beers=</span>
</span>
beerf), </span>
</span>
interval=</span>
</span>
"prediction"</span>
</span>
))</span>
</span>
 **head</span>
(BBPI)</span>


</span>

##<spanstyle='mso-spacerun:yes'>
???????????? </span>
fit<spanstyle='mso-spacerun:yes'>
???????? </span>
lwr<spanstyle='mso-spacerun:yes'>
??????? </span>
upr</span>
</span>

 **## 1 -0. 0127006040 -0. 06424420 0. 03884300</span>
 **## 2 -0. 0065062033 -0. 05721943 0. 04420702</span>
 **## 3 -0. 0003118027 -0. 05024406 0. 04962046</span>
 **## 4<span style='mso-spacerun:yes'>
? </span>
0. 0058825980-0. 04332045 0. 05508564</span>
 **## 5<span style='mso-spacerun:yes'>
?</span>
0. 0120769986 -0. 03645092 0. 06060492</span>
 **## 6<span style='mso-spacerun:yes'>
?</span>
0. 0182713992 -0. 02963777 0. 06618057</span>


</span>


</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


The rest of the code is just making ascatterplot and adding the five lines with a legend. 

</span>


```{r}



```


BB, </span>

"Beers"</span>

ylab=</span>

, </span>

20</span>

, </span>
col=</span>
"gold4"</span>
main=</span>

"Scatterplot ofestimated regression line with 95% CI and PI"</span>
)</span>
 **?</span>
</span>
lines</span>
col=</span>

, </span>

3</span>

)</span>
 **?</span>
</span>
lines</span>
col=</span>

, </span>

2</span>

, </span>
lwd=</span>
3</span>
)</span>
 **?</span>
</span>
lines</span>
col=</span>

, </span>

2</span>

, </span>
lwd=</span>
3</span>
)</span>
 **?</span>
</span>
lines</span>
col=</span>

, </span>

3</span>

, </span>
lwd=</span>
3</span>
)</span>
 **?</span>
</span>
lines</span>
col=</span>

, </span>

3</span>

, </span>
lwd=</span>
3</span>
)</span>
 **legend</span>
"topleft"</span>

c</span>

(</span>
"Estimate"</span>
, </span>
, </span>

), </span>

3</span>

, </span>
lty=</span>
c</span>
(</span>
1</span>
, </span>
2</span>
, </span>
3</span>
), </span>
col=</span>

 </span>
c</span>
(</span>
, </span>

, </span>

))</span>

</div>


<!--[if gte vml 1]>
<v:shape id="_x0000_i1026" type="#_x0000_t75" style='width:403. 15pt;height:252pt;
 visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter6_files/image195. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=538 height=336src="chapter6_files/image196. jpg" v:shapes="_x0000_i1026">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 6-22: Estimated SLR
for BAC data with 95% confidence (dashed lines) and 95% prediction (lighter, 
dotted lines) intervals. </span>
*

</span>


&nbsp;
</span>

More importantly, note that the CI inFigure 6-22 clearly shows widening as we move further away from the mean of the
x?s to the edges of the observed x-values. This reflects a decrease in
knowledge of the true mean as we move away from the mean of the x?s. The PI
also is widening slightly but not as clearly in this situation. The difference
in widths in the two types of intervals becomes extremely clear when they are
displayed together. </span>

font-family:"Lucida Console";mso-fareast-font-family:"Times New Roman";mso-bidi-font-family:"Courier New";color:black;background:#E1E2E5;mso-no-proof:
yes'>
 

</span>


Similarly, the 95% CI andPIs for the Bozeman yearly average maximum temperatures in Figure 6-23 provide
interesting information on the uncertainty in the estimated mean temperature
over time. It is also interesting to explore how many of the observations fall
within the 95% prediction intervals. The PIs are for new observations, but you
can see how the PIs that were constructed to contain almost all the
observations in the original data set but not all of them. In fact, only 2 of
the 109 observations (1. 8%) fall outside the 95% PIs. Since the PI needs to be
concerned with unobserved new observations it makes sense that it might contain
more than 95% of the observations used to make it. 

</span>


<!--[if gte vml 1]>
<v:shape id="Picture" o:spid="_x0000_i1025" type="#_x0000_t75" style='width:383. 65pt;
 height:239. 65pt;visibility:visible;mso-wrap-style:square'>

 <v:imagedata src="chapter6_files/image197. png" o:title=""/>

</v:shape>
<![endif]-->
<![if !vml]>
<img border=0 width=512 height=320src="chapter6_files/image198. jpg" v:shapes="Picture">
<![endif]>
</span>


</span>

*<spanstyle='font-size:12. 0pt;line-height:107%;font-family:"Calibri Light", sans-serif;
"Times New Roman";mso-bidi-theme-font:minor-bidi'>
Figure 6-23: Estimated SLR
for Bozeman temperature data with 95% confidence (dashed lines) and 95%
prediction (lighter, dotted lines) intervals. 

</span>
*


&nbsp;
</span>


We can also use thesesame methods to do a prediction for the year after the data set ended, 2015 and
in 2050:</span>




```{r}



```


data. frame</span>

Year=</span>

), </span>

"confidence"</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
fit<span style='mso-spacerun:yes'>
????</span>
lwr<span style='mso-spacerun:yes'>
????? </span>
upr</span>
 **## 1 58. 31967 57. 7019 58. 93744</span>



?</span>
</span>
(temp1, </span>

data. frame</span>

Year=</span>

), </span>

"prediction"</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
fit<span style='mso-spacerun:yes'>
?????</span>
lwr<span style='mso-spacerun:yes'>
????? </span>
upr</span>
 **## 1 58. 31967 55. 04146 61. 59787</span>



?</span>
</span>
(temp1, </span>

data. frame</span>

Year=</span>

), </span>

"confidence"</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
fit<span style='mso-spacerun:yes'>
?????</span>
lwr<span style='mso-spacerun:yes'>
????? </span>
upr</span>
 **## 1 60. 15514 59. 23631 61. 07397</span>



?</span>
</span>
(temp1, </span>

data. frame</span>

Year=</span>

), </span>

"prediction"</span>

##<spanstyle='mso-spacerun:yes'>
??????? </span>
fit<span style='mso-spacerun:yes'>
?????</span>
lwr<span style='mso-spacerun:yes'>
????? </span>
upr</span>
 **## 1 60. 15514 56. 80712 63. 50316

</span>


&nbsp;
</span>

</div>


style='font-size:3. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;mso-bidi-font-family:"Times New Roman"'>

&nbsp;
</span>


These results tell us that we are 95%confident that the true mean yearly average maximum temperature in 2015 is (or
was?) between 55. 04</span>

font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;mso-char-type:
symbol;mso-symbol-font-family:Symbol'>
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F and 61. 6</span>
107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;mso-hansi-theme-font:minor-latin;
<span style='mso-char-type:symbol;mso-symbol-font-family:Symbol'>
?</span>
</span>
F. And we are 95% sure that the observed yearlyaverage maximum temperature in 2015 will be (or would have been?) between 59. 2</span>

style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
F and 61. 1</span>

style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
F. Obviously, 2015 has occurred, but
since the data were not published when the data set was downloaded in July
2016, we can probably best treat 2015 as a potential "future" observation. The
results for 2050 are clearly for the future mean and a new observation in 2050. 
Note that up to 2014, no values of this response had been observed above 60</span>

style='font-size:12. 0pt;line-height:107%;font-family:Symbol;mso-ascii-font-family:Calibri;mso-ascii-theme-font:minor-latin;mso-hansi-font-family:Calibri;
F and the predicted mean in 2050 is
over 60</span>

F if the trend persists. It is also
easy to criticize the use of this model for 2050 because of its extreme amount
of extrapolation. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020944">
6. 7: Chapter summary</a>


</h2>


????????????? </span>
Inthis chapter, we raised our SLR modeling to a new level, considering inference
techniques for relationships between two variables. The next chapter will build
on these same techniques but add in additional explanatory variables for what
we call  ***multiple linear regression***  (MLR). For example, in the *Beers* vs *BAC* study, it would have been useful to control for the weight ofthe subjects since people of different sizes metabolize alcohol at different
rates and body size might explain some of the variability. We still would want
to study the effects of beer consumption but also control for the differences
in subject?s weights. Or if they had studied both male and female students, we
might need to change the slope or intercept for each sex, allowing the
relationship between *Beers* and *BAC* to change between these groups. Thatwill also be handled using MLR techniques but result in two simple linear regression
equations ? one for females and one for males. And in this chapter you learned
how to interpret SLR models. The next chapter will feel like it is completely
new initially but it actually contains very little new material, just more
complicated models that use the same concepts. There will be a couple of new
issues to consider for MLR and we?ll need to learn how to work with categorical
variables in a regression setting ? but we actually fit linear models with
categorical variables in Chapters 2 and 3 so that isn?t actually completely new
either. 

</span>


????????????? </span>
SLRis a simple (thus its name) tool for analyzing the relationship between two
quantitative variables. It contains assumptions about the estimated regression
line being reasonable and about the distribution of the responses around that
line to do inferences for the population regression line. Our diagnostic plots
help us to carefully assess those assumptions. If we cannot trust the
assumptions, then the estimated line and any inferences for the population are
un-trustworthy. Transformations can fix things so that we can use SLR to fit
regression models. Transformations can complicate the interpretations on the
original, untransformed scale but have minimal impact on the interpretations on
the transformed scale. It is important to be careful with the units of the
variables, especially when dealing with transformations, as this can lead to
big changes in the results depending on which scale (original or transformed) the
results are being interpreted on. 

</span>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020945">
6. 8: Important R code</a>


</h2>


The maincomponents of the R code used in this chapter follow with the components to
modify in red where y is a response variable, x is an explanatory variable, and
the data are in DATASETNAME. 

</span>


<ul style='margin-top:0in' type=disc>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``scatterplot(<span style='color:red'>
y`` ~<span   style='color:red'>
x</span>
, data=<span style='color:red'>
DATASETNAME</span>
, smooth=F)

</span>
</li>
 <ul style='margin-top:0in' type=circle>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides a scatter plot with a regression line. 

</span>
</li>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Require the </span>
``car``  package. 

</span>
</li>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Turn on </span>
``smooth=T``  to add a smoothing line to help detect nonlinear relationships. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``Modelname`` ``&lt;-lm(<span   style='color:red'>
y`` ~<span style='color:red'>
x</span>
, data=<span   style='color:red'>
DATASETNAME</span>
)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Estimates a regression model using least squares. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``summary(<span style='color:red'>
Modelname`` )

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides parameter estimates and R-squared (used heavily in Chapter   7 as well). 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``par(mfrow=c(2, 2)); plot(<span style='color:   red'>
Modelname`` )

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides four regression diagnostic plots in one plot. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``confint(<span style='color:red'>
Modelname`` , level=0. 95)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides 95% confidence intervals for the regression model   coefficients. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Change </span>
``level``  if you want   other confidence levels. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``plot(allEffects(<span style='color:red'>
Modelname`` ))

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides a term-plot of the estimated regression line with 95%   confidence interval for the mean. 

</span>
</li>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Requires the </span>
``effects``  package. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``DATASETNAME$logy`` ``&lt;-log(<span   style='color:red'>
DATASETNAME$y`` )

</span>
</li>

 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Creates a transformed variable called logy ? change this to be more   specific to your "y" or "x". 

</span>
</li>

 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``predict(<span style='color:red'>
Modelname`` , se. fit=T)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides fitted values for all observed x?s with SEs for the mean. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``predict(<span style='color:red'>
Modelname`` , newdata=data. frame(<span   style='color:red'>
x</span>
=<span style='color:red'>
XNEW</span>
), interval="confidence")

</span>
</li>
 <ul style='margin-top:0in' type=circle>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides fitted value for a specific x (XNEW) with CI for the mean.    Replace </span>

``x``  with name of   explanatory variable. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``predict(<span style='color:red'>
Modelname`` , newdata=data. frame(<span   style='color:red'>
x</span>
=<span style='color:red'>
XNEW</span>
), interval="prediction")

</span>
</li>
 <ul style='margin-top:0in' type=circle>
 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Provides fitted value for a specific x (XNEW) with PI for a new   observation. Replace </span>

``x``  with name of explanatory variable. 

</span>
</li>
 </ul>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level1 lfo3'>
``qt(. 975, df=<span style='color:red'>
n`` -2)

</span>
</li>
 <ul style='margin-top:0in' type=circle>

 <li class=MsoNormalCxSpMiddle style='margin-bottom:0in;margin-bottom:. 0001pt;
   mso-add-space:auto;mso-list:l0 level2 lfo3'>
Gets the t* multiplier for making a 95% interval with *n*-2 replaced by the sample size ?
   2. </span>


</li>

 </ul>

</ul>



&nbsp;
</span>

<h2 style='margin-left:0in;text-indent:0in;mso-list:l3 level2 lfo1'>
<a
name="_Toc459020946">
6. 9: Practice problems</a>


</h2>


We will continue with the treadmilldata set introduced in Chapter 0 and the SLR fit in the practice problems in
Chapter 5. The following code will get you back to where we stopped at the end
of Chapter 5:

</span>


style='font-size:10. 0pt;mso-bidi-font-size:12. 0pt;line-height:107%;font-family:"Lucida Console"'>
treadmill&lt;-read. csv("http://www. math. montana. edu/courses/s217/documents/treadmill. csv", header=T)

</span>


``plot(TreadMillOx~RunTime, data=treadmill)

`` 
``tm1&lt;-lm(TreadMillOx~RunTime, data=treadmill)

`` 
``summary(tm1)

`` 
6. 1. Use the output to test for a
linear relationship between treadmill oxygen and run time, writing out all 6+
steps of the hypothesis test. Make sure to address scope of inference and
interpret the p-value. 

</span>


6. 2. Form and interpret a 95%
confidence interval for the slope coefficient "by hand" using the provided
multiplier:

</span>


```{r}



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
```

</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
(. </span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
975</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
, </span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
df=</span>
</span>
<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
29</span>
</span>

<span style='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
)</span>
</span>



<spanstyle='mso-bidi-font-family:"Times New Roman";mso-bidi-theme-font:minor-bidi'>
##[1] 2. 04523</span>
</span>




</div>


``
&nbsp;
`` 
6. 3. Use the </span>

``confint``  function to find a similarconfidence interval, checking your previous calculation. 

</span>


6. 4. Use the predict function to find
fitted values, 95% confidence and 95% prediction intervals for run times of 11
and 16 minutes. 

</span>


6. 5. Interpret the CI and PI for the
11 minute run time. 

</span>


6. 6. Compare the width of either set
of CIs and PIs ? why are they different? For the two different predictions, why
are the intervals wider for 16 minutes than for 11 minutes?

</span>


6. 7. The Residuals vs Fitted plot
considered in Chapter 5 should have suggested slight non-constant variance and
maybe a little missed nonlinearity. Perform a log-transformation of the
treadmill oxygen response variable and re-fit the SLR model. Remake the
diagnostic plots and discuss whether the transformation changed any of them. 

</span>


Summarize the log(y)~x model and interpret the slope coefficient on thetransformed and original scales, regardless of the answer to the previous
question. </span>


</div>


<hr align=left size=1 width="33%">


<![endif]>


<div style='mso-element:footnote' id=ftn1>


<a style='mso-footnote-id:ftn1' href="#_ftnref1"name="_ftn1" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[1]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn2>


<a style='mso-footnote-id:ftn2' href="#_ftnref2"name="_ftn2" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[2]</span>
</span>
<![endif]>
</span>
</span>
</a>
  



</div>


<div style='mso-element:footnote' id=ftn3>


<a style='mso-footnote-id:ftn3' href="#_ftnref3"name="_ftn3" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[3]</span>
</span>
<![endif]>
</span>
</span>
</a>





</div>


<div style='mso-element:footnote' id=ftn4>


<a style='mso-footnote-id:ftn4' href="#_ftnref4"name="_ftn4" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[4]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn5>


<a style='mso-footnote-id:ftn5' href="#_ftnref5"name="_ftn5" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[5]</span>
</span>
<![endif]>
</span>
</span>
</a>
  



</div>


<div style='mso-element:footnote' id=ftn6>


<a style='mso-footnote-id:ftn6' href="#_ftnref6"name="_ftn6" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[6]</span>
</span>
<![endif]>
</span>
</span>
</a>
 



</div>


<div style='mso-element:footnote' id=ftn7>


<a style='mso-footnote-id:ftn7' href="#_ftnref7"name="_ftn7" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[7]</span>
</span>
<![endif]>
</span>
</span>
</a>

 



</div>


<div style='mso-element:footnote' id=ftn8>


<a style='mso-footnote-id:ftn8' href="#_ftnref8"name="_ftn8" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[8]</span>
</span>
<![endif]>
</span>
</span>
</a>

This silly nomenclature was adopted from De Veaux, Velleman, and Bock?s *Stats: Data and Models* text. If you findthis too cheesy, you can just call it x-vee. 



</div>


<div style='mso-element:footnote' id=ftn9>


<a style='mso-footnote-id:ftn9' href="#_ftnref9"name="_ftn9" title="">
<span style='mso-special-character:style='font-size:10. 0pt;line-height:107%;font-family:"Times New Roman", serif;mso-fareast-font-family:"Times New Roman";mso-ansi-language:EN-US;mso-fareast-language:
EN-US;mso-bidi-language:AR-SA'>
[9]</span>
</span>
<![endif]>
</span>
</span>
</a>
 I
have suppressed some of the plotting code in this and the last chapter to make
"pretty" pictures - which you probably are happy to not see it until you want
to make a pretty plot on your own. All the code used is available upon request. 



</div>


</div>


** 

</html>

